### 2023-06-19 chapter_2/src/appendix.qmd

# Chapter 2 Appendix {#sec-2-a}

```{r}
#| label: setup

# Load helpful functions
box::use(
  modelsummary[
    modelsummary
  ]
)

# Load fitted models and ppc checks
load(
  file = "../data/temp/models/study_1_fitted.RData"
)
```

## Measures

{{< pagebreak >}}

## Hypothesis 1

### Models

{{< pagebreak >}}

### Posterior Predictive Checks

{{< pagebreak >}}

## Hypothesis 2

### Models

```{r}
#| label: tbl-hypothesis-2-main
#| tbl-cap: Full results for reported models in $H_2$
# Redefine names
list_fitted_split_1 <- list(
  "Red treatment" = list_fitted[["h_2"]][["red"]]
  , "Blue treatment" = list_fitted[["h_2"]][["blue"]]
)

# Change goodness of fit stuff
gm <- list(
  list("raw" = "nobs", "clean" = "Num.Obs.", "fmt" = 1)
  , list("raw" = "r.squared", "clean" = "$R^2$", "fmt" = 2)
  , list("raw" = "waic", "clean" = "WAIC", "fmt" = 2)
)
# Make table
modelsummary(
  list_fitted_split_1
  , coef_map = c(
    "b_trial_1_red_stimuli" = "Red yard sign"
    , "b_trial_1_blue_stimuli" = "Blue yard sign"
    , "b_Intercept[1]" = "Intercept 1"
    , "b_Intercept[2]" = "Intercept 2"
    )
  , statistic = "conf.int"
  , gof_map = gm 
  , notes = c(
    "Data source: Pavlovia"
    , "Positive coefficients reflect the logged-odds"
    , "of claiming candidate is Republican."
    , "Numbers inside brackets reflect the 5th and 95th percentile"
    , "of the draws from the posterior distribution."
  )
  , escape = FALSE
  , output = "latex"
)
```

```{r}
#| label: tbl-interacted-models-h-2
#| tbl-cap: $H_2$ models moderated by age
# Redefine names
list_fitted_split_2 <- list(
  "Red treatment - Age moderating" = list_fitted[["h_2"]][["red_interact"]]
  , "Blue treatment - Age moderating" = list_fitted[["h_2"]][["blue_interact"]]
)

# Make table
modelsummary(
  list_fitted_split_2
  , coef_map = c(
    "b_trial_1_red_stimuli" = "Red yard sign"
    , "b_trial_1_blue_stimuli" = "Blue yard sign"
    , "b_age" = "Age"
    , "b_trial_1_red_stimuli:age" = "Red yard sign $\\times$ Age"
    , "b_trial_1_blue_stimuli:age" = "Blue yard sign $\\times$ Age"
    , "b_Intercept[1]" = "Intercept 1"
    , "b_Intercept[2]" = "Intercept 2"
    )
  , statistic = "conf.int"
  , gof_map = gm 
  , notes = c(
    "Data source: Pavlovia"
    , "Positive coefficients reflect the logged-odds"
    , " of claiming candidate is Republican."
    , "Numbers inside brackets reflect the 5th and 95th percentile"
    , " of the draws from the posterior distribution."
  )
  , escape = FALSE
  , output = "latex"
)
```
```{r}
#| label: tbl-robustness-alt-specification-h-2
#| tbl-cap: $H_2$ models with alternative specification
# Redefine names
list_fitted_split_3 <- list(
  "Single model" = list_fitted[["h_2"]][["single_model"]]
)

# Make table
modelsummary(
  list_fitted_split_3
  , coef_map = c(
    "b_trial_1_red_stimuli_alt" = "Red yard sign"
    , "b_trial_1_blue_stimuli_alt" = "Blue yard sign"
    , "b_Intercept[1]" = "Intercept 1"
    , "b_Intercept[2]" = "Intercept 2"
  )
  , statistic = "conf.int"
  , gof_map = gm
  , notes = c(
    "Data source: Pavlovia"
    , "Positive coefficients reflect the logged-odds"
    , " of claiming candidate is Republican."
    , "Numbers inside brakcets reflect the 5th and 95th percentile"
    , " of the draws from the posterior distribution."
  )
  , escape = FALSE
  , output = "latex"
)
```

```{r}
#| label: tbl-colorblindness
#| tbl-cap: $H_2$ moderated by colorblindness

# Redefine names
list_fitted_split_4 <- list(
  "Red - Color blindness moderating" = list_fitted[["h_2"]][["red_colorblind"]]
  , "Blue - Color blindness moderating" = list_fitted[["h_2"]][["blue_colorblind"]]
)
# Make table
modelsummary(
  list_fitted_split_4
  , coef_map = c(
    "b_trial_1_red_stimuli" = "Red yard sign"
    , "b_trial_1_blue_stimuli" = "Blue yard sign"
    , "b_trial_1_red_stimuli:color_blind_dummy" = "Red yard sign $\\times$ Color blind"
    , "b_trial_1_blue_stimuli:color_blind_dummy" = "Blue yard sign $\\times$ Color blind"
    , "b_Intercept[1]" = "Intercept 1"
    , "b_Intercept[2]" = "Intercept 2"
  )
  , statistic = "conf.int"
  , gof_map = gm 
  , notes = c(
    "Data source: Pavlovia"
    , "Positive coefficients reflect the logged-odds"
    , " of claiming candidate is Republican."
    , "Numbers inside brackets reflect the 5th and 95th percentile"
    , " of the draws from the posterior distribution."
  )
  , escape = FALSE
  , output = "latex"
)
```

### Posterior Predictive Checks
```{r}
#| label: fig-h-2-ppc
#| layout-ncol: 2
#| fig-cap: Posterior predictive checks of models in $H_2$
#| fig-subcap:
#|   - Red treatment
#|   - Blue treatment
#|   - Red treatment (Moderated by age)
#|   - Blue treatment (Moderated by age)
#|   - Single model
#|   - Red treatment (Moderated by colorblindness)
#|   - Blue treatment (Moderated by colorblindness)

list_ppc[["h_2"]][["red"]]
list_ppc[["h_2"]][["blue"]]
list_ppc[["h_2"]][["red_interact"]]
list_ppc[["h_2"]][["blue_interact"]]
list_ppc[["h_2"]][["single_model"]]
list_ppc[["h_2"]][["red_colorblind"]]
list_ppc[["h_2"]][["blue_colorblind"]]
```

{{< pagebreak >}}
\blandscape

```{r}
#| label: tbl-hypothesis-2a
#| tbl-cap: Full results for models testing bounds of $H_2$
# Redefine names
model_names <- list(
  "Trial 2 - Red treatment" = list_fitted[["h_2a"]][["trial_2_red"]]
  , "Trial 2 - Blue treatment" = list_fitted[["h_2a"]][["trial_2_blue"]]
  , "Trial 2 - Single model" = list_fitted[["h_2a"]][["trial_2_single_model"]]
)

# Change goodness of fit stuff
gm <- list(
  list("raw" = "nobs", "clean" = "Num.Obs.", "fmt" = 1)
  , list("raw" = "r.squared", "clean" = "$R^2$", "fmt" = 2)
  , list("raw" = "waic", "clean" = "WAIC", "fmt" = 2)
)
# Make table
modelsummary(
  model_names
  , coef_map = c(
    "b_trial_2_red_stimuli" = "Red yard sign"
    , "b_trial_2_blue_stimuli" = "Blue yard sign"
    , "b_trial_2_red_stimuli_alt" = "Red yard sign"
    , "b_trial_2_blue_stimuli_alt" = "Blue yard sign"
    , "b_Intercept[1]" = "Intercept 1"
    , "b_Intercept[2]" = "Intercept 2"
    )
  , statistic = "conf.int"
  , gof_map = gm 
  , notes = c(
    "Data source: Pavlovia"
    , "Positive coefficients reflect the logged-odds"
    , " of claiming candidate is Republican."
    , "Numbers inside brackets reflect the 5th and 95th percentile"
    , " of the draws from the posterior distribution."
  )
  , escape = FALSE
  , output = "latex"
)
```

```{r}
#| label: tbl-hypothesis-2a-cont
#| tbl-cap: Full results for models testing bounds of $H_2$ (cont.)
# Redefine names
model_names <- list(
  "Trial 3 - Red treatment" = list_fitted[["h_2a"]][["trial_3_red"]]
  , "Trial 3 - Blue treatment" = list_fitted[["h_2a"]][["trial_3_blue"]]
  , "Trial 3 - Single model" = list_fitted[["h_2a"]][["trial_3_single_model"]]
)

# Change goodness of fit stuff
gm <- list(
  list("raw" = "nobs", "clean" = "Num.Obs.", "fmt" = 1)
  , list("raw" = "r.squared", "clean" = "$R^2$", "fmt" = 2)
  , list("raw" = "waic", "clean" = "WAIC", "fmt" = 2)
)
# Make table
modelsummary(
  model_names
  , coef_map = c(
    "b_trial_3_red_stimuli" = "Red yard sign"
    , "b_trial_3_blue_stimuli" = "Blue yard sign"
    , "b_trial_3_red_stimuli_alt" = "Red yard sign"
    , "b_trial_3_blue_stimuli_alt" = "Blue yard sign"
    , "b_Intercept[1]" = "Intercept 1"
    , "b_Intercept[2]" = "Intercept 2"
    )
  , statistic = "conf.int"
  , gof_map = gm 
  , notes = c(
    "Data source: Pavlovia"
    , "Positive coefficients reflect the logged-odds"
    , " of claiming candidate is Republican."
    , "Numbers inside brackets reflect the 5th and 95th percentile"
    , " of the draws from the posterior distribution."
  )
  , escape = FALSE
  , output = "latex"
)
```

\elandscape
{{< pagebreak >}}

### Posterior Predictive Checks
```{r}
#| label: fig-h-2a-ppc
#| layout-ncol: 2
#| fig-cap: Posterior predictive checks of models in $H_2$
#| fig-subcap:
#|   - Red treatment (Trial 2)
#|   - Blue treatment (Trial 2)
#|   - Combined model (Trial 2)
#|   - Red treatment (Trial 3)
#|   - Blue treatment (Trial 3)
#|   - Combined model (Trial 3)

list_ppc[["h_2a"]][["trial_2_red"]]
list_ppc[["h_2a"]][["trial_2_blue"]]
list_ppc[["h_2a"]][["trial_2_single_model"]]
list_ppc[["h_2a"]][["trial_3_red"]]
list_ppc[["h_2a"]][["trial_3_blue"]]
list_ppc[["h_2a"]][["trial_3_single_model"]]
```

{{< pagebreak >}}

## Hypothesis 3

### Models

```{r}
#| label: tbl-h-3-main
#| tbl-cap: Support of candidate based on partisanship

# Redefine names
model_names <- list(
  "Red stimuli" = list_fitted[["h_3"]][["red"]]
  , "Blue stimuli" = list_fitted[["h_3"]][["blue"]]
)

# Make table
modelsummary(
  model_names
  , coef_map = c(
    "b_trial_1_red_stimuli" = "Red yard sign"
    , "b_trial_1_blue_stimuli" = "Blue yard sign"
    , "b_pid_7" = "Party ID"
    , "b_trial_1_red_stimuli:pid_7" = "Red yard sign $\\times$ Party ID"
    , "b_trial_1_blue_stimuli:pid_7" = "Blue yard sign $\\times$ Party ID"
    , "b_age" = "Age"
    , "b_gender_idMale" = "Gender - Male"
    , "b_gender_idNonMbinary" = "Gender - Non-binary"
    , "b_gender_idPrefernottochoose" = "Gender - Prefer not to choose"
    , "b_race_idBlacknonMhispanic" = "Black"
    , "b_race_idHispanice" = "Hispanic"
    , "b_race_idMultipleraces" = "Multiple races"
    , "b_race_idNativeAmericanDAlaskaNativeOrOtherRace" = "Other race"
    , "b_race_idWhitenonMhispanic" = "White"
  )  
  , statistic = "conf.int"
  , gof_map = gm 
  , notes = c(
    "Data source: Pavlovia"
    , "Positive coefficients reflect the logged-odds"
    , " of claiming candidate is Republican."
    , "Numbers inside brackets reflect the 5th and 95th percentile"
    , " of the draws from the posterior distribution."
  )
  , escape = FALSE
  , output = "latex"
)
```

{{< pagebreak >}}

### Posterior predictive Checks

{{< pagebreak >}}

## Hypothesis 4

### Models

{{< pagebreak >}}

### Posterior Predictive Checks