### 2023-06-20 assets/_theory.qmd
adjust title of theory fig and then add some slides pre-FRAPSA

---
format: html
---
```{mermaid}
%%| label: fig-theory
%%| fig-cap: Snap-judgment model of politically-relevant visual information
%%| fig-width: 6.5
%%{init: {'theme':'base', 'themeVariables':{'primaryColor':'#ffffff', 'primaryBorderColor': '#000000'}}}%%
graph TD
    A[Detection of politically-relevant visual information] --> B(Activation of retrieval mechanism)
    B --> C[Memory and contiguous paths hot]
    C --> D[Appraisal of hot autonomous nervous system]
    D --> E1[Negative Valence]
    D --> E2[Positive Valence]
    E1 --> F1[Disengagement motivation]
    E2 --> F2[Attention activated]
    F1 --> |Neg. new information| G1[Encoded as negative]
    F1 --> |Pos. new information| G2[Encoded as positive]
    F2 --> |Neg. new information| G3[Encoded as negative]
    F2 --> |Pos. new information| G4[Encoded as positive]
    G1 --> H1[Strengthens negative path]
    G2 --> H2[Weakens negative path]
    G3 --> H3[Weakens positive path]
    G4 --> H4[Strengthens positive path]
    G2 --> |forgotten| H1
    H2 --> |reinforced| G2
    H3 --> |reinforced| G3
    G3 --> |forgotten| H4
```