### 2023-06-20 chapter_3/src/slides/slides.qmd
add slides for prr of chapter_3 pre-FRAPSA/ISPP

---
title: "How does color influence our political conversations?"
author:
  - name: Damon C. Roberts
    title: PhD Candidate
    affiliations:
      - id: CU
        name: University of Colorado Boulder
        department: Political Science
        address: 333 UCB
        city: Boulder
        region: CO 
        postal-code: 80309-0333
    attributes:
      corresponding: true
format:
  revealjs: 
    embed-resources: true
    theme: [dark, ../../assets/slides/custom.scss]
    transition: slide
    scrollable: true
    incremental: true
    slide-number: true
    toc: true
    toc-depth: 1
    slide-level: 3
    bibliography: ../../../assets/references.bib
    footer: |
        <span class="faux-block">{{< fa brands creative-commons >}} 2021-2023 Damon Charles Roberts</span>
        <span class="faux-block"><br>All content licensed under {{< fa brands creative-commons >}} {{< fa brands creative-commons-by >}} {{< fa brands creative-commons-sa >}} [Creative Commons CC BY-SA 4.0](https://creativecommons.org/licenses/by-sa/4.0/)</span>
    self-contained: true
    output-file: color_and_conversation.html
execute:
  echo: false
---

```{r}
#| label: setup-block

# Load some nifty functions
box::use(
  ../../../chapter_2/src/R/predicted_prob_bar[...]
)
# Load fitted models
load(
  file = "../../../chapter_2/data/temp/models/study_1_fitted.RData"
)
```

# Background

---

- Chapter from a larger book project
- Book examines
  - does color convey political information?
    - particularly partisan affiliation

## Key findings from other chapters

---

```{r}
#| label: fig-associations
#| layout-ncol: 2
#| fig-cap: People associate red yard signs with Republicans and blue yard signs with Democrats
#| fig-subcap:
#|   - Red treatment
#|   - Blue treatment
red <- predicted_prob_bar(
  fitted_model = list_fitted[["h_2"]][["red"]]
  , x_axis = "trial_1_red_stimuli"
)
blue <- predicted_prob_bar(
  fitted_model = list_fitted[["h_2"]][["blue"]]
  , x_axis = "trial_1_blue_stimuli"
  , treatment = "Blue"
)
red
blue
```

---

```{r}
#| label: fig-associations-weaker
#| layout-ncol: 2
#| fig-cap: Mixing red and blue in a yard sign increases uncertainty about partisanship
#| fig-subcap:
#|   - Trial 2 Red treatment
#|   - Trial 2 Blue treatment
#|   - Trial 3 Red treatment
#|   - Trial 3 Blue treatment
plot_trial_2_red <- predicted_prob_bar(
  fitted_model = list_fitted[["h_2a"]][["trial_2_red"]]
  , x_axis = "trial_2_red_stimuli"
)
plot_trial_2_blue <- predicted_prob_bar(
  fitted_model = list_fitted[["h_2a"]][["trial_2_blue"]]
  , x_axis = "trial_2_blue_stimuli"
  , treatment = "Blue"
)
plot_trial_3_red <- predicted_prob_bar(
  fitted_model = list_fitted[["h_2a"]][["trial_3_red"]]
  , x_axis = "trial_3_red_stimuli"
)
plot_trial_3_blue <- predicted_prob_bar(
  fitted_model = list_fitted[["h_2a"]][["trial_3_blue"]]
  , x_axis = "trial_3_blue_stimuli"
  , treatment = "Blue"
)
plot_trial_2_red
plot_trial_2_blue
plot_trial_3_red
plot_trial_3_blue
```

---

```{r}
#| label: fig-vote
#| layout-ncol: 2
#| fig-cap: Republicans report a higher probability of voting for candidates with red yard signs; Democrats report a higher probability of voting for candidates with blue yard signs
#| fig-subcap:
#|   - Red treatment
#|   - Blue treatment
red <- predicted_prob_bar(
  fitted_model = list_fitted[["h_3"]][["red"]]
  , x_axis = "trial_1_red_stimuli"
  , treatment = "Red"
  , hypothesis = "H3"
  , x_label = "Party ID"
  , y_label = "Pr(Vote for candidate)"
  , legend_title = "Treatment"
)
blue <- predicted_prob_bar(
  fitted_model = list_fitted[["h_3"]][["blue"]]
  , x_axis = "trial_1_blue_stimuli"
  , treatment = "Blue"
  , hypothesis = "H3"
  , x_label = "Party ID"
  , y_label = "Pr(Vote for candidate)"
  , legend_title = "Treatment"
)
red
blue
```

# Question guiding this chapter

- Does color influence whether we talk about politics?

# Building a theory of snap-judgment based on politically-relevant visual information

## A common language

- Valance: a positive or negative response to an object.
- Affect (or affective state): feeling positive or negative in response to an object.
  - Pre-conscious response
- Emotion: Much more complex labels assigned to affective responses.
  - Often requires some amount of conscious processing to assign an appropriate label.
  - Slower
- Attitudes are an expression of valanced responses to some object.

## Learning from our psychology and neuroscience friends 😊

- There are a lot of useful things that we can borrow from social and affective neuroscientists and social psychologists

---

1. Memory reflects stored prior experiences with various objects
  - These memories are not just factual recollections of things
    - e.g., the balloon was red.
  - Memories often encode your affective responses to that object.
2. Relatedly, memories are multi-dimensional
3. How we form an attitude with new information (or a combination of it) is by accessing memories that share features in common with the object.
4. Memories are a complex web of prior experiences that have a whole lot of strands connecting them.
5. Social groups and social hiearchies are a dimension that we pay attention to and encode along with a memory.

## A snap-judgment model of politically-relevant visual information

---

{{< include ../../../assets/_theory.qmd >}}

# References

:::{#refs}
:::