### 2023-06-21 chapter_3/src/slides/slides.qmd
some more iterations of slides for ch3

---
title: "How does color influence our political conversations?"
author:
  - name: Damon C. Roberts [{{< fa brands orcid size=xs >}}](https://orcid.org/0000-0002-4360-3675)
format:
  revealjs: 
    embed-resources: true
    theme: [dark, ../../assets/slides/custom.scss]
    transition: slide
    scrollable: false
    incremental: true
    slide-number: true
    toc: true
    toc-depth: 1
    slide-level: 3
    bibliography: ../../../assets/references.bib
    footer: |
        <span class="faux-block">{{< fa brands creative-commons >}} 2021-2023 Damon Charles Roberts</span>
        <span class="faux-block"><br>All content licensed under {{< fa brands creative-commons >}} {{< fa brands creative-commons-by >}} {{< fa brands creative-commons-sa >}} [Creative Commons CC BY-SA 4.0](https://creativecommons.org/licenses/by-sa/4.0/)</span>
    self-contained: true
    output-file: color_and_conversation.html
execute:
  echo: false
---

```{r}
#| label: setup-block

# Load some nifty functions
box::use(
  ../../../chapter_2/src/R/predicted_prob_bar[...]
)
# Load fitted models
load(
  file = "../../../chapter_2/data/temp/models/study_1_fitted.RData"
)
```

# Background

::: footer
:::
---

::: {.notes}
- This is a chapter from a larger book project.
- The book addresses the question: 
  - does color convey political information? 
  - does it have pre-conscious effects on information that comes later?
:::

::: footer
:::

## Key findings from other chapters

::: {.notes}
- Ran a study on prolific
- Counterbalanced questions of attention, age, gender, race, colorblindness, partisan identity, etc.
- Ran three trials that used mousetracking software
- See one of three yard signs in each trial
  - Red, blue, or white yard sign
:::

::: footer
:::
---

```{r}
#| label: fig-associations
#| layout-ncol: 2
#| fig-cap: People associate red yard signs with Republicans and blue yard signs with Democrats
#| fig-subcap:
#|   - Red treatment
#|   - Blue treatment
red <- predicted_prob_bar(
  fitted_model = list_fitted[["h_2"]][["red"]]
  , x_axis = "trial_1_red_stimuli"
)
blue <- predicted_prob_bar(
  fitted_model = list_fitted[["h_2"]][["blue"]]
  , x_axis = "trial_1_blue_stimuli"
  , treatment = "Blue"
)
red
blue
```

::: {.notes}
- Average predicted probabilities from two ordered logistic regressions
- Outcome is the respondent's belief that the candidate was either a Republican, Independent, or a Democrat.
- Read each figure by comparing the same colored bars. 
- Candidates using red yard signs are more likely to be seen as Republican
- Candidates using blue yard signs are more likely to be seen as Democrat
:::

::: footer
:::
---

```{r}
#| label: fig-associations-weaker
#| layout-ncol: 2
#| fig-cap: Mixing red and blue in a yard sign increases uncertainty about partisanship
#| fig-subcap:
#|   - Trial 2 Red treatment
#|   - Trial 2 Blue treatment
#|   - Trial 3 Red treatment
#|   - Trial 3 Blue treatment
plot_trial_2_red <- predicted_prob_bar(
  fitted_model = list_fitted[["h_2a"]][["trial_2_red"]]
  , x_axis = "trial_2_red_stimuli"
)
plot_trial_2_blue <- predicted_prob_bar(
  fitted_model = list_fitted[["h_2a"]][["trial_2_blue"]]
  , x_axis = "trial_2_blue_stimuli"
  , treatment = "Blue"
)
plot_trial_3_red <- predicted_prob_bar(
  fitted_model = list_fitted[["h_2a"]][["trial_3_red"]]
  , x_axis = "trial_3_red_stimuli"
)
plot_trial_3_blue <- predicted_prob_bar(
  fitted_model = list_fitted[["h_2a"]][["trial_3_blue"]]
  , x_axis = "trial_3_blue_stimuli"
  , treatment = "Blue"
)
plot_trial_2_red
plot_trial_2_blue
plot_trial_3_red
plot_trial_3_blue
```

::: {.notes}
- The other two trials used less of the primary color and brought in more red or blue
- Do this as a test of the strength of these effects
- See that the results (kind of) weaken.
  - See that these effects weaken for the blue treatment. Much less change for the red treatment. Likely American flag.
:::

::: footer
:::
---

```{r}
#| label: fig-vote
#| layout-ncol: 2
#| fig-cap: Republicans report a higher probability of voting for candidates with red yard signs; Democrats report a higher probability of voting for candidates with blue yard signs
#| fig-subcap:
#|   - Red treatment
#|   - Blue treatment
red <- predicted_prob_bar(
  fitted_model = list_fitted[["h_3"]][["red"]]
  , x_axis = "trial_1_red_stimuli"
  , treatment = "Red"
  , hypothesis = "H3"
  , x_label = "Party ID"
  , y_label = "Pr(Vote for candidate)"
  , legend_title = "Treatment"
)
blue <- predicted_prob_bar(
  fitted_model = list_fitted[["h_3"]][["blue"]]
  , x_axis = "trial_1_blue_stimuli"
  , treatment = "Blue"
  , hypothesis = "H3"
  , x_label = "Party ID"
  , y_label = "Pr(Vote for candidate)"
  , legend_title = "Treatment"
)
red
blue
```

::: {.notes}
- After viewing the three trials, participants are asked which of the three yard signs they prefer.
- Make a dichotomous variable of whether or not they like the first trial's yard sign.
- run logistic regressions to see whether votes are explained by the stimuli moderated by partisan identification
- See that there is a low predicted probability of strong Democrats voting for a candidate with a red yard sign.
  - See that the predicted probability of strong Republicans voting for a candidate with a red yard sign is high.
- See that there is a low predicted probability of strong republicans voting for a candidate with a blue yard sign.
  - See that the predicted probability of a strong Democrat voting for a candidate with a blue yard sign is high.
:::

::: footer
:::

# Question guiding this chapter

- Does color influence whether we talk about politics?

::: {.notes}
- Color seems to hold information about partisan affilations
  - Red = Republican
  - Blue = Democrat
- Unclear what sorts of effects this will have on attitudes and behaviors, though.
:::

::: footer
:::

# Building a theory of snap-judgment based on politically-relevant visual information

::: footer
:::

## A common language

- Valance: a positive or negative response to an object.
- Affect (or affective state): feeling positive or negative in response to an object.
  - Pre-conscious response
- Emotion: Much more complex labels assigned to affective responses.
  - Often requires some amount of conscious processing to assign an appropriate label.
  - Slower
- Attitudes are an expression of valanced responses to some object [@fazio_2007_sc].

::: footer
:::

## A cognitive architecture

::: footer
:::

---

1. Memories are stored prior experiences [@cunningham_2021_yup].
  - often encode affective responses
2. Memories are multi-dimensional [@kahana_et-al_2022_ohhum].
3. Sensory and affective information is often encoded in memories [@kensinger_fields_2022_ohhum].
4. Memories are a complex network [@kahana_et-al_2022_ohhum].
  - Interconnected by different bits of information
5. Social groups and hierarchies are reflected in this network  [@santavirta_et-al_2023_n].

::: footer
:::

## A snap-judgment model of politically-relevant visual information

::: footer
:::

---

{{< include ../../../assets/_theory.qmd >}}

::: {.notes}
Scenario.
- At the airport and watching the news
- Someone sits next to you before boarding
- They are wearing a red hat
- You might have one of two reactions
- A small conversation starts up since they see you watching the news.
- You see its a MAGA hat but they tell you they are having a hard time squaring some of their other beliefs with the stances of the party on immigration.
- You may shift your initial impression a bit. Feel more positive or more negative. However, you haven't fully reversed course on that. You have learned something though.

Implications.
- Color has downstream effects on our attitudes towards an object related to politics.
- The degree to which color matters in politics is whether or not it is primed.
:::

::: footer
:::

# Proposed research design

## Study 1

+---------------------:+:-----------------:+
| 🚫 Prime, Red shirt  | Prime, Red shirt  |
+----------------------+-------------------+
| 🚫 Prime, Blue shirt | Prime, Blue shirt |
+----------------------+-------------------+

: $2 \times 2$ factorial design {#tbl-study-1}

::: footer
:::

## Study 2

+-------------------------:+:---------------------:+
| Red shirt, 🚫 congruent  | Red shirt, congruent  |
+--------------------------+-----------------------+
| Blue shirt, 🚫 congruent | Blue shirt, congruent |
+--------------------------+-----------------------+

: $2 \times 2$ factorial design {#tbl-study-2}

::: footer
:::

# References

::: {#refs}
:::

::: footer
:::