### 2023-06-24 chapter_2/src/manuscript.qmd

# Does color convey political information?
<!--
    - Description: QMD script for empirical chapter 1
    - Updated: 2023-07-16
    - Notes:
        * Code:
        * Writing:
-->

## Introduction

::: {#fig-party-brand layout-ncol=2}

![Republicans pre-2000](assets/img/rnc_logo_pre-2000.png){#fig-republican}

![Democrats pre-2000](assets/img/dnc_logo_1960.png){#fig-democrat}

![Republicans 2023](assets/img/rnc_logo_2023.png){#fig-republican-post}

![Democrats 2023](assets/img/dnc_logo_2023.png){#fig-democrat-post}

Party logos
:::

Now that I have presented some background and have introduced the snap-judgment model of politically-relevant visual information, this chapter is a first attempt to tackle the question at hand. Do colors provide political information that shape attitudes? There are two goals that I am hoping to achieve in this chapter. The first is to convince the reader that the colors red and blue convey affiliative information. Second, that these connections between colors and political groups can shape political outcomes.

This chapter applies the snap-judgment model to a simple case. The use of color on yard signs used by campaigns in elections. I focus on yard signs as they are a common form of campaign marketing and evidence suggests that they can shape individual behavior [@makse_et-al_2019_oup] and electoral outcomes [@green_et-al_2016_es]. Additionally, I start by examining the colors red and blue. While there are other colors that may provide political information -- black and white for Black Lives Matter and Blue Lives Matter movements, the overarching goal of this book is to lay the foundation for a larger research agenda in visual political psychology [see @bucy_joo_2021_ijpp]. Without such foundations, we are studying complex cases and circumstances upon assumptions without a grasp of how, and if they even do, occur in the cases that might seem trivial to some. While there indeed is empirical and experiential evidence suggesting that the Republicans use the color red and Democrats use the color blue [@schloss_palmer_2014_pbr;@williams_et-al_2022_jomp], there is yet to be a theoretical framework that explains how these connections happen at the individual level. It may also be unsatisfying that I do not address *how* these connections came about. While the connections between Republicans with the color red and Democrats with the color blue seemed to quicken after the 2000 Presidential election where news networks covering the election began to consistently use those colors [@elving_2014_npr], it is not clear if this was a deliberate choice or whether this was yet another symptom of an increasingly polarized public who just want easily digestable information about politics so they can move on to other matters in their lives. An undertaking to address such a question is for a different book, perhaps.

To give a preview for this chapter, I propose two studies to explore the claims I make in this chapter. Yard signs are a simple yet effective source of campaign branding for political candidates as they convey information that is often useful for voters -- partisanship [@campbell_et-al_1960_jws]. As yard signs are a simple and static form of campaign advertising, they cannot rely heavily on more complicated forms of information. However, one form of information they rely on and vary in is their color. With these characteristics, yard signs are an ideal candidate both from a statistical standpoint and also substantive for an experimental test of my claims in this chapter. From a statistical standpoint, as they are static forms of party branding they have fewer moving parts, which help isolate the effect that color has on shaping attitudes. From a substantive standpoint, while we understand that yard signs matter to campaigns, we have yet to systematically explore more simple design choices such as color.

In [Study 1](#sec-study-1), I conduct an online experiment that examines whether the public notice color and whether their attitudes are shaped by "Republican Red" and "Democratic Blue" on the yard signs of a fictional candidate. Based on informal conversations with campaign marketing consultants, it appears that campaigns are aware that the colors red and blue betray the partisanship of a candidate to voter. So in [Study 2](#sec-study-2), I examine whether campaigns react to the partisan composition of an electorate as a response to these individual-level patterns. I do this by examining how electoral context shapes variation in using these partisan colors. While these two studies highlight endogenous processes, I am not necessarily arguing that either of these patterns only flow one way. In fact, I believe that there is a necessary feedback loop between the parties and their base of support. Voters support candidates who signal strong allegiance to the party to increase their chances for electoral success [@utych_2020_es] and the parties are motivated to distinguish themselves [@lee_2009;@clifford_2020_pb] as a result of their high levels of ideological and affective polarization [@dietrich_2021_pa;@enders_2021_jop]. Again, to tease out the direction of blame is more appropriate for another time.

In the first study, I find that individuals do shift their perceptions of political candidates in response to the use of the colors red and blue, they consistently and strongly associate Republicans with the color red and Democrats with the color blue, and their willingness to vote for a candidate can be dramatically different depending on whether the color used aligns with the individual's partisan identification. The second study suggests that candidates in districts where Democrats have a higher vote share use more blue on their yard signs while also using less red. The implications of this are that the colors red and blue convey political information. More generally, I suspect other colors and more complex forms of visual information such as symbols are capable of conveying political information as well. Perhaps most interesting among these results is that the color red and blue alone can make individuals willing to vote for a candidate. The association between the colors red and blue are strong enough that individuals express a willingness to cast a vote for a candidate based on the color of their yard sign without confirming any of their policy stances or their partisan affiliation. These tendencies, even if based on a hunch for the folks on the ground, seem to lead to descriptive differences in the use of these colors as it guides strategy. The following chapters dig a bit deeper into the idea that colors can shape attitudes by pitting the information individuals get from color against information gleaned from explicit policy stances.

## The role of visual information in politics

A central theme of an edited issue of *The International Journal of Press/Politics* is that visual politics is important and understudied [@lilleker_2019_pm]. Those who are engaged in these questions attribute these challenges to methodological sophistication and the difficult task of interdisciplinary theorizing [@gerodimos_2019_pm; @bucy_joo_2021_ijpp].

The rise of television consumption changed the focus of the medium in the research for political communication scholars [@hall-jamieson_2014_ohpc]. In a similar way, so does the rise of image-based social media and has set new agendas. For several methodological and disciplinary reasons, the visual aspects of television were of little focus in the literature [@bucy_joo_2021_ijpp]. However, the news organizations and politicians are responding to the ubiquitous use of the public to use image-dominant social media platforms like TikTok and Instagram, by joining such platforms. They are relatively active on these platforms as well. Scholars need to make this transition to integrate the role of simple visual information into our theories of political information processing (abbreviated as *pip*) and attitude formation.

How does politically-relevant visual information matter to politics? From an evolutionary-biological perspective, visual information has been a common source of information for millions of years that a variety of single-and-multi-cell organisms rely upon to evaluate their environment [see @grabe_bucy_2009_oup, Chapter 1 for a useful discussion]. Visual information processing, as an ancient biological invention, the human brain is organized around the processing of it. Reflecting on this, many scholars of neuroscience argue that visual information is the fastest form of information processing for humans. For example, even complex visual information, such as the warmth expressed in someone's facial features, is automatically and pre-consciously processed in only about 33 milliseconds (abbreviated as ms) [@ames_et-al_2012_ohsn].

Approached from a different perspective, as humans are cognitive misers \index{cognitive misers}, visual information in the realm of politics provides efficient information to voters about politically-relevant actors and events [@lilleker_2019_pm]. Evidence suggests that even simple party branding on yard signs has an emotional appeal and encourages a shift in attitudes toward the person who owns the yard sign [@makse_et-al_2019_oup]. That is, even branding as simple as yard signs appears to influence the attitudes of those who view them. However, like all types of party branding, there remain questions about the individual visual components' contribution.

More simple forms of visual information are likely even more efficient forms of information. In the context of the United States, the "Republican red" and "Democratic blue" is a relatively recent invention that likely has significant import in an era of significant effort by the parties to distinguish themselves from each other [@clifford_2020_pb] and where voters toe the party line [@utych_2020_es]. Since the 2000 presidential election, the media have consistently used red on their electoral maps in "horserace" journalism to represent Republicans and blue to represent Democrats [@elving_2014_npr]. The supposed consequence is that Democrats now report a preference for the color blue over the color red, and Republicans report a preference for the color red over the color blue [@schloss_palmer_2014_pbr]. Others have demonstrated the strategic choices that candidates make on branding choices -- including color -- based on the types of information they want to convey to voters to distinguish them from their rivals [see @williams_et-al_2022_jomp].

In western Europe, voters are better at connecting the ideological positioning of a party with the color they use in their branding [@casiraghi_et-al_2022_pp]. The ability to do so is strengthened when the party is longer-surviving and is more prominent [@casiraghi_et-al_2022_pp]. The use of politically-relevant colors activates biases toward pre-existing ideological and partisan preferences among voters in a Spanish sample [@maestre_medero_2022_pr]. And male and female candidates in the United States use different colors -- among other things such as fonts -- in their campaign branding to convey distinguishing information about themselves [@williams_et-al_2022_jomp]. What remains unclear is whether these patterns in the United States are just simply patterns or whether there are identifiable causes to these patterns. While political psychologists are increasingly interdisciplinary and rely on psychology and neuroscience, it is relatively unclear how color can convey information as complex as politics. Further, it is unclear whether such a simple type of information has enough efficacy to shape political outcomes such as voting intention. This chapter is a first cut at addressing these open questions.

## Integrating color into a model of political information processing

Existing models of political information processing (abbreviated here as *pip*) primarily focus on complex forms of political information such as policy positions communicated verbally or through text. As individuals process visual information before other sorts of information, we might expect that they may form a snap judgment or the initial appraisal of an object.

Color and other simple visual information are processed much quicker and occur more frequently than text-based information that may be communicated through a news article [@mehta_zhu_2009_s]. As the color and other visual information are processed differently, we should also consider its use as political information. As visual information is affectively encoded [@cimbalo_et-al_1978_jgp], it can affect the affective state and processing of more complex information, such as text. The visual information provides a snap judgment or an impression of the object through faster processing and activates particular neurological processes that influence subsequent information appraisals [@ames_et-al_2012_ohsn]. This implies that snap-judgments from visual forms of political information may have downstream effects that shape the way in which we engage with forms of political information that we traditionally consider.

Before expanding upon the role of colors in shaping political attitudes, let me first define an attitude as a concept. An attitude represents an accessible, valanced evaluation of associated prior information and experiences. This conceptualization fits with that of the Object-Evaluation Associations Model [@fazio_2007_sc]. As opposed to viewing attitudes as a latent collection of memories, it views attitudes as measurable evaluations of memories. As memories are at the core of an attitude, the association of memories with its evaluative component [see @kensinger_fields_2022_ohhum] contributes to the perspective that attitudes are affective. This implies that we should be able to measure attitudes but that such an operationalization requires careful consideration of the context's role in any given measure of an attitude as they result from memories [@fazio_2007_sc]. In this chapter, I consider colors' effect on shaping attitudes about a political object -- a yard sign. As I hold this constant in the chapter, concerns about variation in what people are thinking about when seeing the color red and blue is less of a concern in this chapter. Later chapters grapple with this consideration.

In line with existing theories of *pip*, I conceptualize attitudes as associative. This means that attitudes may be unstable - not randomly, in any case. As attitudes are associative, they manifest slightly differently depending on the associative paths activated [@fazio_2007_sc]. The retrieval of relevant memories to the attitude depends on many factors, such as the recency of the event, the similarity of the context, and the importance or salience of the memory [@kahana_et-al_2022_ohhum]. This means that the memories retrieved to contribute to an attitude are quite variable. However, to understand where that variability comes from, we must understand the deeper processes influencing how information is encoded and later retrieved. This illustrates my need to build upon the reigning models of *pip*. Colors may act as a contextual feature that may lead to this variability in how a given set of political information may shape attitudes.

Colors are associative and are affectively encoded [@cimbalo_et-al_1978_jgp]. When individuals access a memory, they do not just recall an object, but they may recall visual information such as the color of an object [@mehta_zhu_2009_s]. Due to these features, colors can be processed pre-consciously and very quickly [@mehta_zhu_2009_s]. As they are affectively encoded, their associations with particular memory contribute to the evaluative component of the memory. For example, individuals associate colors like red with anger and arousal [@valdez_mehrabian_1994], whereas individuals associate blue with things like happiness and pleasure [@dandrade_egan_1974_ae]. This means that colors are compelling as contextual information shaping the subsequent processing and integration of "traditional" forms of political information to construct an attitude. As the preference for a particular color correlates with political attitudes [@schloss_palmer_2014_pbr], they may have some causal influence upon political attitudes; rather than just a correlation with them.

According to the literature in affective neuroscience, visual information is processed in various parts of the brain, such as the visual cortex [@goldstein_brockmole_2017_cl]. These types of processes do not occur in particular areas of the brain as some like to think. As attitudes are associative, components of our brain works as this complex highway connecting different types of information. As a result, the processing of visual information will activate other areas of the brain and will make associated paths "hot." One such area is the amygdala. Neuroscientists believe that as visual information is quickly and subconsciously processed, the amygdala takes and appraises it based on the paths it activated; this generates a simple affective response to such information [@winkielman_et-al_2011_ohsn]. More complex, categorical emotion occurs later in conscious processing [@winkielman_et-al_2011_ohsn].

To provide a less technical description of the processes occuring here, once politically-relevant visual information is detected, the retina passes it to the brain. Once there, the brain attempts to classify the visual information by activating networks of neurons associated with the current information. With these activated pathways, the brain also attempts to appraise such information based on quick classification. For attitude formation, this implies that, as memories are affectively encoded [@kensinger_fields_2022_ohhum], these memories help areas such as the amygdala to appraise the current information.

These fast affective classifications are valanced rather than the more laborious categorization of specific emotions (e.g., anger, anxiety, fear, happy). What this means is that rather than conceptualizing the affective component of this process in line with the popular conceptualizations in political science, such as affective intelligence theory [@marcus_2000_arps], I conceptualize affect as a more extensive system that considers emotion only as a conscious classification process that occurs later than initial valanced appraisals of an object [@sander_2013_chhan; @ralph_anderson_2018_pup]. That is, emotion -- the complex classification of appraisals -- is a conscious component of affect. The pre-conscious process of affect occurs first with the simple and automatic valanced classification of an object [@winkielman_et-al_2011_ohsn; @dror_2017_er].

The pre-conscious appraisals of the visual information one encounters encourages particular behavioral and attitudinal motivations [@valentino_et-al_2011_jop; @ralph_anderson_2018_pup]. This has evolutionary roots for survival [@ralph_anderson_2018_pup; @parker_2003_p]. While affective appraisals can lead to complex motivations, such as anxiety leading to motivations for information seeking [@marcus_2000_arps], affective appraisals are valanced and are more automatic [@winkielman_et-al_2011_ohsn]. These affective appraisals lead to a desire to either retract or engage more with the object [@valentino_et-al_2011_jop]. The snap judgment resulting from the automatic processing of politically relevant visual information should likely lead to an affective response that motivates either a desire to engage more with the object or disengage.

While the visual information may encourage a particular immediate reaction to engage or disengage from the information, subsequent information processing and more conscious processing adjusts this initial appraisal generated by the snap judgment [@kensinger_fields_2022_ohhum]. While subsequent information may amend one's snap-judgment, the snap-judgment nevertheless influences the processing of subsequent information by activating particular paths, which is later encoded as associated with the object as it converts to a memory [@lodge_taber_2013_cup; @kensinger_fields_2022_ohhum].

Beyond a motivated backlash effect, the popular application of motivated reasoning to political science finds evidence suggesting that individuals take more time processing information that runs against their pre-existing preferences [@taber_lodge_2006_ajps]. The explanation for this is that people are finding ways to come up with counterarguments to such information. Therefore, it is possible that we may expect that those interacting with visual information of a political party for which they disagree with spend more time looking at those objects.

@fig-snap-judgement-model presents an illustration of the snap-judgment model.

```{mermaid}
%%| label: fig-snap-judgement-model
%%| fig-cap: Snap judgement model
%%| fig-width: 6
%%{init: {'theme':'base', 'themeVariables':{'primaryColor':'#ffffff', 'primaryBorderColor': '#000000'}}}%%
graph TD;
    A[Detection of politically-relevant visual information] --> B(Activation of retrieval mechanism);
    B --> C[Memory and contiguous paths hot];
    C --> D[Appraisal of hot autonomous nervous system];
    D --> E1[Negative Valence];
    D --> E2[Positive Valence];
    E1 --> F1[Disengagement motivation];
    E2 --> F2[Attention activated];
    F1 --> |Neg. new information| G1[Encoded as negative];
    F1 --> |Pos. new information| G2[Encoded as positive];
    F2 --> |Neg. new information| G3[Encoded as negative];
    F2 --> |Pos. new information| G4[Encoded as positive];
    G1 --> H1[Strengthens negative path];
    G2 --> H2[Weakens negative path];
    G3 --> H3[Weakens positive path];
    G4 --> H4[Strengthens positive path];
    G2 --> |forgotten| H1;
    H2 --> |reinforced| G2;
    H3 --> |reinforced| G3;
    G3 --> |forgotten| H4;
```

Let me illustrate the snap-judgment model with a common experience for residents of the United States. Say you are driving down a highway. At 65 miles an hour, you are traveling at about 95 feet per second at this speed. You split your attention. You focus your eyes on the conditions of the road in front of you, the cars in front of you, and the rear-view mirror where your kids are either dropping food in the crevice between the seats or trying to grab your attention. Out of the corner of your eye, you see a sign. It is not a road sign because it is not on the familiar white or yellow background with black lettering. It is election season. You correctly infer that it is a political yard sign. In this split second, you notice the sign's color and may see a name: Mitch McConnell. You now are racking your brain to think about who that is. If you are politically engaged, you might come to that recognition of the name quickly or it may take you significantly longer if you are less politically engaged [see @kahana_et-al_2022_ohhum]. You figure out that they are a Republican politician. You may have come to this with the help of the fact that every year you have seen yard signs on this stretch of highway. You know that when you see those electoral maps pop up on your news app on your phone that the electoral forecasts always represent Republican support with red and blue for Democrats. Once you have figured out who this person is, with the help of this other information, you react: "ugh, that guy is too loyal to Trump" or "yeah! He's loyal to Trump". You've expressed a political attitude.

What the snap-judgment model predicts is happening in your head is that as soon as the light that bounces off the sign to produce a particular wavelength hits your eyes, your brain is already trying to make sense of this information. This is a valuable tool for survival that biology has optimized for millions of years [@parker_2003_p]. Rather than slowly processing the visual information and finding yourself in the jaws of a predator or processing it quickly but forming the wrong impression and running away from a friend, the brain processes the information quickly and subconsciously [@newell_1990_hup]. To make sense of such information, it accesses familiar information similar to what it is currently attempting to process for efficiency [@kahana_et-al_2022_ohhum]. This is accessing memories and contains valanced information [@kensinger_fields_2022_ohhum]: should I avoid this, or is it pleasant? Once the brain has finished such processing, it can pass its prediction to your conscious memory. Once you form a reflex of avoid or approach, this opens up space for your brain to process the more complex information: to take the patterns of the light as shapes that construct symbols and letters. This comes later because this information not only requires access to information about what it *is* but also what it *means*; once you understand what it means, you have the information necessary to evaluate it.

The snap-judgment model predicts that you first process the colors of the yard sign. You access associative memory to figure out what those particular wavelengths represent: red, white, blue? As these colors are associated with different emotional states [see @cimbalo_et-al_1978_jgp] and the resulting behavioral consequences, your brain starts sending signals to the rest of your body to prepare it to react [see @sander_2013_chhan; @dror_2017_er]. You now need to figure out what the rest of that information was. What were the patterns of that light? It appears that there were some white letters on the sign. There was an "E," a "L," an "E," a "C," and a "T." That creates the word "ELECT." Meaning to vote for. There were some more letters on the sign: an "M," an "I," a "T," a "C," and an "H." A name. The full name is "Mitch McConnell". Since it is about politics, it must be a politician named Mitch McConnell. Now imagine the information was the same, except the color was blue. You may take more time to figure out how that Mitch McConnell person is and come to your reaction to seeing their yard sign. This is because, without the color red, you first are thinking about Democrats who are named Mitch McConnell. Only when you come up empty on your mental Rolodex do you figure out that it is the Republican Mitch McConnell.

How do you react to the color and then to the name? Social groupings are not simply abstract concepts invented by social psychologists; our neurobiology reflects them. For example, researchers find activation of parts of the brain, such as the anterior insula, when we see someone in our social group outperformed by someone from the out-group [see @zink_barter_2012_ohsn]. The anterior insula activity is associated with physical and emotional pain, not just for ourselves but also for others [@adolphs_janowski_2011_ohsn]. Others have also observed that when seeing someone part of a high-status social group, there is an increase in activity in the sensorimotor cortex and supplementary motor area, indicating more activity in the areas of the brain that encourage movement [@zink_et-al_2008_n]. Visual information about someone in your social group speeds up processing, is more salient, and demands more attention than visual information about an object outside your social group [@zink_barter_2012_ohsn].

There is significant evidence supporting the theory that our partisan identification reflects more than just our attitudes about politics but a social identity [see @campbell_et-al_1960_jws; @mason_2018_cup] that guides our attitudes [see @achen_bartels_2016_pup; @white_et-al_2014_apsr; also @bullock_2011_apsr]. As our political attitudes reflect shared views among co-partisans [@pickup_et-al_2020_pb], the congruence between political information and our partisan identification influences our reactions to such political information. This means that the visual information we glean from politics will likely motivate those neurological features of social groups and explain the resulting behavioral manifestations reacting to such information. That information is also likely to be processed at different rates as well. That is, while visual information carries general affective associations, we should also expect associations between politically-relevant colors with partisan identification.

+--------------+-----------------------------------------------------------------------------------------------------------------------------------------------------+
| Hypotheses   | Expectation                                                                                                                                         |
+==============+:===================================================================================================================================================:+
| $H_1$        | People notice color                                                                                                                                 |
+--------------+-----------------------------------------------------------------------------------------------------------------------------------------------------+
| $H_2$        | Colors shape perceptions of candidate based on partisan associations                                                                                |
+--------------+-----------------------------------------------------------------------------------------------------------------------------------------------------+
| $H_3$        | Candidates using Red are more supported by Republicans; candidates using Blue are more supported by Democrats                                       |
+--------------+-----------------------------------------------------------------------------------------------------------------------------------------------------+
| $H_4$        | Republicans spend less time evaluating candidates using Red; Democrats spend less time evaluating candidates using Blue                             |
+--------------+-----------------------------------------------------------------------------------------------------------------------------------------------------+
| $H_5$        | Campaigns recognize the importance of color to voters' evaluations of candidates and respond strategically                                          |
+--------------+-----------------------------------------------------------------------------------------------------------------------------------------------------+

: Summary of hypotheses {#tbl-hypotheses}

From this discussion, I expect the following: that the average potential voter pays attention to the colors used in campaign branding ($H_{1}$); that these colors that they notice shape perceptions about the person and ideological symbol represented in the branding -- meaning that they express different levels of preference for receiving more information that is similar to what they saw and their levels of preference for supporting such a campaign ($H_{2}$); that the consistency of information explains more positive perceptions -- simple visual information with more complex "traditional" information ($H_{3}$); this positive and consistent information is processed quicker than negative and inconsistent, negative and consistent, and positive and inconsistent information ($H_{4}$); and finally that campaigns make strategic choices about their branding to attract voters ($H_{5}$) in line with their primary objective of reelection [@fenno_1973_lb; @mayhew_1974_yup]. A summary of these hypotheses are included in @tbl-hypotheses.

## Study 1

```{r}
#| label: study-1-setup

# Load some handy functons
box::use(
    data.table[...]
    , modelsummary[
        datasummary_skim
        , datasummary_balance
    ]
    , ./R/predicted_prob_bar[...]
)

# Execute the cleaning script
source("./eda/study_1/cleaning.R")
# Load fitted models
load("../data/temp/models/study_1_fitted.RData")
```

The purpose of [Study 1](#sec-study-1) is to examine whether the color choices on an almost ubiquitus form of campaign branding -- yard signs -- influence people's perceptions of political candidates. Evidence suggests that yard signs *do* matter in shaping political attitudes [@makse_et-al_2019_oup], vote intentions [@makse_et-al_2019_oup], and even electoral outcomes [@green_et-al_2016_es]. As they are simple, cheap, intrusive, and common forms of campaign branding, they provide a conservative test of the effects of color in campaign branding. Using yard signs in this study, I test the first four hypotheses I derive from the snap-judgment model. First, [Study 1](#sec-study-1) tests the claim that individuals do indeed notice the color of electoral yard signs. Next, it tests the claim that these colors that individuals detect influence the viewer's evaluations of the yard sign and the candidate represented on it. [Study 1](#sec-study-1) then tests the claim that the effects on perception are moderated by how consistent the color is with the more complex information displayed on the yard sign. Finally, [Study 1](#sec-study-1) examines whether positive information that contains consistency is processed quicker than negative or inconsistent information. That is, do partisans quickly detect and encode information from a clearly co-partisan political candidate?

### Research design

I recruit participants from Prolific.[^color_influence_pre-reg_anon-3] After providing informed consent to participate in the study, Prolific redirects subjects to Pavlovia[^color_influence_pre-reg_anon-4] and are provided with a demographics and political attitudes questionnaire. As I am concerned about priming effects introduced by these questions, while I am simultaneously concerned about bias introduced by post-treatment control [@montgomery_et-al_2018_ajps], I randomly select half of the participants to receive the questionnaire post-treatment and the other half receive it pre-treatment. This questionnaire includes common questions about the participant's ascriptive and descriptive characteristics and about the participant's political ideology, partisan identification, interest in politics, patriotism, and political knowledge.

[^color_influence_pre-reg_anon-3]: I pay subjects a rate of \$12.00 per hour. On top of the price per participant, Prolific charges a 30% servicing fee.

[^color_influence_pre-reg_anon-4]: Pavolovia allows for researchers to host and run open source experiments for about \$0.20 per participant (to cover their server costs). I use it primarily to integrate the JavaScript components from the jsPsych package [@jspsych] for my experimental design.

I include those questions in the questionnaire due to expectations that they may act as confounds in my hypotheses. I include some questions collecting information on participants' ascriptive and descriptive characteristics such as age, education, gender identity, and racial identity, as a number of these are correlates with partisan identification [see @campbell_et-al_1960_jws; @mason_2018_cup]. Though political knowledge does predict the strength of an individual's partisan identification [@lodge_taber_2013_cup], attention paid to politics is likely a more important confounding variable here. The purpose of color in politics is to act as a heuristic for people (a cognitive shortcut). If individuals are aware of the tendency for Republicans to use and be associated with the color red and Democrats with the color blue, the degree to which they know factual information about politics is not entirely relevant. On the other hand, those that pay no attention to politics, whatsoever, may be less likely to be aware of such associations. Additionally, those who express less interest in politics are also likely to identify as more moderate when asked about their ideology and partisan identity [@klar_krupnikov_2016_cup].

Additionally, I include a question about the respondent's sex assigned at birth and about whether they have received a diagnosis of any color blindness. As some individuals may possess undiagnosed colorblindness, asking about their sex assists in covariate balance. I additionally include an open-ended question asking participants to describe their "first memory of a political event." The use of an open-ended question helps provide an attention check and identify duplicated responses for those spoofing IP addresses with a VPN [see @kennedy_et-al_2021_poq]. Given estimates using these exclusion criteria, I would expect that upwards of 40% of my original sample will fail these attention checks [@kennedy_et-al_2021_poq]; thus, the large sample size.

I then present participants with an instruction screen informing them of the task for the experiment. In the first trial of the experiment, participants I randomly presented participants with two of three possible yard signs, one at a time. These yard signs are simple with the text "Riley Ready to Lead" and a solid background color of either "Republican Red," "Democratic Blue," or White.[^color_influence_pre-reg_anon-6] There is an added component to this, however.

[^color_influence_pre-reg_anon-6]: See the supplementary information to view all of the stimuli used in [Study 1](#sec-study-1).

Rather than use eye-tracking devices and software, I instead use Mouseview.js [see @anwyl-irvine_et-al_2022_brm], which either blocks out or blurs a large portion of the participant's screen and encourages them to move their mouse to view different parts of the screen in isolation. As the participants move their cursor around the screen, it tracks the coordinates of the cursor along with the "dwell" time of the cursor in that particular coordinate. One primary benefit of Mouseview.js is that it allows researchers to field their experiments outside of a lab-based setting -- while providing results that robustly correlate with the results from a design employing eye-tracking hardware [@anwyl-irvine_et-al_2022_brm]. This allows researchers to rely less on convenience samples, which are common with eye-tracking studies. For my design, I am particularly concerned about reliance on a convenience sample due to variations in participants' ability to detect and process color in the U.S. population relative to a student sample. This means that Mouseview.js is a handy tool for my study. A published pilot version of the experiment used in [Study 1](#sec-study-1) on [Pavlovia](https://run.pavlovia.org/damoncroberts/diss_ch_1_pre-reg/?__pilotToken=c74d97b01eae257e44aa9d5bade97baf&__oauthToken=9a645f267f9b83c8ebb48e60093bae5a746da56a6c824b8343389a9329fa2365).

When viewing each yard sign in the first trial, there is a blur over a substantial portion of the screen. At any given point in time, participants can view only 8% of the image without an obstruction, which simulates the observation that we typically foveate on about 8% of our available visual field at any given time [@wedel_pieters_2008_rmr].[^color_influence_pre-reg_anon-7] Participants move their cursor to explore the yard sign. I allot 5000 ms to perform the exploration until the image goes away to encourage a consistent and short duration to explore the image.[^color_influence_pre-reg_anon-8] After exploring each image, I asked participants what colors were on the yard sign and whether they felt that the candidate represented on the yard sign was a Democrat, a Republican, or Neither. After viewing both images, I ask subjects to indicate their preference among the two images. To ensure that participants have a standardized initial placement of their cursor, I display a blank page before viewing the yard sign that requires participants to click a "Next" button. Immediately after clicking "Next," participants are shown the yard sign. The goal is to ensure that variation in where participants explore the image is not dependent on a non-standard starting point for their cursor. I additionally utilize a gaussian blur for the overlay of the image rather than a solid overlay obstructing the participant's view. This gaussian blur allows participants to see a blurred visual field beyond the cursor. This allows participants to see enough to take purposeful action to explore blurred parts of the image that attract them [@anwyl-irvine_et-al_2022_brm]. The use of the gaussian blur requires that participants use a web browser other than Safari because of a known issue [@anwyl-irvine_et-al_2022_brm]. This will require participants to either switch browsers or to not participate in the study if they are using Safari at the time they are recruited by Prolific to participate in the study. As this requirement is enforced *before* joining the study, this should not have an effect on the number of excluded participants from my original sample.

[^color_influence_pre-reg_anon-7]: I include a screenshot providing an example of what the participants are able to see with the blur included in the supplementary materials.

[^color_influence_pre-reg_anon-8]: In marketing research, some studies give participants about 6000 milliseconds in eye-tracking studies to examine a brand and to formulate an intention to purchase a product or not [@wedel_pieters_2008_rmr]. With Mouseview.js, a study examining the tool's correlation with optical responses to viewing disgust-and-pleasure-evoking images uses 10000 milliseconds; but is intended to be an extended amount of time [@anwyl-irvine_et-al_2022_brm].

There are three more trials that are much like the first trial. What is different between the two other trials is that I vary the amount of color that is on the yard signs (trials 2 and 3). I provide more textual information that deviates from the association of Republicans with red and Democrats with blue (trial 4). Examples of all of these yard signs are included in the appendix. For readers interested in the more technical details, the appendix acommpanying this chapter also includes a table of the data collected in the experiment -- @tbl-study-1-measures; reports full results, posterior predictive checks, and more detailed discussions for all of the models used in the remainder of this section.^[Unless noted otherwise, the reflection of uncertainty in the models are credible intervals at the 90% level. This should not be equated with the frequentist practice of often reporting confidence intervals at the 95% level as credible intervals at this level often are extremely unstable and, more importantly, are interpreted differently. For example, a credible interval at the 90% level can be interpreted as: the probability that the true effect is in that interval. This contrasts with a confidence interval used in the frequentist framework which is interpreted as: upon repeated samples, 90% of the confidence intervals contain the true effect. As I am using two-sided credible intervals here, at the 90% level, the probability that 0 would be the true value given the data would be less than 5% as it would be on one side of my posterior distribution.]

@tbl-study-1-descriptives presents some characteristics of the sample on the whole. The sample is relatively young with a mean age of 36.6 years old, few report being diagnosed with colorblindness, about 40% of the sample identifies as female, 70% of the sample identifies as White while only 10% report being Hispanic and Black. The average respondent in the sample reports leaning Democratic.

```{r}
#| label: tbl-study-1-descriptives
#| tbl-cap: Descriptive statistics of Prolific sample

df_demographics_subset <- list_df[["cleaned"]][
    , .(
        Age = age
        , `Color blind` = color_blind_dummy
        , Female = female_sex
        , White = white
        , Hispanic = hispanic
        , Black = black
        , `Party ID` = pid_7
    )
] |> 
as.data.frame()

datasummary_skim(
    df_demographics_subset
    , histogram = FALSE
    , notes = c(
      "Data source: Prolific."
      , "Sample characteristics."
    )
    , output = "latex"
)
```

### Do individuals notice color in political branding?

To address this first question, I use a couple of proxy measures for people's attention paid to the yard signs. I collect the first measure through recording the movement of the subjects' cursors along the x-and-y-axes. The second measure accounts for the amount of time participants' cursor lingers over a particular coordinate on their screen. Presumably, the more participants explore the yard sign, the more they are finding visual information that they can use to evaluate the candidate. While these two measures are simple proxies and are not sufficient measures of "noticing the yard sign," they allow us to get a sense of the patterns by which people explore the yard signs across treatment conditions. It also helps with my confidence in the degree to which variation between the treatments among different groups of participants are due to actual differences in how people processed these treatments and not due to random chance.

```{r}
#| label: tbl-h1-trial-1
#| tbl-cap: Average time between cursor movements (Trial 1)
df_movement_trial_1 <- list_df[["cleaned"]][
  , .(
    trial_1_difference_Time
    , trial_1_stimulus
    , trial_1_difference_X
    , trial_1_difference_Y
  )
]
setnames(
  df_movement_trial_1
  , old = c(
    "trial_1_difference_Time"
    , "trial_1_difference_X"
    , "trial_1_difference_Y"
  )
  , new = c(
    "$\\Delta_{t}$ (milliseconds)"
    , "$\\Delta_{x}$ (pixels)"
    , "$\\Delta_{y}$ (pixels)"
  )
)

datasummary_balance(
  ~trial_1_stimulus
  , data = df_movement_trial_1
  , notes = "Data source: Prolific sample."
  , escape = FALSE
  , output = "latex"
)
```

```{r}
#| label: tbl-h1-trial-2
#| tbl-cap: Average time between cursor movements (Trial 2)
df_movement_trial_2 <- list_df[["cleaned"]][
  , .(
    trial_2_difference_Time
    , trial_2_stimulus
    , trial_2_difference_X
    , trial_2_difference_Y
  )
]
setnames(
  df_movement_trial_2
  , old = c(
    "trial_2_difference_Time"
    , "trial_2_difference_X"
    , "trial_2_difference_Y"
  )
  , new = c(
    "$\\Delta_{t}$ (milliseconds)"
    , "$\\Delta_{x}$ (pixels)"
    , "$\\Delta_{y}$ (pixels)"
  )
)
modelsummary::datasummary_balance(
  ~trial_2_stimulus
  , data = df_movement_trial_2
  , notes = "Data source: Prolific sample."
  , escape = FALSE
  , output = "latex"
)
```

```{r}
#| label: tbl-h1-trial-3
#| tbl-cap: Average time between cursor movements (Trial 3)
df_movement_trial_3 <- list_df[["cleaned"]][
  , .(
    trial_3_difference_Time
    , trial_3_stimulus
    , trial_3_difference_X
    , trial_3_difference_Y
  )
]
setnames(
  df_movement_trial_3
  , old = c(
    "trial_3_difference_Time"
    , "trial_3_difference_X"
    , "trial_3_difference_Y"
  )
  , new = c(
    "$\\Delta_{t}$(milliseconds)"
    , "$\\Delta_{x}$ (pixels)"
    , "$\\Delta_{y}$ (pixels)"
  )
)

modelsummary::datasummary_balance(
  ~trial_3_stimulus
  , data = df_movement_trial_3
  , notes = "Data source: Prolific sample."
  , escape = FALSE
  , output = "latex"
)
```

@tbl-h1-trial-1, @tbl-h1-trial-2, and @tbl-h1-trial-3 report the average difference in milliseconds and pixels along the x-and-y-axis of the subjects' screen when exploring the yard signs. These tables demonstrate a few things. First, the average participant in each treatment condition appear to be spending more time exploring the yard signs when there are more "conflicting" colors (trials 2 and 3). Second, we see that, for the most part, the average participant in the Red and Blue yard sign conditions spend more time exploring those yard signs than the participants viewing the White yard sign.

Overall, it appears that people viewed the yard signs differently depending on which yard sign they were looking at. Now, I turn to exploring how features about the subjects themselves may shape the way they interact and respond to this visual information on the yard signs. Though, I will note that the data on the timing and cursor location is much more noisy than I had originally anticipated. However, as the tests of the next few questions demonstrate, there seem to be systematic differences between individuals that saw either a red, white, or blue yard sign despite holding everything else but the color constant between them. While I am a bit skeptical of the data here as it is noisy, I think the next peices of empirical evidence suggest support for this hypothesis as well.

### Do colors shape perceptions of political objects?

To address the question of whether the colors affect perceptions of the candidate and the yard sign, I ask participants to report whether they perceived the candidate to be a partisan -- either Republican, Democrat, or as non-partisan immediately after viewing each image. Everything on the yard signs remains constant except for the color. As representations of ideology are associated with more than just political views but things like space [@mills_et-al_2016_bbr] and color [@maestre_medero_2022_pr], differences between respondents on the perceived political affiliations of the candidate should be more than "by chance" differences. However, differences occur based on the associations between red and blue with partisanship and the lack of political information that the color white conveys.

I examine differences in respondents' reported perceptions of the candidate's partisan affiliation. Evidence supporting $H_2$ would appear as differences between those who saw the white yard signs and the "partisan color" yard signs in reported perceptions of the partisanship of the candidates on average. I estimate two ordered logistic regression where the main independent variables of interest are two indicator variables for whether respondents received the "red yard sign" or the "blue yard sign" in each round of Trial 1. The baseline category for both of those indicator variables is if they instead saw the "white yard sign".

@fig-h-2-predictions present the predicted probabilities of an average subject reporting the candidate was either a Democrat, Independent, or Republican when viewing the different yard signs. The error bars reflect the high density credible interval for 95% of my posterior draws. Each bar reflects the predicted probability of an outcome -- listing the candidate is either a Democrat, Independent, or Republican. When interpreting the error bars, you will want to look at the difference in the error bars *between* yard sign treatments rather than *within* treatment. We see that the models indicate statistically meaningful differences between the yard sign treatments for each outcome. That is, the red and blue yard signs did shift people's perceptions of the partisanship of the candidate relative to the white yard sign. The credible intervals of these models are reported in the appendix in @tbl-hypothesis-2-main which contains the rest of the results from these models.

Digging in a little bit more about what these perceptions were, the predicted probability that subjects perceive the candidate with a Red yard sign to be Republican is almost 70%, whereas it is only about 30% when viewing a white yard sign. When given a red yard sign, the predicted probability the candidate is marked to be an independent is less than 30% and less than 5% to be a Democrat. However, when viewing a blue yard sign, the predicted probability of a subject perceiving the candidate to be a Democrat is at about 50% whereas it is less than 20% with a white yard sign. When viewing a blue yard sign, the predicted probability of subjects perceiving the candidate to be a Republican is less than 10% and about 45% to be an Independent. The full results are reported in @tbl-hypothesis-2-main in the appendix.

```{r}
#| label: fig-h-2-predictions
#| layout-nrow: 1
#| fig-cap: The partisanship of candidates change depending on the color of the yard sign
#| fig-subcap:
#|   - Red treatment
#|   - Blue treatment
red <- predicted_prob_bar(
  fitted_model = list_fitted[["h_2"]][["red"]]
  , x_axis = "trial_1_red_stimuli"
)
blue <- predicted_prob_bar(
  fitted_model = list_fitted[["h_2"]][["blue"]]
  , x_axis = "trial_1_blue_stimuli"
  , treatment = "Blue"
)
red
blue
```

As the consistency by which the parties have used either the color red or blue has increased since the 2000's, I suspect that the age of my subjects may moderate this relationship. That is, the tendency for subjects to label red yard signs as a Republican candidate versus a different color may be stronger for those who are younger as the consistency in branding is the political environment they have been around more. I expect this particular relationship may be weaker for older subjects as they may place current branding strategies in a larger political context where this was not always the case.

I fit two ordered logistic regressions where the independent variables are indicator variables reporting whether the participant received the red (the first model) or they recieved the blue treatment (the second model). Those included in the baseline category for both models are those who received the white yard sign treatment. Since I believe that age may moderate this effect, I include an interaction term for both of these models that multiplies these treatment indicator variables with age.

Though my pre-registered expectations were that age would moderate this effect, these models suggest that there are not substantive or statistically meaningful differences in how older and younger participants responded to these treatments. The results of these models are included in the appendix in @tbl-hypothesis-2-main.

I also explore whether these effects may be weaker for those with colorblindness. The results of this model is reported @tbl-colorblindness, which is located in the appendix. The results do not suggest that they are. However, given that I had a very small number of respondents reporting colorblindness in my sample, the lack of substantive or statistical effects here are likely because I do not have enough participants with colorblindness in each group to compare.

I also test whether my effects here are sensitive to my choice to run two separate regression models -- one for each treatment. I fit a single ordered logistic regression with two indicator variables for the treatment coded as: 0 did not receive red/blue treatment, 1 did receive treatment. The full results of this alternative model specification is included in @tbl-robustness-alt-specification-h-2 This means that the intercept term for the regression would be the predicted probability of the average respondent labeling the owner of the yard sign as either Democrat, Independent, or Republican. The conclusions I draw from this regression model do not change the ones I make here with @fig-h-2-predictions.

In summary, I find that subjects perceive candidates with red yard signs as Republican, candidates with blue yard signs as Democratic, and candidates with white yard signs as being neither -- an Independent. These yard signs are *very* red and blue, however. In the next section, I examine how these effects may be weaker when we use less red or blue and include more blue or red. That is, see if using both colors on the yard signs has the capacity to increase uncertainty about the perception of partisan affiliation.

### Do these perceptions require consistency between information types?

The design of the latter trials in the study presents yard signs with mixtures of non-partisan and out-partisan colors. For example, primarily red and presumably Republican yard signs, but has some blue or white in them. I should expect that the trials that use less "consistent" visual information demonstrate more ambivalence among respondents in their reported perceptions of the politician's positions. Specifically, I expect that participants will perceive the yard signs that have both red and blue on them as more moderate and that the higher proportion of the color red or blue among the two colors will lead respondents, on average, to be more likely to believe that the candidate leans more Republican or Democratic. This is not to say that I expect the results to change significantly, but rather that the certainty by which I can predict individuals' guesses of the candidate's partisan affiliation to be weaker.

I fit four ordered logistic regression models. Two for each trial. Where each model for each trial calculates whether an average subject in the study reported that the yard sign they saw in the trial was a Democrat, Indepdent, or a Republican. The predicted probabilities for each outcome occuring under different treatments and for different trials are presented in @fig-h-2a-predictions. @tbl-hypothesis-2a and @tbl-hypothesis-2a-cont located in the appendix reports the full results of these models while @fig-h-2a-ppc located in the appendix reports the posterior predictive checks of these models.

```{r}
#| label: fig-h-2a-predictions
#| layout-nrow: 2
#| fig-cap: Mixing red and blue in a yard sign increases uncertainty about partisanship
#| fig-subcap:
#|   - Trial 2 Red treatment
#|   - Trial 2 Blue treatment
#|   - Trial 3 Red treatment
#|   - Trial 3 Blue treatment
plot_trial_2_red <- predicted_prob_bar(
  fitted_model = list_fitted[["h_2a"]][["trial_2_red"]]
  , x_axis = "trial_2_red_stimuli"
)
plot_trial_2_blue <- predicted_prob_bar(
  fitted_model = list_fitted[["h_2a"]][["trial_2_blue"]]
  , x_axis = "trial_2_blue_stimuli"
  , treatment = "Blue"
)
plot_trial_3_red <- predicted_prob_bar(
  fitted_model = list_fitted[["h_2a"]][["trial_3_red"]]
  , x_axis = "trial_3_red_stimuli"
)
plot_trial_3_blue <- predicted_prob_bar(
  fitted_model = list_fitted[["h_2a"]][["trial_3_blue"]]
  , x_axis = "trial_3_blue_stimuli"
  , treatment = "Blue"
)
plot_trial_2_red
plot_trial_2_blue
plot_trial_3_red
plot_trial_3_blue
```

The results from these models offer partial support for my expectations. First, I see that yard signs using less blue are not as distinguishably Democratic as the white yard sign. In other words, voters are not meaningfully more likely to report that the yard sign with less blue on it belongs to a Democratic candidate as a white and black yard sign. So, the strength of using a blue yard sign to communicate alignment with the Democratic party weakens if you use less blue relative to white or include red on it. With red yard signs, however, we see that voters still see them to be belonging to Republican candidates even when we reduce the amount of red on the yard sign and even include blue.

One potential reason for this may be the association of the American flag when using all three colors. Particularly given the Republican party's tendency to condemn those who do not adhere to social norms about showing respect to the American flag. This, however, is untestable based on the data I did not collect for this study. So, this explanation is speculation. However, what we do see is that there are bounds by which at least the color blue on a yard sign indicates to voters that the candidate is a Democrat. What remains unclear is whether these perceptions of ownership by a partisan candidate influences people's willingness to support that candidate or to express favoritism towards it ($H_3$).

### Does the color of a yard sign explain support for a candidate?

An important implication of the snap-judgment model suggests that if individuals perceive candidates using the color red as Republican while candidates using the color blue as Democratic, that partisans may be more attracted to yard signs displaying the color that fits with these associations. That is, if the color red conveys group membership to the Republican party, we should expect that Republican individuals should like the red yard signs more and be more willing to vote for candidates using those colors in their campaign materials. We should expect a similar thing with Democrats but with blue yard signs. Another expectation is that we should expect these relationships to be relatively strong despite little substantive information (e.g., policy positions and name recognition) about the candidate if color does indeed matter as a form of political information.

After the three trials, I asked participants which of the three yard signs they would vote for. I create an indicator variable from this question that documents whether participants would vote for the candidate that owns the first yard sign (1) or not (0). I fit two logistic regression models with the vote variable as the outcome I am trying to explain. In the first of the two logistic regression models, I examine whether the effect that receiving a red versus a white stimuli is moderated by the participants' partisanship. As partisanship and its moderating effect on the treatment is of interest here while not being randomly assigned, I control for a number of confounding variables such as attention paid to politics, age, gender identification, and racial identification. The second logistic regression model is much like the first. The difference is that instead of examining the moderated effect of the red treatment, I examine the moderated effect of the blue treatment. As I am still interested in examining how partisan identification modifies the relationship between support for the candidate when comparing a blue versus a white yard sign, I include the same controls as I did with the first logistic regression model. I include the full table of results for both of these models in @tbl-h-3-main which is located in the appendix.

```{r}
#| label: fig-h-3-predictions
#| layout-nrow: 1
#| fig-cap: Probability of voting for candidate with yard sign
#| fig-subcap:
#|   - Red treatment
#|   - Blue treatment
red <- predicted_prob_bar(
  fitted_model = list_fitted[["h_3"]][["red"]]
  , x_axis = "trial_1_red_stimuli"
  , treatment = "Red"
  , hypothesis = "H3"
  , x_label = "Party ID"
  , y_label = "Pr(Vote for candidate)"
  , legend_title = "Treatment"
)
blue <- predicted_prob_bar(
  fitted_model = list_fitted[["h_3"]][["blue"]]
  , x_axis = "trial_1_blue_stimuli"
  , treatment = "Blue"
  , hypothesis = "H3"
  , x_label = "Party ID"
  , y_label = "Pr(Vote for candidate)"
  , legend_title = "Treatment"
)
red
blue
```

@fig-h-3-predictions presents the average predicted probability that respondents who vary on treatment assignment and partisan identification would support the candidate. The results fit with my expectations. First, I see that the predicted probability that a strong Democrat would vote for a candidate with a red yard sign is less than 10% while for strong Republicans it is at about 70%. Likewise, we see that strong Democrats much prefer a candidate with a blue yard sign than a candidate with a white yard sign; 80% relative to 40%, respectively. Strong Republicans are unlikely to vote for a candidate with a blue yard sign and a white yard sign with their predicted probability of voting for either of those candidates below 30%.

### Do partisans process co-partisan branding faster?

The other analytical tasks do not examine expectations derived from the motivated reasoning portion of the model. That is, do people process in-group information faster than out-group information? There is evidence of these tendencies in political circumstances [@lodge_taber_2013_cup].

To examine whether motivated reasoning is also active with the processing of politically-relevant color, I examine the difference between the amount of time between the start of viewing a stimulus and clicking "Next" to stop viewing the stimulus among those who were viewing a presumed co-partisan yard sign relative to those viewing a presumed out-partisan yard sign. As the outcome is a count of the amount of time participants spent looking at a yard sign, I turn to a class of regression models that work well for explaining outcomes measured as count data. By the nature of my measure, I do not believe that participants can record a value of 0 as time elapsed is measured in milliseconds. Additionally, I expect that there will be what is referred to over-dispersion as I expect to have a large bulk of participants recording a very small time elapsed while there may be others who may record quite a large time elapsed. In the appendix, I include @tbl-h-4-dispersion which summarizes the time elapsed outcome variable; which indeed demonstrates over-dispersion in this variable with the outcome being the time elapsed. Given this, rather than using a count model assuming a poisson-distributed outcome, I instead assume an outcome distributed as poisson-gamma. I therefore fit two poisson-gamma models (sometimes referred to as negative binomial regression). I am interested in testing whether time elapsed can be explained by not only the color of the yard sign that the subject was viewing, but also that this effect is moderated by their partisanship. I therefore include an interaction term between stimuli and the partisan identification of the respondent. Like before, as partisan identification is not randomly assigned, I control for confounds such as attention paid to politics, age, gender identity, and racial identity. @tbl-h-4 located in the appendix is the full table of results for these two poisson-gamma regression models. @fig-h-4-ppc also indicates that the model does a quite good job at predicting the distribution of the time elapsed outcome variable.

```{r}
#| label: fig-h-4
#| fig-cap: The effect of party congruency on time spent looking at yard sign
#| layout-ncol: 2
#| fig-subcap:
#|   - Red treatment
#|   - Blue treatment

red <- predicted_prob_bar(
  fitted_model = list_fitted[["h_4"]][["red"]]
  , x_axis = "trial_1_red_stimuli"
  , treatment = "Red"
  , hypothesis = "H4"
  , x_label = "Party ID"
  , y_label = "Predicted amount of time elapsed"
  , legend_title = "Treatment"
)
blue <- predicted_prob_bar(
  fitted_model = list_fitted[["h_4"]][["blue"]]
  , x_axis = "trial_1_blue_stimuli"
  , treatment = "Blue"
  , hypothesis = "H4"
  , x_label = "Party ID"
  , y_label = "Predicted amount of time elapsed"
  , legend_title = "Treatment"
)
red
blue
```

@fig-h-4 presents the average predicted elapsed time for participants in different contexts. Overall, these results communicate that I do not find conclusive support for my expectations on $H_4$ in either regression model. Strong Democrats do appear to spend more time looking at red yard signs than white ones, once I account for the variation in the data, I cannot be sure that these differences are not due to random chance. I also see that there are not really any differences (substantive or statistical) between partisans looking at blue versus white yard signs.

## Study 2

[Study 1](#sec-study-1) presents evidence in support of the idea that the colors red and blue matter to politics -- they convey information to people. Second, the evidence from [Study 1](#sec-study-1) suggests that the information can shape the attitudes of voters. What remains unclear is whether or not these patterns have large enough impacts that it shapes how campaigns market themselves. In informal conversations with a campaign marketing firm which specializes in campaign marketing, I was told:

> There obviously is a strong association with blue and the Democratic Party and red and the Republican Party.  We've seen in some of our testing that specifically MAGA red tends to elicit a slightly stronger association with Republicans/conservatives/Trump than other shades. 
> The rest of color spectrum doesn't seem to have as strong an association, though greens and purples tend to read a little more left, the combination of red, white and blue tends to read a little more right, black and white tends to read a little left, black and red tend to read a little more right etc.

While this was good information to hear that seemed to align with what I found in [Study 1](#sec-study-1), I wasn't sure how much this actually influenced strategy. It appears it does, however.

> Beyond that,  I generally suggest that considering associations with party is generally a good place to start.  For example, if you're running in a Democratic primary and that's the race that's going to most likely decide who ultimately gets elected, then blues are a good place to start.

While there is some qualification in that there are other colors that can be useful beyond red and blue. The value of colors associated with a particular place can be a way for candidates to communicate attachment to the district and to the constituents. Though, there is a preference for campaigns to use as few colors as possible to save on costs - simpler is better. What particularly stood out to me was when I was told

> You want voters to be able to look at whatever collateral you have, including lawn signs, and be able to identify that it's your material without ever having to read it. If you do one thing long enough, voters tend to start to associate specific color schemes with specific candidates.

So, unsurprisingly, campaigns care about how they market themselves. And it seems as though campaigns want to tap into quick and efficient information processing for voters and potential voters. However, as noted during the conversation

> A lot of decisions are still made on purely gut feelings as opposed to taking a more multi-disciplinary evidence-based approach.

The goal for the second peice of empirical evidence in the chapter is to explore whether there is any evidence not from the "lab" suggesting that campaigns use the colors red and blue systematically. If the color of a yard sign does indeed matter and its electoral effects depend on the perceptions of potential voters, we should expect that campaigns are less likely to place yard signs with a large amount of Red or Blue in more moderate districts. However, in districts where there is a strong partisan base, we should expect that the candidate of the popular party will use the party's color more, while the disadvantaged candidate will use their party's color less as a way to either play up or down their loyalty to their party.

First, I collect data from the MIT Election Lab that reports the electoral results of House of Representative elections from the years 1970-2020. The data here tell me which candidate got how many votes in each House election across the United States. With this data, I calculate the rolling 5-year average of the Democratic party's vote share for each district.

Next, I download around 900 yard signs from House elections from the years 2016 to 2024 from the website provided by the Center for American Politics and Design. I use the `OpenCV` library from python to examine the proportion of the colors near the "Republican Red"^[lower values: 93, 9, 12; upper values 236, 69, 75] and "Democratic Blue"^[lower values: 0, 18, 26; upper values: 102, 212, 255] in each of these yard signs.

For example, I downloaded the GOP logo used on their official Twitter account during the 2022 midterm election cycle. What `OpenCV` does is it loads the image and converts it to a three-dimensional `numpy` array that contains the GBR (reverse of RBG) values for the pixels in the image. I standardize the image to be 224 $\times$ 224 pixels. The color is trained to detect the range of Republican Red and Democrat Blue colors. Once trained, the computer detects the pixels that do not contain values within the pre-specified range and converts those pixels to be black.

```{python}
#| label: example detection
#| echo: false

from sys import path # for path management
from cv2 import imread, imwrite # for color detection
    #* user-defined
#path.append("../")
from PY.helper import colorDetector

# Define colors to detect
    #* White
        #** Not defined. Default for colorDetector()
    #* Red
republican_red = [232, 27, 35] # target color
red_lower = [93, 9, 12] # lower end of spectrum for red
red_higher = [237, 69, 75] # higher end of spectrum for red

# Load gop image to read
img = imread("./assets/img/gop_2022.png")

# Detect colors
percent, img_transformed, result = colorDetector(img = img, color_upper = red_higher, color_lower = red_lower)

transformed= imwrite("../data/temp/study_2/gop_2022_transformed.png", img_transformed) # save transformed image

masked= imwrite("../data/temp/study_2/gop_2022_detected.png", result) # save masked image
```

::: {#fig-color-detection-example layout-ncol="2"}
![Resized original image](../data/temp/study_2/gop_2022_transformed.png){#fig-resized}

![Masked](../data/temp/study_2/gop_2022_detected.png){#fig-masked}

Detecting colors in the GOP logo
:::

I then extract the non-black pixels in the array and calculate the percentage of pixels in the image that are not black.^[$\text{Color} \% = \frac{\text{Non-black}}{\text{Transformed}} \times \frac{\text{Original}_{\text{Height}} + \text{Original}_{\text{Width}}}{2\text{Transformed}_{\text{Height}} + 2\text{Transformed}_{\text{Width}}}$]

```{python}
#| output: asis
# Print calculated percentage of example
print("For the example in @fig-color-detection-example, about {:.2f}".format(percent) + " of the image is red.")
```

With the data on the proportion of each of the yard signs using the red and blue colors, I merge this data with the MIT Election Lab Data. This leaves me with data from the House elections between 2018 and 2022. While the data are limited in the number of elections it reflects, it allows me to hold the district composition constant as redistricting occurs after the Census every 10 years. However, each district and state is different in their laws surrounding elections and are different in terms of political context. I want to examine whether there are meaningful differences between districts that are partisan "strongholds" (Democrats or Republicans tend to get elected in a district) relative to districts that are much less so. To do this, I include a model with intercepts with no pooling. That is a model that includes an intercept for each state-district unit. Though my outcomes of interest (proportion of the yard sign that is red or blue) are continous (the outcomes can take any real number, so not an integer), the outcomes are bounded between 0 and 1. While the traditional approach may be to use ordinary least squares, such a technique often can lead to poor model fit and introduce systematic error into my model and lead to results that are quite unstable. To not bog down readers with a lot of technical jargon, the model that I end up using is called a ordered beta regression which allows me to efficiently and accurately predict outcomes that can be 0 or 1 or any real number between 0 and 1 [@kubinec_2022_pa]. For readers in a more technical discussion about why I chose this over other techniques in the beta regression family, I encourage you to look at the appendix.

Once I have fit these models, I use the `marginaleffects` [@marginaleffects] package to calculate the grand mean of the predicted proportion of the use of the color red or blue for these yard signs as explained by the 5-year rolling average Democratic candidates' vote share. The grand mean here just reflects that I am not showing the predicted proportion of the use of these outcomes for each state-district, but rather I am just averaging across all of those to provide a single line. These results are included in @fig-study-2-pred.

```{r}
#| label: fig-study-2-pred
#| layout-ncol: 1
#| fig-cap: More blue on the yard signs in Democratic strongholds
#| fig-subcap:
#|   - Proportion of red
#|   - Proportion of blue
red <- predicted_prob_bar(
  fitted_model = list_fitted[["h_5"]][["red"]]
  , treatment = "Red"
  , hypothesis = "H5"
)
blue <- predicted_prob_bar(
  fitted_model = list_fitted[["h_5"]][["blue"]]
  , treatment = "Blue"
  , hypothesis = "H5"
)
red
blue
```

The results demonstrate that districts where Democrats are historically more successful see yard signs with a lower proportion of red on them, while those same districts see more Blue on them. Again, these results do not claim that campaigns are doing this either because voters are encouraging them to take this strategy or that they are doing it as a way to proactively get voters to do it. To tease out which of those questions is the right one requires more sophisticated approaches and data that could allow me to leverage redistricting to see how these outcomes change due to some unrelated, external shock. The data available to me do not allow me to do that at this time. Furthermore, teasing out such an association is not a primary goal of this chapter.

## Conclusions

This chapter argues that the colors red and blue provide useful information to voters about politics. Specifically, the colors red and blue provide information about a candidates' partisanship. Second, the chapter argues that the information gleaned from these colors can shape how people engage with politics. While I had a number of other sets of expectations from this chapter, these were the two primary questions I wanted to address in the chapter.

Evidence from my experiment suggests that the use of the colors red and blue convey partisan information, while a third color like white provides very little useful political information. The experiment provides evidence that individuals see candidates that use red on a yard sign as more Republican and candidates that use blue as more Democratic. Evidence from the experiment also suggested that using less of these partisan colors and using both partisan colors do make it a bit harder for individuals to make those same associations. Further, the experiment demonstrated that these associations that we make with color and partisanship are powerful enough that they can shape whether people express a willingness to vote for a candidate. That is, individuals that identify with the Republican party are much more likely to vote for a candidate with a red yard sign than a white yard sign and Democrats are much more likely to vote for a candidate with a blue yard sign than yard signs with other colors.

Since it appears that colors can influence vote choice, I turn to data that capture the real world messiness to examine whether or not campaigns are responsive to such patterns. It appears that they are. The empirical evidence suggest that campaigns use more blue on their yard signs if it is a district that has a higher vote share for Democrats the five previous years and use red much less often in those same circumstances. This fits with strategies communicated to me by a campaign marketing firm.

What this chapter demonstrates is that the colors red and blue convey information and they shape outcomes. It appears as though that it does not simply shape outcomes for individuals such as their personal attitudes or their vote choice in a lab, but that it has enough of an aggregative effect that can shape strategy for campaigns. The next chapter seeks to dig into this latter claim a bit more. The next chapter examines whether the snap-judgment model holds up in more complex settings. Specifically, the next chapter examines whether color can shape whether and how we have conversations about politics with others and the degree to which color can be a tool inhibiting political persuasion.