### 2023-09-15 chapter_2/src/slides/index.qmd
slide drafting

---
title: "The Shape and Color of Politics"
author:
  - name: Damon C. Roberts [{{< fa brands orcid size=xs >}}](https://orcid.org/0000-0002-4360-3675)
format:
  revealjs: 
    embed-resources: true
    theme: [dark, ../../assets/slides/custom.scss]
    transition: slide
    scrollable: true
    incremental: true
    slide-number: true
    toc: true
    toc-depth: 2
    slide-level: 3
    bibliography: ../../../assets/references.bib
    footer: |
        <span class="faux-block">{{< fa brands creative-commons >}} 2021-2023 Damon Charles Roberts</span>
        <span class="faux-block"><br>All content licensed under {{< fa brands creative-commons >}} {{< fa brands creative-commons-by >}} {{< fa brands creative-commons-sa >}} [Creative Commons CC BY-SA 4.0](https://creativecommons.org/licenses/by-sa/4.0/)</span>
    self-contained: true
    output-file: color_and_conversation.html
execute:
  echo: false
---

```{r}
#| label: setup-block
box::use(
  ggplot2[
    ggplot
    , aes
    , geom_density
    , theme_minimal
    , geom_abline
  ]
)
```

## Visual politics is not new politics, but "hot" politics

---

[![1860s Ballots](../../assets/slides/imgs/ballots.avif){#fig-ballots}](https://slate.com/news-and-politics/2020/02/ballot-history-photo-essay.html)

---

::: {#fig-1980-convention layout-ncol=2}

[![Democratic National Convention](../../assets/slides/imgs/1980_dem_convention.jpg){#fig-1980-dem height=100}](https://untappedcities.com/2021/08/12/1980-democratic-national-convention/)

[![Republican National Convention](../../assets/slides/imgs/1980_rep_convention.jpg){#fig-1980-rep height=100}](https://reuther.wayne.edu/node/9207)

1980 Presidential Party National Conventions
:::

---

[![2000 Presidential Election map](../../assets/slides/imgs/2000_election_map.webp){#fig-2000-election-map width=200}](https://www.google.com/url?sa=i&url=https%3A%2F%2Fwww.smithsonianmag.com%2Fhistory%2Fwhen-republicans-were-blue-and-democrats-were-red-104176297%2F&psig=AOvVaw3mlqWWKalyCl0gzVJDCIV5&ust=1694829600914000&source=images&cd=vfe&opi=89978449&ved=0CA8QjRxqFwoTCJCNg7bCq4EDFQAAAAAdAAAAABAT)

---

::: {#fig-party-brand layout-ncol=2}

![Republicans pre-2000](../assets/img/rnc_logo_pre-2000.png){#fig-republican}

![Democrats pre-2000](../assets/img/dnc_logo_1960.png){#fig-democrat}

![Republicans 2023](../assets/img/rnc_logo_2023.png){#fig-republican-post}

![Democrats 2023](../assets/img/dnc_logo_2023.png){#fig-democrat-post}

Party logos
:::

---

[![The public](../../assets/slides/imgs/division.webp){#fig-public}](https://www.brookings.edu/articles/what-is-a-brokered-convention-what-is-a-contested-convention/)

## Question guiding the project

Is color a source of political information that shapes the public's behaviors and attitudes?

## De-mystifying some of the jargon I might use

- Object
  - refers to anything out in the world that we are interacting with
  - here, these objects are political
- Affect $\noeq$ emotion
  - refers to the larger state of being
  - emotion refers to consciously labeled reactions
  - valance refers to pre-consciously 
- "Hot Cognition"
  - refers to the idea that some memories and experiences are more accessible or come to the top of the mind quicker.
  - this coincides with memories and experiences that are affect-laden

## Dominant models of information processing

### Zaller's [- @zaller_1992_cup] Memory model

---

```{r}
#| label: sim-visualization

df <- fabricatr::fabricate(
  N = 10000,
  pop = rnorm(n = N, mean = 2, sd = 1),
  sample = rnorm(n = N, mean = 2, sd = 0.5),
  updated = rnorm(n = N, mean = 1.9, sd = 1)
)
```

```{r}
#| label: fig-pre-existing-info
#| fig-cap: Set of pre-existing attitudes
ggplot(data = df) +
  geom_density(aes(x = pop)) +
  theme_minimal()
```

---

```{r}
#| label: fig-sample-pre-existing-info
#| fig-cap: Pre-existing attitudes used

ggplot(data = df) +
  geom_density(aes(x = sample)) +
  theme_minimal()
```

---

```{mermaid}
graph TD;
A[Sampled pre-existing info] --> C[Updated attitudes]
B[New info] --> C
```

---

```{r}
#| label: fig-new-attitude
#| fig-cap: Updated set of attitudes
ggplot(data = df) +
  geom_density(aes(x = updated)) +
  theme_minimal()
```

### Lodge and Taber's [- @lodge_taber_2013_cup] Online model

## Amending those models with the Snap-judgment model