### 2023-09-18 chapter_2/src/slides.qmd
update site

---
title: "The Shape and Color of Politics"
author:
  - name: Damon C. Roberts [{{< fa brands orcid size=xs >}}](https://orcid.org/0000-0002-4360-3675)
format:
  revealjs: 
    embed-resources: true
    theme: [dark, ../assets/slides/custom.scss]
    transition: slide
    scrollable: true
    incremental: true
    slide-number: true
    toc: true
    toc-depth: 2
    slide-level: 5
    bibliography: ../../assets/references.bib   
    self-contained: true
    output-file: color_and_conversation.html
execute:
  echo: false
---

```{r}
#| label: setup-block
# Set seed
set.seed(12102023)
# Load libraries
box::use(
  ggplot2[
    ggplot
    , aes
    , geom_density
    , theme_minimal
    , geom_vline
  ]
  , modelsummary[
    datasummary_skim
    , datasummary_balance
    , modelsummary
  ]
  , ./R/predicted_prob_bar[...]
  , data.table[...]
)

# Execute the cleaning script
source("./eda/study_1/cleaning.R")
# Load cleaned data
list_df <- list()
list_df[["cleaned"]] <- read.csv(
  file = "../data/clean/cleaned_pavlovia.csv"
)
# Load fitted models
load("../data/temp/models/study_1_fitted.RData")
```

## Visual politics is not new politics, but "hot" politics

---

[![1860s Ballots](../assets/slides/imgs/ballots.avif){#fig-ballots}](https://slate.com/news-and-politics/2020/02/ballot-history-photo-essay.html)

---

::: {#fig-1980-convention layout-ncol=2}

[![Democratic National Convention](../assets/slides/imgs/1980_dem_convention.jpg){#fig-1980-dem height=100}](https://untappedcities.com/2021/08/12/1980-democratic-national-convention/)

[![Republican National Convention](../assets/slides/imgs/1980_rep_convention.jpg){#fig-1980-rep height=100}](https://reuther.wayne.edu/node/9207)

1980 Presidential Party National Conventions
:::

---

[![2000 Presidential Election map](../assets/slides/imgs/2000_election_map.webp){#fig-2000-election-map width=200}](https://www.google.com/url?sa=i&url=https%3A%2F%2Fwww.smithsonianmag.com%2Fhistory%2Fwhen-republicans-were-blue-and-democrats-were-red-104176297%2F&psig=AOvVaw3mlqWWKalyCl0gzVJDCIV5&ust=1694829600914000&source=images&cd=vfe&opi=89978449&ved=0CA8QjRxqFwoTCJCNg7bCq4EDFQAAAAAdAAAAABAT)

---

::: {#fig-party-brand layout-ncol=2}

![Republicans pre-2000](./assets/img/rnc_logo_pre-2000.png){#fig-republican}

![Democrats pre-2000](./assets/img/dnc_logo_1960.png){#fig-democrat}

![Republicans 2023](./assets/img/rnc_logo_2023.png){#fig-republican-post}

![Democrats 2023](./assets/img/dnc_logo_2023.png){#fig-democrat-post}

Party logos
:::

---

[![The public](../assets/slides/imgs/division.webp){#fig-public}](https://www.brookings.edu/articles/what-is-a-brokered-convention-what-is-a-contested-convention/)

## Question guiding the project

Is color a source of political information that shapes the public's behaviors and attitudes?

## De-mystifying some of the jargon I might use

- Object
  - refers to anything out in the world that we are interacting with
  - here, these objects are political
- Affect $\neq$ emotion
  - refers to the larger state of being
  - emotion refers to consciously labeled reactions
  - valance refers to pre-consciously 
- "Hot Cognition"
  - refers to the idea that some memories and experiences are more accessible or come to the top of the mind quicker.
  - this coincides with memories and experiences that are affect-laden

## Dominant models of information processing

### Zaller's [-@zaller_1992_cup] Memory model

```{mermaid}
%%| label: fig-zaller
%%| fig-cap: Memory model of info processing
graph TD;
A[Sampled pre-existing info] -->|Stronger impacts when less pre-existing knowledge| C[Updated attitudes]
B[New info] --> C
```

---

```{r}
#| label: sim-visualization

df <- fabricatr::fabricate(
  N = 10000,
  pop = rnorm(n = N, mean = 2, sd = 1),
  sample_strong = rnorm(n = N, mean = 2, sd = 0.5),
  updated_strong = rnorm(n = N, mean = 1.9, sd = 1),
  sample_weak = rnorm(n = N, mean = 2, sd = 2),
  updated_weak = rnorm(n = N, mean = 1.5, sd = 1)
)
```

```{r}
#| label: fig-pre-existing-info
#| fig-cap: Set of pre-existing attitudes
ggplot(data = df) +
  geom_density(aes(x = pop)) +
  geom_vline(aes(xintercept = 2), linetype = 1) +
  theme_minimal()
```

---

```{r}
#| label: fig-sample-pre-existing-info
#| fig-cap: Pre-existing attitudes used
#| fig-subcap:
#|   - Unknowledgeable person
#|   - Knowledgeable person
ggplot(data = df) +
  geom_density(aes(x = sample_weak)) +
  geom_vline(aes(xintercept =  2), linetype = 1) +
  theme_minimal()
ggplot(data = df) +
  geom_density(aes(x = sample_strong)) +
  geom_vline(aes(xintercept =  2), linetype = 1) +
  theme_minimal()
```

---

```{r}
#| label: fig-new-attitude
#| fig-cap: Updated set of attitudes
#| fig-subcap:
#|   - Unknowledgeable person
#|   - Knowledgeable person
ggplot(data = df) +
  geom_density(aes(x = updated_weak)) +
  geom_vline(aes(xintercept =  2), linetype = 1) +
  geom_vline(aes(xintercept =  1.5), linetype = 2) +
  theme_minimal()
ggplot(data = df) +
  geom_density(aes(x = updated_strong)) +
  geom_vline(aes(xintercept =  2), linetype = 1) +
  geom_vline(aes(xintercept =  1.9), linetype = 2) +
  theme_minimal()
```

### Lodge and Taber's [-@lodge_taber_2013_cup] Online model

```{mermaid}
%%| label: fig-online-model
%%| fig-cap: Online model of info processing
graph TD;
A[Pre-existing beliefs] --> C[Updated beliefs]
B[New information] -->|If it aligns with current beliefs| C
```

---

```{r}
#| label: sim-visualization-online

df <- fabricatr::fabricate(
  N = 10000,
  pop = rnorm(n = N, mean = 2, sd = 1),
  sample_strong = rnorm(n = N, mean = 2, sd = 0.5),
  updated_strong = rnorm(n = N, mean = 2, sd = 1),
  sample_weak = rnorm(n = N, mean = 2, sd = 2),
  updated_weak = rnorm(n = N, mean = 1.5, sd = 1)
)
```

---

```{r}
#| label: fig-pre-existing
#| fig-cap: Pre-existing beliefs
#| fig-subcap:
#|   - Weak accuracy, strong pre-existing
#|   - Strong accuracy, weak pre-existing

ggplot(data = df) +
  geom_density(aes(x = sample_strong)) +
  geom_vline(aes(xintercept =  2), linetype = 1) +
  theme_minimal()

ggplot(data = df) +
  geom_density(aes(x = sample_weak)) +
  geom_vline(aes(xintercept =  2), linetype = 1) +
  theme_minimal()
```

---

```{r}
#| label: fig-updated
#| fig-cap: Updated beliefs
#| fig-subcap:
#|   - Weak accuracy, strong pre-existing
#|   - Strong accuracy, weak pre-existing

ggplot(data = df) +
  geom_density(aes(x = updated_strong)) +
  geom_vline(aes(xintercept =  2), linetype = 1) +
  geom_vline(aes(xintercept =  2), linetype = 2) +
  theme_minimal()

ggplot(data = df) +
  geom_density(aes(x = updated_weak)) +
  geom_vline(aes(xintercept =  2), linetype = 1) +
  geom_vline(aes(xintercept =  1.5), linetype = 2) +
  theme_minimal()
```

## Amending those models with the Snap-judgment model

---

{{< include ../../assets/_theory.qmd >}}

---

```{r}
#| label: sim-snap-judgment

df <- fabricatr::fabricate(
  N = 10000,
  pop = rnorm(n = N, mean = 2, sd = 1),
  sample_strong = rnorm(n = N, mean = 2.5, sd = 0.5),
  updated_strong = rnorm(n = N, mean = 2.5, sd = 1),
  sample_weak = rnorm(n = N, mean = 2.5, sd = 2),
  updated_weak = rnorm(n = N, mean = 1.5, sd = 1)
)
```

```{r}
#| label: fig-snap-sample
#| fig-cap: Pre-existing beliefs based on visual information
#| fig-subcap:
#|   - Weak accuracy, strong pre-existing
#|   - Strong accuracy, weak pre-existing

ggplot(data = df) +
  geom_density(aes(x = sample_strong)) +
  geom_vline(aes(xintercept = 2.5), linetype = 1) +
  theme_minimal()
ggplot(data = df) +
  geom_density(aes(x = sample_weak)) +
  geom_vline(aes(xintercept = 2.5), linetype = 1) +
  theme_minimal()
```

---

```{r}
#| label: fig-snap-updated
#| fig-cap: Updated beliefs with visual information
#| fig-subcap:
#|   - Weak accuracy, strong pre-existing
#|   - Strong accuracy, weak pre-existing

ggplot(data = df) +
  geom_density(aes(x = updated_strong)) +
  geom_vline(aes(xintercept = 2.5), linetype = 1) +
  geom_vline(aes(xintercept = 2.5), linetype = 2) +
  theme_minimal()
ggplot(data = df) +
  geom_density(aes(x = updated_weak)) +
  geom_vline(aes(xintercept = 2.5), linetype = 1) +
  geom_vline(aes(xintercept = 1.5), linetype = 2) +
  theme_minimal()
```

---

- Implications
  - Refines existing info. processing models to consider pre-conscious influences.
  - Outlines the cognitive processes undergirding how we:
    - have made associations between the colors red and blue with the two parties.
    - form attitudes and behaviors based on less substantive information (e.g., policy stances).

## Study 1

### Motivation of the experiment

- Do we associate the color red with Republicans and the color blue with Democrats?
- Do these associations provide information that the public act on in politics?

### Setup of the experiment

---

#### Background information

- Age
- Education
- Gender Identity
- Sex
- Visual impairments
  - Including colorblindness
- Party Identification

---

#### Treatment

- Three trials

---

::: {#fig-trial-1}

![Red condition](./study_1/assets/rep_yard_sign_round_1.png)
![Blue condition](./study_1/assets/dem_yard_sign_round_1.png)
![White condition](./study_1/assets/cont_yard_sign_round_1.png)

Trial 1
:::

---

::: {#fig-trial-2}

![Red condition](./study_1/assets/rep_yard_sign_round_2.png)
![Blue condition](./study_1/assets/dem_yard_sign_round_2.png)
![White condition](./study_1/assets/cont_yard_sign_round_2_3.png)

Trial 2
:::

---

::: {#fig-trial-3}

![Red condition](./study_1/assets/rep_yard_sign_round_3.png)
![Blue condition](./study_1/assets/dem_yard_sign_round_3.png)
![White condition](./study_1/assets/cont_yard_sign_round_2_3.png)

Trial 3
:::

#### Design

- Recruited 1,000 participants from Prolific
- Counterbalance treatment and background information ($\frac{1}{2}$ see treatment first; other $\frac{1}{2}$ see background questions first)
- Within each trial, participants have $\frac{1}{3}$ chance of seeing any given yard sign.
- MouseView.js

---

![Example of treatment](../assets/slides/imgs/treatment.gif){#fig-treatment-example}

### Do people make these associations?

---

```{r}
#| label: fig-h-2-predictions
#| layout-nrow: 1
#| fig-cap: The partisanship of candidates change depending on the color of the yard sign
#| fig-subcap:
#|   - Red treatment
#|   - Blue treatment
red <- predicted_prob_bar(
  fitted_model = list_fitted[["h_2"]][["red_multinomial"]]
  , x_axis = "trial_1_red_stimuli"
)
blue <- predicted_prob_bar(
  fitted_model = list_fitted[["h_2"]][["blue_multinomial"]]
  , x_axis = "trial_1_blue_stimuli"
  , treatment = "Blue"
)
red
blue
```

### What does conflicting colors do to these associations?

---

```{r}
#| label: fig-h-2a-predictions
#| layout-nrow: 2
#| fig-cap: Mixing red and blue in a yard sign increases uncertainty about partisanship
#| fig-subcap:
#|   - Trial 2 Red treatment
#|   - Trial 2 Blue treatment
#|   - Trial 3 Red treatment
#|   - Trial 3 Blue treatment
plot_trial_2_red <- predicted_prob_bar(
  fitted_model = list_fitted[["h_2a"]][["trial_2_red_multinomial"]]
  , x_axis = "trial_2_red_stimuli"
)
plot_trial_2_blue <- predicted_prob_bar(
  fitted_model = list_fitted[["h_2a"]][["trial_2_blue_multinomial"]]
  , x_axis = "trial_2_blue_stimuli"
  , treatment = "Blue"
)
plot_trial_3_red <- predicted_prob_bar(
  fitted_model = list_fitted[["h_2a"]][["trial_3_red_multinomial"]]
  , x_axis = "trial_3_red_stimuli"
)
plot_trial_3_blue <- predicted_prob_bar(
  fitted_model = list_fitted[["h_2a"]][["trial_3_blue_multinomial"]]
  , x_axis = "trial_3_blue_stimuli"
  , treatment = "Blue"
)
plot_trial_2_red
plot_trial_2_blue
plot_trial_3_red
plot_trial_3_blue
```

### Do these associations inform important political behavior?

```{r}
#| label: fig-h-3-predictions
#| layout-nrow: 1
#| fig-cap: Probability of voting for candidate with yard sign
#| fig-subcap:
#|   - Red treatment
#|   - Blue treatment
red <- predicted_prob_bar(
  fitted_model = list_fitted[["h_3"]][["red"]]
  , x_axis = "trial_1_red_stimuli"
  , treatment = "Red"
  , hypothesis = "H3"
  , x_label = "Party ID"
  , y_label = "Pr(Vote for candidate)"
  , legend_title = "Treatment"
)
blue <- predicted_prob_bar(
  fitted_model = list_fitted[["h_3"]][["blue"]]
  , x_axis = "trial_1_blue_stimuli"
  , treatment = "Blue"
  , hypothesis = "H3"
  , x_label = "Party ID"
  , y_label = "Pr(Vote for candidate)"
  , legend_title = "Treatment"
)
red
blue
```

## Study 2

### Motivation of Study

---

> There obviously is a strong association with blue and the Democratic Party and red and the Republican Party. We've seen in some of our testing that specifically MAGA red tends to elicit a slightly stronger association with Republicans/conservatives/Trump than other shades. The rest of color spectrum doesn't seem to have as strong an association, though greens and purples tend to read a little more left, the combination of red, white and blue tends to read a little more right, black and white tends to read a little left, black and red tend to read a little more right, etc.

---

> A lot of decisions are still made on purely gut feelings as opposed to taking a more multi-disciplinary evidence-based approach.

---

- Do these associations have systematic influences on campaign strategy?

### Outcome variable - How much red and blue do campaigns use?

---

- Collect around 1,000 yard signs from House elections
  - 2016-2024
- Measuring the colors used on a yard sign:
  - Use a machine learning Python Library, `OpenCV`, that reads images and classifies their components.
    - Trained to classify each pixel in the image
    - Classifies it based on BGR value (reverse of RGB)
    - Looks for a range of values around the "Republican red" and "Democratic blue"

---

```{python}
#| label: example detection
#| echo: false

from sys import path # for path management
from cv2 import imread, imwrite # for color detection
    #* user-defined
#path.append("../")
from PY.helper import colorDetector

# Define colors to detect
    #* White
        #** Not defined. Default for colorDetector()
    #* Red
republican_red = [232, 27, 35] # target color
red_lower = [93, 9, 12] # lower end of spectrum for red
red_higher = [237, 69, 75] # higher end of spectrum for red

# Load gop image to read
img = imread("./assets/img/gop_2022.png")

# Detect colors
percent, img_transformed, result = colorDetector(img = img, color_upper = red_higher, color_lower = red_lower)

transformed= imwrite("../data/temp/study_2/gop_2022_transformed.png", img_transformed) # save transformed image

masked= imwrite("../data/temp/study_2/gop_2022_detected.png", result) # save masked image
```

::: {#fig-color-detection-example layout-ncol="2"}
![Resized original image](../data/temp/study_2/gop_2022_transformed.png){#fig-resized}

![Masked](../data/temp/study_2/gop_2022_detected.png){#fig-masked}

Detecting colors in the GOP logo
:::

```{python}
#| output: asis
# Print calculated percentage of example
print("For the example in @fig-color-detection-example, about {:.2f}".format(percent) + " of the image is red.")
```

### Predictor variable - What is the partisan composition of the district?

---

- MIT Election Lab data
  - House election results
  - 1970-2020
- Measuring partisan composition of district:
  - Calculate rolling 5-year average of the Democratic party's vote share for each district

### Do campaigns use a different proportion of the colors red and blue based on partisan composition?

---

```{r}
#| label: fig-study-2-pred
#| layout-ncol: 1
#| fig-cap: More blue on the yard signs in Democratic strongholds
#| fig-subcap:
#|   - Proportion of red
#|   - Proportion of blue
red <- predicted_prob_bar(
  fitted_model = list_fitted[["h_5"]][["red"]]
  , treatment = "Red"
  , hypothesis = "H5"
)
blue <- predicted_prob_bar(
  fitted_model = list_fitted[["h_5"]][["blue"]]
  , treatment = "Blue"
  , hypothesis = "H5"
)
red
blue
```

## Questions and Comments

## References

::: {#refs}
:::

## Appendix

### Study 1

---

#### Characteristics of sample

---

```{r}
#| label: tbl-study-1-descriptives
#| tbl-cap: Descriptive statistics of Prolific sample

df_demographics_subset <- list_df[["cleaned"]][
    , c(
        "Age" = "age"
        , `Color blind` = "color_blind_dummy"
        , "Female" = "female_sex"
        , "White" = "white"
        , "Hispanic" = "hispanic"
        , "Black" = "black"
        , `Party ID` = "pid_7"
    )
] |> 
as.data.frame()

datasummary_skim(
    df_demographics_subset
    , histogram = FALSE
    , notes = c(
      "Data source: Prolific."
      , "Sample characteristics."
    )

)
```

---

#### Measures

---

+-----------------------------+-----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| Measure                     | Question                                                                                                                                                                                                                                                                                                  | Coded as                                                                                                                                                           |
+=============================+:=========================================================================================================================================================================================================================================================================================================:+:==================================================================================================================================================================:+
| Age                         | Confirm your age                                                                                                                                                                                                                                                                                          | Text entry                                                                                                                                                         |
+-----------------------------+-----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| Gender                      | What is your gender identity?                                                                                                                                                                                                                                                                             | Non-binary, Female, Male, Prefer not to choose, Other                                                                                                              |
+-----------------------------+-----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| Sex                         | What is your sex assigned at birth?                                                                                                                                                                                                                                                                       | Female, Male, Other                                                                                                                                                |
+-----------------------------+-----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| Race                        | Which ethnic or racial category best describes you?                                                                                                                                                                                                                                                       | White, non-Hispanic; Black, non-Hispanic; Hispanic; Asian or Native Hawaiian/other Pacific Islander; Native American/Alaska Native or other race; Multiple races   |
+-----------------------------+-----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| Color blind                 | Have you been diagnosed with any of the following visual impairments that are not treated with corrective lenses?                                                                                                                                                                                         | Color blindness (any form); Blurred vision (not treatable by corrective lenses); Macular degeneration; Glaucoma; Cataract; Diabetic retinopathy; None of the above |
+-----------------------------+-----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| Attention                   | How often do you pay attention to what is going on in government and politics?                                                                                                                                                                                                                            | Always; Most of the time; About half the time; Some of the time; Never                                                                                             |
+-----------------------------+-----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| Party ID                    | Generally speaking, do you usually think of yourself as a Democrat, a Republican, an Independent, or what?; Would you say that you are a strong [Republican/Democrat] or not a very strong [Republican/Democrat]?; Do you think of yourself as closer to the Republican party or to the Democratic party? | Republican, Democrat, Independent, Other; Strong, Not very strong; Closer to the Republican party, Neither, Closer to Democratic                                   |
+-----------------------------+-----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| Attn. Check                 | In a few words describe what is your first memory of a political event.                                                                                                                                                                                                                                   | Text entry                                                                                                                                                         |
+-----------------------------+-----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| Trial stimuli               | Yard sign assigned                                                                                                                                                                                                                                                                                        | 0 = White, 1 = Red/Blue                                                                                                                                            |
+-----------------------------+-----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| Trial stimuli alt           | Yard sign assigned                                                                                                                                                                                                                                                                                        | 0 = Not red/blue, 1 = Red/Blue                                                                                                                                     |
+-----------------------------+-----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| Time elapsed                | NA                                                                                                                                                                                                                                                                                                        | Difference between time at start of trial and stop                                                                                                                 |
+-----------------------------+-----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| $\Delta_{t}$ (milliseconds) | NA                                                                                                                                                                                                                                                                                                        | Time elapsed between cursor movements                                                                                                                              |
+-----------------------------+-----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| $\Delta_{x}$ (pixels)       | NA                                                                                                                                                                                                                                                                                                        | Number of pixels moved along x-axis                                                                                                                                |
+-----------------------------+-----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| $\Delta_{y}$ (pixels)       | NA                                                                                                                                                                                                                                                                                                        | Number of pixels moved along y-axis                                                                                                                                |
+-----------------------------+-----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| Candidate party             | Is this candidate a Republican, a Democrat, or Neither?                                                                                                                                                                                                                                                   | Republican, Democrat, Neither                                                                                                                                      |
+-----------------------------+-----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| Candidate vote              | Of the three candidates who would you most likely vote for?                                                                                                                                                                                                                                               | The first yard sign, The second yard sign, The third yard sign                                                                                                     |
+-----------------------------+-----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| Candidate vote indicator    | NA                                                                                                                                                                                                                                                                                                        | 0 = Would not vote for first yard sign, 1 = Would vote for first yard sign                                                                                         |
+-----------------------------+-----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------+

: Study 1 Measures {#tbl-study-1-measures}

#### $H_1$

```{r}
#| label: tbl-h1-trial-1
#| tbl-cap: Average time between cursor movements (Trial 1)
df_movement_trial_1 <- list_df[["cleaned"]][
  , c(
    "trial_1_difference_Time"
    , "trial_1_stimulus"
    , "trial_1_difference_X"
    , "trial_1_difference_Y"
  )
]
setnames(
  df_movement_trial_1
  , old = c(
    "trial_1_difference_Time"
    , "trial_1_difference_X"
    , "trial_1_difference_Y"
  )
  , new = c(
    "Δ<sub>t</sub> (milliseconds)"
    , "Δ<sub>x</sub> (pixels)"
    , "Δ<sub>y</sub> (pixels)"
  )
)

datasummary_balance(
  ~trial_1_stimulus
  , data = df_movement_trial_1
  , notes = "Data source: Prolific sample."
  , escape = FALSE
)
```

---

#### $H_2$

---

##### Multinomial Logit

```{r}
#| label: tbl-hypothesis-2-main
#| tbl-cap: Full results for reported models in $H_2$
# Redefine names
list_fitted_split_1 <- list(
  "Red treatment" = list_fitted[["h_2"]][["red_multinomial"]]
  , "Blue treatment" = list_fitted[["h_2"]][["blue_multinomial"]]
)

# Change goodness of fit stuff
gm <- list(
  list("raw" = "nobs", "clean" = "Num.Obs.", "fmt" = 1)
  , list("raw" = "r.squared", "clean" = "R<sup>2</sup>", "fmt" = 2)
  , list("raw" = "waic", "clean" = "WAIC", "fmt" = 2)
)
# Make table
modelsummary(
  list_fitted_split_1
  , coef_map = c(
    "b_mu2_trial_1_red_stimuli" = "Red yard sign (Independent)"
    , "b_mu3_trial_1_red_stimuli" = "Red yard sign (Republican)"
    , "b_mu2_trial_1_blue_stimuli" = "Blue yard sign (Independent)"
    , "b_mu3_trial_1_blue_stimuli" = "Blue yard sign (Republican)"
    , "b_mu2_Intercept" = "Intercept - Independent"
    , "b_mu3_Intercept" = "Intercept - Republican"
    )
  , statistic = "conf.int"
  , conf_level = 0.9
  , gof_map = gm 
  , notes = c(
    "Data source: Pavlovia"
    , "Positive coefficients reflect the logged-odds"
    , "of claiming candidate is not a Democrat."
    , "Numbers inside brackets reflect the 5th and 95th percentile"
    , "of the draws from the posterior distribution."
  )
  , escape = FALSE

)
```

---

```{r}
#| label: fig-h-2-ppc-multinomial
#| fig-cap: Posterior predictive check of Multinomial models for $H_2$
#| fig-subcap:
#|   - Red treatment
#|   - Blue treatment

list_ppc[["h_2"]][["red_multinomial"]]
list_ppc[["h_2"]][["blue_multinomial"]]
```

##### Ordered Logit

```{r}
#| label: tbl-hypothesis-2-ordered
#| tbl-cap: Full results for reported ordered logistic regression models for $H_2$
# Redefine names
list_fitted_split_1_alt <- list(
  "Red treatment" = list_fitted[["h_2"]][["red_ordinal"]]
  , "Blue treatment" = list_fitted[["h_2"]][["blue_ordinal"]]
)

# Change goodness of fit stuff
gm <- list(
  list("raw" = "nobs", "clean" = "Num.Obs.", "fmt" = 1)
  , list("raw" = "r.squared", "clean" = "R<sup>2</sup>", "fmt" = 2)
  , list("raw" = "waic", "clean" = "WAIC", "fmt" = 2)
)
# Make table
modelsummary(
  list_fitted_split_1_alt
  , coef_map = c(
    "b_trial_1_red_stimuli" = "Red yard sign"
    , "b_trial_1_blue_stimuli" = "Blue yard sign"
    , "b_Intercept[1]" = "Intercept 1"
    , "b_Intercept[2]" = "Intercept 2"
    )
  , statistic = "conf.int"
  , conf_level = 0.9
  , gof_map = gm 
  , notes = c(
    "Data source: Pavlovia"
    , "Positive coefficients reflect the logged-odds"
    , "of claiming candidate is Republican."
    , "Numbers inside brackets reflect the 5th and 95th percentile"
    , "of the draws from the posterior distribution."
  )
  , escape = FALSE

)
```

---

```{r}
#| label: fig-h-2-ppc-ordered
#| fig-cap: Posterior predictive checks of Ordered Logit models in $H_2$
#| fig-subcap:
#|   - Red treatment
#|   - Blue treatment

list_ppc[["h_2"]][["red_ordinal"]]
list_ppc[["h_2"]][["blue_ordinal"]]
```

##### Single model

```{r}
#| label: tbl-robustness-alt-specification-h-2
#| tbl-cap: $H_2$ models with alternative specification
# Redefine names
list_fitted_split_3 <- list(
  "Single model" = list_fitted[["h_2"]][["single_model"]]
)

# Make table
modelsummary(
  list_fitted_split_3
  , coef_map = c(
    "b_trial_1_red_stimuli_alt" = "Red yard sign"
    , "b_trial_1_blue_stimuli_alt" = "Blue yard sign"
    , "b_Intercept[1]" = "Intercept 1"
    , "b_Intercept[2]" = "Intercept 2"
  )
  , statistic = "conf.int"
  , conf_level = 0.9
  , gof_map = gm
  , notes = c(
    "Data source: Pavlovia"
    , "Positive coefficients reflect the logged-odds"
    , " of claiming candidate is Republican."
    , "Numbers inside brackets reflect the 5th and 95th percentile"
    , " of the draws from the posterior distribution."
  )
  , escape = FALSE

)
```

---

```{r}
#| label: fig-h-2-ppc-single
#| fig-cap: Posterior predictive checks of single model specification for $H_2$

list_ppc[["h_2"]][["single_model"]]

```

---

##### Moderated by age

```{r}
#| label: tbl-interacted-models-h-2
#| tbl-cap: $H_2$ models moderated by age
# Redefine names
list_fitted_split_2 <- list(
  "Red treatment - Age moderating" = list_fitted[["h_2"]][["red_interact"]]
  , "Blue treatment - Age moderating" = list_fitted[["h_2"]][["blue_interact"]]
)

# Make table
modelsummary(
  list_fitted_split_2
  , coef_map = c(
    "b_trial_1_red_stimuli" = "Red yard sign"
    , "b_trial_1_blue_stimuli" = "Blue yard sign"
    , "b_age" = "Age"
    , "b_trial_1_red_stimuli:age" = "Red yard sign × Age"
    , "b_trial_1_blue_stimuli:age" = "Blue yard sign × Age"
    , "b_Intercept[1]" = "Intercept 1"
    , "b_Intercept[2]" = "Intercept 2"
    )
  , statistic = "conf.int"
  , conf_level = 0.9
  , gof_map = gm 
  , notes = c(
    "Data source: Pavlovia"
    , "Positive coefficients reflect the logged-odds"
    , " of claiming candidate is Republican."
    , "Numbers inside brackets reflect the 5th and 95th percentile"
    , " of the draws from the posterior distribution."
  )
  , escape = FALSE

)
```

---

```{r}
#| label: fig-h-2-ppc-age
#| fig-cap: Posterior predictive checks of models in $H_2$ moderated by age
#| fig-subcap:
#|   - Red treatment moderated by age
#|   - Blue treatment moderated by age

list_ppc[["h_2"]][["red_interact"]]
list_ppc[["h_2"]][["blue_interact"]]
```

---

##### Moderated by colorblindness

```{r}
#| label: tbl-colorblindness
#| tbl-cap: $H_2$ moderated by colorblindness

# Redefine names
list_fitted_split_4 <- list(
  "Red - Color blindness moderating" = list_fitted[["h_2"]][["red_colorblind"]]
  , "Blue - Color blindness moderating" = list_fitted[["h_2"]][["blue_colorblind"]]
)
# Make table
modelsummary(
  list_fitted_split_4
  , coef_map = c(
    "b_trial_1_red_stimuli" = "Red yard sign"
    , "b_trial_1_blue_stimuli" = "Blue yard sign"
    , "b_trial_1_red_stimuli:color_blind_dummy" = "Red yard sign × Color blind"
    , "b_trial_1_blue_stimuli:color_blind_dummy" = "Blue yard sign × Color blind"
    , "b_Intercept[1]" = "Intercept 1"
    , "b_Intercept[2]" = "Intercept 2"
  )
  , statistic = "conf.int"
  , conf_level = 0.9
  , gof_map = gm 
  , notes = c(
    "Data source: Pavlovia"
    , "Positive coefficients reflect the logged-odds"
    , " of claiming candidate is Republican."
    , "Numbers inside brackets reflect the 5th and 95th percentile"
    , " of the draws from the posterior distribution."
  )
  , escape = FALSE

)
```

---

```{r}
#| label: fig-h-2-ppc-colorblind
#| fig-cap: Posterior predictive checks of models in $H_2$ moderated by colorblindness
#| fig-subcap:
#|   - Red treatment moderated by colorblindness
#|   - Blue treatment moderated by colorblindness

list_ppc[["h_2"]][["red_colorblind"]]
list_ppc[["h_2"]][["blue_colorblind"]]
```

---

#### $H_{2a}$

---

##### Multinomial

```{r}
#| label: tbl-hypothesis-2a
#| tbl-cap: Full results for models testing bounds of $H_2$
# Redefine names
model_names <- list(
  "Trial 2 - Red treatment" = list_fitted[["h_2a"]][["trial_2_red_multinomial"]]
  , "Trial 2 - Blue treatment" = list_fitted[["h_2a"]][["trial_2_blue_multinomial"]]
  , "Trial 2 - Single model" = list_fitted[["h_2a"]][["trial_2_single_model"]]
)

# Change goodness of fit stuff
gm <- list(
  list("raw" = "nobs", "clean" = "Num.Obs.", "fmt" = 1)
  , list("raw" = "r.squared", "clean" = "R<sup>2</sup>", "fmt" = 2)
  , list("raw" = "waic", "clean" = "WAIC", "fmt" = 2)
)
# Make table
modelsummary(
  model_names
  , coef_map = c(
    "b_mu2_trial_2_red_stimuli" = "Red yard sign (Independent)"
    , "b_mu3_trial_2_red_stimuli" = "Red yard sign (Republican)"
    , "b_mu2_trial_2_blue_stimuli" = "Blue yard sign (Independent)"
    , "b_mu3_trial_2_blue_stimuli" = "Blue yard sign (Republican)"
    , "b_trial_2_red_stimuli_alt" = "Red yard sign"
    , "b_trial_2_blue_stimuli_alt" = "Blue yard sign"
    , "b_mu2_Intercept" = "Intercept - Independent"
    , "b_mu3_Intercept" = "Intercept - Republican"
    , "b_Intercept[1]" = "Intercept 1"
    , "b_Intercept[2]" = "Intercept 2"
    )
  , statistic = "conf.int"
  , conf_level = 0.9
  , gof_map = gm
  , notes = c(
    "Data source: Pavlovia"
    , "Positive coefficients reflect the logged-odds"
    , " of claiming candidate is not a Democrat (Columns 1 and 2) or is a Republican (Column 3)."
    , "Numbers inside brackets reflect the 5th and 95th percentile"
    , " of the draws from the posterior distribution."
  )
  , escape = FALSE

)
```

---

```{r}
#| label: fig-h-2a-multinomial-trial-2-ppc
#| fig-cap: Posterior Predictive Checks for Multinomial models testing $H_{2a}$, Trial 2
#| fig-subcap:
#|   - Red treatment
#|   - Blue treatment
#|   - Single model

list_ppc[["h_2a"]][["trial_2_red_multinomial"]]
list_ppc[["h_2a"]][["trial_2_blue_multinomial"]]
list_ppc[["h_2a"]][["trial_2_single_model"]]
```

---

```{r}
#| label: tbl-hypothesis-2a-cont
#| tbl-cap: Full results for models testing bounds of $H_2$ (cont.)
# Redefine names
model_names <- list(
  "Trial 3 - Red treatment" = list_fitted[["h_2a"]][["trial_3_red_multinomial"]]
  , "Trial 3 - Blue treatment" = list_fitted[["h_2a"]][["trial_3_blue_multinomial"]]
  , "Trial 3 - Single model" = list_fitted[["h_2a"]][["trial_3_single_model"]]
)

# Change goodness of fit stuff
gm <- list(
  list("raw" = "nobs", "clean" = "Num.Obs.", "fmt" = 1)
  , list("raw" = "r.squared", "clean" = "R<sup>2</sup>", "fmt" = 2)
  , list("raw" = "waic", "clean" = "WAIC", "fmt" = 2)
)
# Make table
modelsummary(
  model_names
  , coef_map = c(
    "b_mu2_trial_3_red_stimuli" = "Red yard sign (Independent)"
    , "b_mu3_trial_3_red_stimuli" = "Red yard sign (Republican)"
    , "b_mu2_trial_3_blue_stimuli" = "Blue yard sign (Independent)"
    , "b_mu3_trial_3_blue_stimuli" = "Blue yard sign (Republican)"
    , "b_trial_3_red_stimuli_alt" = "Red yard sign"
    , "b_trial_3_blue_stimuli_alt" = "Blue yard sign"
    , "b_mu2_Intercept" = "Intercept - Independent"
    , "b_mu3_Intercept" = "Intercept - Republican"
    , "b_Intercept[1]" = "Intercept 1"
    , "b_Intercept[2]" = "Intercept 2"
    )
  , statistic = "conf.int"
  , conf_level = 0.9
  , gof_map = gm 
  , notes = c(
    "Data source: Pavlovia"
    , "Positive coefficients reflect the logged-odds"
    , " of claiming candidate is Republican."
    , "Numbers inside brackets reflect the 5th and 95th percentile"
    , " of the draws from the posterior distribution."
  )
  , escape = FALSE

)
```

---

```{r}
#| label: fig-h-2a-multinomial-trial-3-ppc
#| fig-cap: Posterior Predictive Checks for Multinomial models testing $H_{2a}$, Trial 3
#| fig-subcap:
#|   - Red treatment
#|   - Blue treatment
#|   - Single model

list_ppc[["h_2a"]][["trial_3_red_multinomial"]]
list_ppc[["h_2a"]][["trial_3_blue_multinomial"]]
list_ppc[["h_2a"]][["trial_3_single_model"]]
```

---

##### Ordinal model

```{r}
#| label: tbl-hypothesis-2a-ordered
#| tbl-cap: Full results for ordered logistic regressions testing bounds of $H_2$
# Redefine names
model_names <- list(
  "Trial 2 - Red treatment" = list_fitted[["h_2a"]][["trial_2_red_ordinal"]]
  , "Trial 2 - Blue treatment" = list_fitted[["h_2a"]][["trial_2_blue_ordinal"]]
)

# Change goodness of fit stuff
gm <- list(
  list("raw" = "nobs", "clean" = "Num.Obs.", "fmt" = 1)
  , list("raw" = "r.squared", "clean" = "R<sup>2</sup>", "fmt" = 2)
  , list("raw" = "waic", "clean" = "WAIC", "fmt" = 2)
)
# Make table
modelsummary(
  model_names
  , coef_map = c(
    "b_trial_2_red_stimuli" = "Red yard sign"
    , "b_trial_2_blue_stimuli" = "Blue yard sign"
    , "b_Intercept[1]" = "Intercept 1"
    , "b_Intercept[2]" = "Intercept 2"
    )
  , statistic = "conf.int"
  , conf_level = 0.9
  , gof_map = gm
  , notes = c(
    "Data source: Pavlovia"
    , "Positive coefficients reflect the logged-odds"
    , " of claiming candidate is Republican."
    , "Numbers inside brackets reflect the 5th and 95th percentile"
    , " of the draws from the posterior distribution."
  )
  , escape = FALSE

)
```

---

```{r}
#| label: fig-h-2a-ordinal-trial-2-ppc
#| fig-cap: Posterior Predictive Checks for Ordinal models testing $H_{2a}$, Trial 2
#| fig-subcap:
#|   - Red treatment
#|   - Blue treatment

list_ppc[["h_2a"]][["trial_2_red_ordinal"]]
list_ppc[["h_2a"]][["trial_2_blue_ordinal"]]
```
---

```{r}
#| label: tbl-hypothesis-2a-cont-ordered
#| tbl-cap: Full results for models testing bounds of $H_2$ (cont.)
# Redefine names
model_names <- list(
  "Trial 3 - Red treatment" = list_fitted[["h_2a"]][["trial_3_red_ordinal"]]
  , "Trial 3 - Blue treatment" = list_fitted[["h_2a"]][["trial_3_blue_ordinal"]]
)

# Change goodness of fit stuff
gm <- list(
  list("raw" = "nobs", "clean" = "Num.Obs.", "fmt" = 1)
  , list("raw" = "r.squared", "clean" = "R<sup>2</sup>", "fmt" = 2)
  , list("raw" = "waic", "clean" = "WAIC", "fmt" = 2)
)
# Make table
modelsummary(
  model_names
  , coef_map = c(
    "b_trial_3_red_stimuli" = "Red yard sign"
    , "b_trial_3_blue_stimuli" = "Blue yard sign"
    , "b_Intercept[1]" = "Intercept 1"
    , "b_Intercept[2]" = "Intercept 2"
    )
  , statistic = "conf.int"
  , conf_level = 0.9
  , gof_map = gm 
  , notes = c(
    "Data source: Pavlovia"
    , "Positive coefficients reflect the logged-odds"
    , " of claiming candidate is Republican."
    , "Numbers inside brackets reflect the 5th and 95th percentile"
    , " of the draws from the posterior distribution."
  )
  , escape = FALSE

)
```

---
```{r}
#| label: fig-h-2a-ordinal-trial-3-ppc
#| fig-cap: Posterior Predictive Checks for Ordinal models testing $H_{2a}$, Trial 3
#| fig-subcap:
#|   - Red treatment
#|   - Blue treatment

list_ppc[["h_2a"]][["trial_3_red_ordinal"]]
list_ppc[["h_2a"]][["trial_3_blue_ordinal"]]
```

---

#### $H_3$

##### Main results

```{r}
#| label: tbl-h-3-main
#| tbl-cap: Support of candidate based on partisanship

# Redefine names
model_names <- list(
  "Red stimuli" = list_fitted[["h_3"]][["red"]]
  , "Blue stimuli" = list_fitted[["h_3"]][["blue"]]
)

# Make table
modelsummary(
  model_names
  , coef_map = c(
    "b_trial_1_red_stimuli" = "Red yard sign"
    , "b_trial_1_blue_stimuli" = "Blue yard sign"
    , "b_pid_7" = "Party ID"
    , "b_trial_1_red_stimuli:pid_7" = "Red yard sign × Party ID"
    , "b_trial_1_blue_stimuli:pid_7" = "Blue yard sign × Party ID"
    , "b_attention" = "Attention"
    , "b_age" = "Age"
    , "b_gender_idMale" = "Gender - Male"
    , "b_gender_idNonMbinary" = "Gender - Non-binary"
    , "b_gender_idPrefernottochoose" = "Gender - Prefer not to choose"
    , "b_race_idBlacknonMhispanic" = "Black"
    , "b_race_idHispanice" = "Hispanic"
    , "b_race_idMultipleraces" = "Multiple races"
    , "b_race_idNativeAmericanDAlaskaNativeOrOtherRace" = "Other race"
    , "b_race_idWhitenonMhispanic" = "White"
  )  
  , statistic = "conf.int"
  , conf_level = 0.9
  , gof_map = gm 
  , notes = c(
    "Data source: Pavlovia"
    , "Positive coefficients reflect the logged-odds"
    , " of claiming candidate is Republican."
    , "Numbers inside brackets reflect the 5th and 95th percentile"
    , " of the draws from the posterior distribution."
  )
  , escape = FALSE

)
```

---

```{r}
#| label: fig-h-3-ppc
#| fig-cap: Posterior Predictive checks for $H_3$ models
#| fig-subcap:
#|   - Red treatment
#|   - Blue treatment
list_ppc[["h_3"]][["red"]]
list_ppc[["h_3"]][["blue"]]
```

---

#### $H_4$

---

##### ✅ for over-dispersion

```{r}
#| label: tbl-h-4-dispersion
#| tbl-cap: Checking for over-dispersion of data
df_over_dispersed <- list_df[["cleaned"]][, c("trial_1_time_elapsed")]
names(df_over_dispersed)[names(df_over_dispersed) == 'trial_1_time_elapsed'] <- 'Time Elapsed'
datasummary_skim(
  df_over_dispersed
  , notes = "Data source: Prolific sample."
  , histogram = FALSE
)
```

##### Main results

```{r}
#| label: tbl-h-4
#| tbl-cap: The effect of partisan congruency on time spent looking at a yard sign
# Redefine names
model_names <- list(
  "Red stimuli" = list_fitted[["h_4"]][["red"]]
  , "Blue stimuli" = list_fitted[["h_4"]][["blue"]]
)

# Make table
modelsummary(
  model_names
  , coef_map = c(
    "b_trial_1_red_stimuli" = "Red yard sign"
    , "b_trial_1_blue_stimuli" = "Blue yard sign"
    , "b_pid_7" = "Party ID"
    , "b_trial_1_red_stimuli:pid_7" = "Red yard sign × Party ID"
    , "b_trial_1_blue_stimuli:pid_7" = "Blue yard sign × Party ID"
    , "b_attention" = "Attention"
    , "b_age" = "Age"
    , "b_gender_idMale" = "Gender - Male"
    , "b_gender_idNonMbinary" = "Gender - Non-binary"
    , "b_gender_idPrefernottochoose" = "Gender - Prefer not to choose"
    , "b_race_idBlacknonMhispanic" = "Black"
    , "b_race_idHispanice" = "Hispanic"
    , "b_race_idMultipleraces" = "Multiple races"
    , "b_race_idNativeAmericanDAlaskaNativeOrOtherRace" = "Other race"
    , "b_race_idWhitenonMhispanic" = "White"
  )  
  , statistic = "conf.int"
  , conf_level = 0.9
  , gof_map = gm 
  , notes = c(
    "Data source: Pavlovia"
    , "Positive coefficients reflect the logged-odds"
    , " of claiming candidate is Republican."
    , "Numbers inside brackets reflect the 5th and 95th percentile"
    , " of the draws from the posterior distribution."
  )
  , escape = FALSE

)
```

---

```{r}
#| label: fig-h-4-ppc
#| fig-cap: Posterior predictive checks of $H_4$ models
#| fig-subcap:
#|   - Red
#|   - Blue
list_ppc[["h_4"]][["red"]]
list_ppc[["h_4"]][["blue"]]
```

#### $H_5$

##### Main results

```{r}
#| label: tbl-hypothesis-5
#| tbl-cap: The effect of democratic vote share on color choices for yard sign

names(list_fitted[["h_5"]]) <- c("Prop. yard sign Red", "Prop. yard sign Blue")
modelsummary(
  list_fitted[["h_5"]]
  , stat = "conf.int"
  , conf_level = 0.9
  , metrics = c("n.obs", "RMSE")
  , coef_map = c(
    "b_dem_vote_share" = "Democratic vote share (percent)"
    , "b_Intercept" = "Intercept"
  )
  , notes = c(
    "Data source: Center for Political Design and MIT Election Lab"
    , "Positive coefficients reflect the logged-odds"
    , "of claiming candidate is Republican."
    , "Numbers inside brackets reflect the 5th and 95th percentile"
    , "of the draws from the posterior distribution."
  )
  , escape = FALSE

)
```

---

```{r}
#| label: fig-hypothesis-5-ppc
#| layout-ncol: 2
#| fig-cap: Posterior Predictive checks for $H_5$
#| fig-subcap:
#|   - Proportion of yard signs red
#|   - Proportion of yard signs blue

list_ppc[["h_5"]][["red"]]$continuous
list_ppc[["h_5"]][["blue"]]$continuous
```

## Copyright Information

<span class="faux-block">{{< fa brands creative-commons >}} 2021-2023 Damon Charles Roberts</span>
<span class="faux-block"><br>All content licensed under {{< fa brands creative-commons >}} {{< fa brands creative-commons-by >}} {{< fa brands creative-commons-sa >}} [Creative Commons CC BY-SA 4.0](https://creativecommons.org/licenses/by-sa/4.0/)</span>