### 2023-09-19 grants/ddirg/src/roberts_biographical_sketch.qmd
adding drafts of grant applications for DDIRG

---
subtitle: Biographical sketch
format:
  fancy-doc-pdf:
    toc: false
    latex-output-dir: ../out/
    biblatex: true
    filters:
      #- "_extensions/andrewheiss/wordcount/citeproc.lua"
      - "_extensions/andrewheiss/wordcount/wordcount.lua"
author:
  - name: Damon C. Roberts
    affiliations:
      - name: Univeristy of Colorado Boulder
    email: damon.roberts-1@colorado.edu
bibliography: ../../../assets/references.bib
---

I have a track record of successfully earning and using awarded research grants for my projects. I have earned grants from The Time-Sharing Experiments for the Social Science Short Studies Program (TESS), grants from my department, and a grant from the Elections, Voting, Public Opinion, and Behavior (EVOPB) Section of the American Political Science Association. I have published multiple peer-reviewed papers from these grants and have presented over 14 times at academic conferences.

Since the fall of 2019, I have published five peer-reviewed academic papers; on four of these I have been the first author. My research has appeared in Political Behavior, Political Research Quarterly, Social Science Quarterly, and twice in American Politics Research. I have several other papers at various stages of the publishing process. I am committed to expanding the reach of my work beyond academia, and have also engaged in a significant amount of public scholarship in the past year. I have co-authored a public piece discussing my research in the [*Washington Post's Monkey Cage Blog*](https://www.washingtonpost.com/politics/2022/07/21/biden-trump-president-age-2024-election/) and have agreed to a number of interviews about my research -- this includes a recorded interview with Colorado Public Radio and an interview written up in the New York Times.

In conjunction with my research, I have also been the instructor of record for two under-graduate statistics courses focused on introductory material for quantitative political science, as well as more advanced discussions of research design and causal inference. I have been a teaching assistant tasked with grading and leading discussion sections in introductory American Politics, formal modeling, campaigns and elections, and introductory research methodology.

All of these experiences have contributed to my development as a producer of research, but also as an academic who both values and has experience communicating my work to a wide array of audiences within and beyond academia. [An itemized CV is on my website](https://damoncroberts.com/cv/dcr_cv.pdf)
<!--
I have a track record of successfully earning and using awarded research grants for my projects. I have earned grants from The Time-Sharing Experiments for the Social Science Short Studies Program (TESS), research grants from my department five times, and a grant from the Elections, Voting, Public Opinion, and Behavior (EVOPB) Section of the American Political Science Association. I have published multiple peer-reviewed papers from these grants and presented over 14 times at an academic conference.

Since the fall of 2019, I have published five peer-reviewed academic papers, four of which I have been the first author. My research has appeared in Political Behavior, Political Research Quarterly, Social Science Quarterly, and twice in American Politics Research. I have several other papers at various stages of the publishing process. I am committed to expanding the reach of my work beyond academia and have also engaged in a significant amount of public scholarship in the past year. I have co-authored a public piece discussing my research in the [*Washington Post's Monkey Cage Blog*](https://www.washingtonpost.com/politics/2022/07/21/biden-trump-president-age-2024-election/) and have agreed to a number of interviews about my research which includes a recorded interview with Colorado Public Radio and an interview written up in the New York Times.

In conjunction with my research, I have also been the instructor of record for two undergraduate statistics courses focused on introductory material for quantitative political science and more advanced discussions of research design and causal inference. I have been a teaching assistant tasked with grading and leading discussion sections in introductory American Politics, formal modeling, campaigns and elections, and introductory research methodology.

All of these experiences have contributed to my development as a producer of research but also as an academic who has experience completing and publishing research and communicating my work to a wide array of audiences within and beyond academia. [An itemized CV is on my website.](https://damoncroberts.com/cv/dcr_cv.pdf)
-->