### 2023-09-19 grants/ddirg/src/roberts_proposal_summary.qmd

---
subtitle: Proposal Summary
format:
  fancy-doc-pdf:
    toc: false
    latex-output-dir: ../out/
    biblatex: true
    filters:
      #- "_extensions/andrewheiss/wordcount/citeproc.lua"
      - "_extensions/andrewheiss/wordcount/wordcount.lua"
author:
  - name: Damon C. Roberts
    affiliations:
      - name: Univeristy of Colorado Boulder
    email: damon.roberts-1@colorado.edu
bibliography: ../../../assets/references.bib
---

The use of the colors red and blue to distinguish the two dominant political parties in the United States started during the 2000 Presidential Election with the electoral maps [@elving_2014_npr]. Since then, the two political party's logos have changed rather dramatically in that now the Republican party uses the color red, and the Democratic party only uses the color blue. @fig-party-brand nicely illustrates this point. Some existing descriptive research demonstrates that these clear distinctions in color are not limited to party logos but that Democrats use more blue than Republicans. In contrast, Republicans use more red on their campaign's yard signs [@williams_et-al_2022_jomp].^[The first empirical chapter of this dissertation replicates this finding when exploring the rates of color usage among partisans in districts that have more or less electoral competition.] There are two primary questions for this dissertation and book project. The first question seeks to answer whether the colors red and blue carry affiliative information for voters to use as a heuristic. The second question asks whether this affiliative information that these colors may provide influence attitude expression and behavior, even in the face of more traditional substantive political information.

::: {#fig-party-brand layout-ncol=2}

![Republicans pre-2000](../../../chapter_2/src/assets/img/rnc_logo_pre-2000.png){#fig-republican}

![Democrats pre-2000](../../../chapter_2/src/assets/img/dnc_logo_1960.png){#fig-democrat}

![Republicans 2023](../../../chapter_2/src/assets/img/rnc_logo_2023.png){#fig-republican-post}

![Democrats 2023](../../../chapter_2/src/assets/img/dnc_logo_2023.png){#fig-democrat-post}

Party logos
:::

The primary contribution of this project is the cognitive model of visual information that it provides. Existing work on the visual aspects of politics is a newer area of work and comes from several different academic disciplines. This leaves us with a somewhat fragmented theory of how visual information affects mass attitudes and behavior. This project lays out a model that not only illustrates how visual information like color may be a source of information in politics; but as visual information is often processed faster than more symbolic information such as substantive political information, it also illustrates that the first-order information explaining an individual's attitude expression is visual information, such as color. The empirical chapters of the project discuss how this influences our understanding of vote choice and the value of yard signs to campaigns, discuss how this informs the degree to which people engage in conversations about politics, provides a typology of when persuasion can occur, and explains how we can make guesses about the partisanship of a group of people based on the way they have curated the environment in which they reside.

This grant would be a significant help with the costs of data collection. While I have performed the data collection for the first of the three primary empirical chapters, the last two require several experiments that require both online surveys and in-person experiments. Therefore, financial assistance with recruiting sufficiently powered and representative samples, along with adequately paying these participants for participating in these studies, would be a tremendous help for the completion of this dissertation and will remove the need for me to redo a number of these studies when working on the book project after defending the dissertation.