### 2023-09-22 chapter_2/src/prr/si_color_influence_pre-reg_anon.qmd
init gh-pages branch

---
title: |
  Supplementary information to a Pre-analysis plan
subtitle: |
  How do colors convey political information and affect individual attitudes?
#code-repo: |
#  For replication, go to: <https://github.com/DamonCharlesRoberts/dissertation>.
author:
  - name: Anonymized for peer review
  #- name: Damon C. Roberts
  #  email: damon.roberts-1@colorado.edu
  #  orcid: 0000-0002-4360-3675
  #  title: PhD Candidate
  #  affiliations:
  #    - id: CU
  #      name: University of Colorado Boulder
  #      department: Political Science
  #      address: 333 UCB
  #      city: Boulder
  #      region: CO 
  #      postal-code: 80309-0333
abstract: |
  Colors are important to politics as a form of political information. Building upon existing theories of political information processing, attitude formation, and affect in neuroscience, I present a snap-judgement model of political information processing. In this model, colors provide automatic information about a politically-relevant object that may shape subsequent processing of more complex information. The model has important implications for how we consider the role that visual information has on political information processing and attitude formation. The model additionally provides clarity on motivations behind party branding and the ways in which information may activate partisan biases pre-consciously. I test this model using a survey experiment tracking participants' mouse movements to view different parts of yard signs that vary the use of partisan and non-partisan colors. I additionally leverage congressional redistricting to examine changes in color use for political yard signs as a strategic choice for attracting voters by communicating partisanship.
bibliography: "../../../assets/references.bib"
format:
  hikmah-pdf:
    # Spit out in drafts directory
    latex-output-dir: "../../out/prr"
    # Use biblatex-chicago
    biblatex-chicago: true
    biblio-style: authordate
    biblatexoptions: 
      - backend=biber
      - autolang=hyphen
      - isbn=false
      - uniquename=false
code-overflow: wrap
nocite: |
  @quarto
execute:
  echo: false
  warning: false
  message: false
  cache: false
---

# Pre-test

## Stimuli

![Blue](../../data/prr/pre-test/blue_treatment.jpg)

![Red](../../data/prr/pre-test/red_treatment.jpg)

![White](../../data/prr/pre-test/white_treatment.jpg)

{{< pagebreak >}}

## Measures

+-----------------+--------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------+-------------------------------------------------+
| Measure         | Question                                                                                                                                   | Response options                                 | Coded as                                        |
+================:+:==========================================================================================================================================:+:================================================:+:===============================================:+
| PartyId         | -   pid = Generally speaking, do you usually think of yourself as a Republican, a Democrat, an Independent or what?                        | -   1 = Republican                               | -   -3 = Strong democrat                        |
|                 |                                                                                                                                            |                                                  |                                                 |
|                 | -   dem1\[rep1\] (if pid == 2\[1\]) = Would you call yourself a strong Democrat\[Republican\] or a not very strong Democrat\[Republican\]? | -   2 = Democrat                                 |     -   pid == 2 & dem1 == 3                    |
|                 |                                                                                                                                            |                                                  |                                                 |
|                 | -   ind1 (if pid == 3 \| 4) = Do you think of yourself as closer to the Republican or Democratic party?                                    | -   3 = Independent                              | -   -2 = Democrat                               |
|                 |                                                                                                                                            |                                                  |                                                 |
|                 |                                                                                                                                            | -   4 = Other                                    |     -   pid == 2 & dem1 == 2                    |
|                 |                                                                                                                                            |                                                  |                                                 |
|                 |                                                                                                                                            | -   1 = Strong Democrat\[Republican\]            | -   -1 = Leans Democratic                       |
|                 |                                                                                                                                            |                                                  |                                                 |
|                 |                                                                                                                                            | -   2 = Not a very strong Democrat\[Republican\] |     -   pid == 3 & ind1 == 2                    |
|                 |                                                                                                                                            |                                                  |                                                 |
|                 |                                                                                                                                            | -   1 = Republican party                         | -   0 = Independent                             |
|                 |                                                                                                                                            |                                                  |                                                 |
|                 |                                                                                                                                            | -   2 = Democratic party                         |     -   pid == 3 & ind1 == 3                    |
|                 |                                                                                                                                            |                                                  |                                                 |
|                 |                                                                                                                                            | -   3 = Neither party                            | -   1 = Leans Republican                        |
|                 |                                                                                                                                            |                                                  |                                                 |
|                 |                                                                                                                                            |                                                  |     -   pid == 3 & ind1 == 1                    |
|                 |                                                                                                                                            |                                                  |                                                 |
|                 |                                                                                                                                            |                                                  | -   2 = Republican                              |
|                 |                                                                                                                                            |                                                  |                                                 |
|                 |                                                                                                                                            |                                                  |     -   pid == 1 & rep == 2                     |
|                 |                                                                                                                                            |                                                  |                                                 |
|                 |                                                                                                                                            |                                                  | -   3 = Strong Republican                       |
|                 |                                                                                                                                            |                                                  |                                                 |
|                 |                                                                                                                                            |                                                  |     -   pid == 1 & rep == 1                     |
+-----------------+--------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------+-------------------------------------------------+
| Party           | -   dr_pid = If you had to guess, is the politician who owns this yard sign a Republican, a Democrat or neither?                           | -   1 = Republican                               | -   1 = Democrat                                |
|                 |                                                                                                                                            |                                                  |                                                 |
|                 |                                                                                                                                            | -   2 = Democrat                                 | -   2 = Neither - An independent or Third party |
|                 |                                                                                                                                            |                                                  |                                                 |
|                 |                                                                                                                                            | -   3 = Neither - An independent or Third party  | -   3 = Republican                              |
+-----------------+--------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------+-------------------------------------------------+
| Vote            | -   dr_info_4 = When you see this yard sign do you want to: Vote for this candidate?                                                       | -   1 = Yes                                      | -   1 = dr_info_4 == No & dr_info_5 == Yes      |
|                 |                                                                                                                                            |                                                  |                                                 |
|                 | -   dr_info_5 = When you see this yard sign do you want to: Avoid this candidate?                                                          | -   2 = No                                       | -   2 = dr_info_4 == Yes & dr_info_5 == No      |
|                 |                                                                                                                                            |                                                  |                                                 |
|                 |                                                                                                                                            |                                                  | -   2 = dr_info_4 == No & dr_info_5 == No       |
|                 |                                                                                                                                            |                                                  |                                                 |
|                 |                                                                                                                                            |                                                  | -   3 = dr_info_4 == Yes & dr_info_5 == No      |
+-----------------+--------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------+-------------------------------------------------+

: Pre-test measures

## Results

```{r}
#| label: setup-block

box::use(
    modelsummary=modelsummary[modelsummary],
    tibble=tibble[tribble]
)
```

```{r}
#| label: load models

load(
    file="../../data/models/prr/pre-test/pre-test_party_model.RDS"
  )
load(
    file="../../data/models/prr/pre-test/pre-test_vote_model.RDS"
  )
```

```{r}
#| label: tbl-pre-test-table
#| tbl-cap: Color on Yard Signs shape perceptions and vote intentions of candidate
# Table for models
  #* Set up non-coefficient related information from models 
gm = list(
    list('raw' = 'nobs', 'clean' = 'N', fmt = 0)
)
  #* Store models in list object
table_contents = list(
    'Party' = PartyModel,
    'Candidate evaluation' = VotePreTestModel
)
  #* Add row for sectioning out threshold estimates
rows = tribble(~term, ~`Party`, ~`Candidate evaluation`, 'Intercepts', " ", " ",)
attr(rows, 'position') = c(11)
  #* Generate table
modelsummary(table_contents, statistic = 'conf.int', coef_map = c('b_BlueTreatment' = 'Blue treatment', 'b_RedTreatment' = 'Red treatment', 'b_PartyId' = 'Party ID', 'b_BlueTreatment:PartyId' = 'Blue treatment $\\times$ Party ID', 'b_RedTreatment:PartyId' = 'Red treatment $\\times$ Party ID', 'b_Intercept[1]' = 'Threshold 1', 'b_Intercept[2]' = 'Threshold 2'), gof_map = gm, add_rows = rows, notes = c('95-percent credible intervals in brackets.', 'Median estimate from fitted model with 6 chains and 2000 iterations.', 'Data source: Pre-test experiment.'), escape = FALSE)
```

{{< pagebreak >}}

# Study 1

## Stimuli

![Example of treatment with blur](../../assets/img/example_of_treatment.png)

## Trial 1

![Blue](../study_1/assets/dem_yard_sign_round_1.png)

![Red](../study_1/assets/rep_yard_sign_round_1.png)

![White](../study_1/assets/cont_yard_sign_round_1.png)

{{< pagebreak >}}

### Trial 2

![Blue](../study_1/assets/dem_yard_sign_round_2.png)

![Red](../study_1/assets/rep_yard_sign_round_2.png)

![White](../study_1/assets/cont_yard_sign_round_2_3.png)

{{< pagebreak >}}

### Trial 3

![Blue](../study_1/assets/dem_yard_sign_round_3.png)

![Red](../study_1/assets/rep_yard_sign_round_3.png)

![White](../study_1/assets/cont_yard_sign_round_2_3.png)

{{< pagebreak >}}

## Measures

+---------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+-----------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------+
| Measure             | Question                                                                                                                                                                                                                                                                                                                                               | Response options                                                | Coded as                                                                                                                 |
+=====================+:======================================================================================================================================================================================================================================================================================================================================================:+:===============================================================:+:========================================================================================================================:+
| age                 | Confirm your age                                                                                                                                                                                                                                                                                                                                       | Any integer                                                     | Any integer                                                                                                              |
+---------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+-----------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------+
| gender              | gender_id = What is your gender identity?                                                                                                                                                                                                                                                                                                              | 1 = Non-binary                                                  | 1 = Male                                                                                                                 |
|                     |                                                                                                                                                                                                                                                                                                                                                        |                                                                 |                                                                                                                          |
|                     |                                                                                                                                                                                                                                                                                                                                                        | 2 = Transgender                                                 | 2 = Non-binary, prefer not to choose,other                                                                               |
|                     |                                                                                                                                                                                                                                                                                                                                                        |                                                                 |                                                                                                                          |
|                     |                                                                                                                                                                                                                                                                                                                                                        | 3 = Female                                                      | 3 = Female                                                                                                               |
|                     |                                                                                                                                                                                                                                                                                                                                                        |                                                                 |                                                                                                                          |
|                     |                                                                                                                                                                                                                                                                                                                                                        | 4 = Male                                                        |                                                                                                                          |
|                     |                                                                                                                                                                                                                                                                                                                                                        |                                                                 |                                                                                                                          |
|                     |                                                                                                                                                                                                                                                                                                                                                        | 5 = Prefer not to choose                                        |                                                                                                                          |
|                     |                                                                                                                                                                                                                                                                                                                                                        |                                                                 |                                                                                                                          |
|                     |                                                                                                                                                                                                                                                                                                                                                        | 6 = Other                                                       |                                                                                                                          |
+---------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+-----------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------+
| Sex                 | sex = What is your sex assigned at birth?                                                                                                                                                                                                                                                                                                              | 1 = Female                                                      | 0 = Female/Other                                                                                                         |
|                     |                                                                                                                                                                                                                                                                                                                                                        |                                                                 |                                                                                                                          |
|                     |                                                                                                                                                                                                                                                                                                                                                        | 2 = Male                                                        | 1 = Male                                                                                                                 |
|                     |                                                                                                                                                                                                                                                                                                                                                        |                                                                 |                                                                                                                          |
|                     |                                                                                                                                                                                                                                                                                                                                                        | 3 = Other                                                       |                                                                                                                          |
+---------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+-----------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------+
| White               | race_id = Which ethnic or racial category best describes you?                                                                                                                                                                                                                                                                                          | 1 = White, non-Hispanic                                         | 0 = Black, non-Hispanic \| Hispanic \| Asian or Native Hawaiian/other Pacific Islander \| Multiple Races                 |
|                     |                                                                                                                                                                                                                                                                                                                                                        |                                                                 |                                                                                                                          |
|                     |                                                                                                                                                                                                                                                                                                                                                        | 2 = Black, non-Hispanic                                         | 1 = White, non-Hispanic                                                                                                  |
|                     |                                                                                                                                                                                                                                                                                                                                                        |                                                                 |                                                                                                                          |
|                     |                                                                                                                                                                                                                                                                                                                                                        | 3 = Hispanic                                                    |                                                                                                                          |
|                     |                                                                                                                                                                                                                                                                                                                                                        |                                                                 |                                                                                                                          |
|                     |                                                                                                                                                                                                                                                                                                                                                        | 4 = Asian or Native Hawaiian/other Pacific Islander             |                                                                                                                          |
|                     |                                                                                                                                                                                                                                                                                                                                                        |                                                                 |                                                                                                                          |
|                     |                                                                                                                                                                                                                                                                                                                                                        | 5 = Native American/Alaska Native or other race                 |                                                                                                                          |
|                     |                                                                                                                                                                                                                                                                                                                                                        |                                                                 |                                                                                                                          |
|                     |                                                                                                                                                                                                                                                                                                                                                        | 6 = Multiple races                                              |                                                                                                                          |
+---------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+-----------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------+
| ColorBlind          | -   color_blind = Have you been diagnosed with any of the following visual impairments?                                                                                                                                                                                                                                                                | 1 = Blurred vision                                              | 0 = None of the above                                                                                                    |
|                     |                                                                                                                                                                                                                                                                                                                                                        |                                                                 |                                                                                                                          |
|                     |                                                                                                                                                                                                                                                                                                                                                        | 2 = Macular degeneration                                        | 1 = Blurred vision \| Macular degeneration \| Glaucoma \| Cataract \| Diabetic retinopathy \| Color blindness (any form) |
|                     |                                                                                                                                                                                                                                                                                                                                                        |                                                                 |                                                                                                                          |
|                     |                                                                                                                                                                                                                                                                                                                                                        | 3 = Glaucoma                                                    |                                                                                                                          |
|                     |                                                                                                                                                                                                                                                                                                                                                        |                                                                 |                                                                                                                          |
|                     |                                                                                                                                                                                                                                                                                                                                                        | 4 = Cataract                                                    |                                                                                                                          |
|                     |                                                                                                                                                                                                                                                                                                                                                        |                                                                 |                                                                                                                          |
|                     |                                                                                                                                                                                                                                                                                                                                                        | 5 = Diabetic retinopathy                                        |                                                                                                                          |
|                     |                                                                                                                                                                                                                                                                                                                                                        |                                                                 |                                                                                                                          |
|                     |                                                                                                                                                                                                                                                                                                                                                        | 6 = Color blindness (any form)                                  |                                                                                                                          |
|                     |                                                                                                                                                                                                                                                                                                                                                        |                                                                 |                                                                                                                          |
|                     |                                                                                                                                                                                                                                                                                                                                                        | 7 = None of the above                                           |                                                                                                                          |
+---------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+-----------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------+
| AttentionFlipped    | -   attention = How often do you pay attention to what is going on in government and politics                                                                                                                                                                                                                                                          | 1 = Always                                                      | 1 = Never                                                                                                                |
|                     |                                                                                                                                                                                                                                                                                                                                                        |                                                                 |                                                                                                                          |
|                     |                                                                                                                                                                                                                                                                                                                                                        | 2 = Most of the time                                            | 2 = Some of the time                                                                                                     |
|                     |                                                                                                                                                                                                                                                                                                                                                        |                                                                 |                                                                                                                          |
|                     |                                                                                                                                                                                                                                                                                                                                                        | 3 = About half the time                                         | 3 = About half the time                                                                                                  |
|                     |                                                                                                                                                                                                                                                                                                                                                        |                                                                 |                                                                                                                          |
|                     |                                                                                                                                                                                                                                                                                                                                                        | 4 = Some of the time                                            | 4 = Most of the time                                                                                                     |
|                     |                                                                                                                                                                                                                                                                                                                                                        |                                                                 |                                                                                                                          |
|                     |                                                                                                                                                                                                                                                                                                                                                        | 5 = Never                                                       | 5 = Always                                                                                                               |
+---------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+-----------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------+
| Knowledge           | -   know_promise =                                                                                                                                                                                                                                                                                                                                     | -   1 = I promise to try my best without looking up any answers | For each question:                                                                                                       |
|                     |                                                                                                                                                                                                                                                                                                                                                        |                                                                 |                                                                                                                          |
|                     |     We are interested in the guesses people make when they do not know the answer to a question. We will ask you several questions. Some may be easy, but some are meant to be so difficult that you will have to guess.\</p\>\<p\> In fact, for some of these questions, if you answer correctly, we will know that you probably looked up the answer | -   2 = I do not want to make the promise                       | 0 = Incorrect                                                                                                            |
|                     |                                                                                                                                                                                                                                                                                                                                                        |                                                                 |                                                                                                                          |
|                     |     Please do not look up the answers you do not know. Instead, please just make your best guess.                                                                                                                                                                                                                                                      | -   Any integer                                                 | 1 = Correct                                                                                                              |
|                     |                                                                                                                                                                                                                                                                                                                                                        |                                                                 |                                                                                                                          |
|                     |     Will you please promise to try your best without looking up any answers? Or do you not want to make that promise?                                                                                                                                                                                                                                  | -   1 = Foreign aid                                             | Average of correct responses                                                                                             |
|                     |                                                                                                                                                                                                                                                                                                                                                        |                                                                 |                                                                                                                          |
|                     | -   know_senate = For how many years is a United States Senator elected \-- that is, how many years are there in one full term of office for a U.S. Senator?                                                                                                                                                                                           | -   2 = Medicare                                                |                                                                                                                          |
|                     |                                                                                                                                                                                                                                                                                                                                                        |                                                                 |                                                                                                                          |
|                     | -   know_spend = On which of the following does the U.S. federal government currently spend the least?                                                                                                                                                                                                                                                 | -   3 = National defense                                        |                                                                                                                          |
|                     |                                                                                                                                                                                                                                                                                                                                                        |                                                                 |                                                                                                                          |
|                     | -   know_house = Do you happen to know which party currently has the most members in the U.S. House of representatives in Washington?                                                                                                                                                                                                                  | -   4 = Social security                                         |                                                                                                                          |
|                     |                                                                                                                                                                                                                                                                                                                                                        |                                                                 |                                                                                                                          |
|                     | -   know_catch = In what year did the Supreme Court of the united states decide Geer v. Connecticut?                                                                                                                                                                                                                                                   | -   1 = Democrats                                               |                                                                                                                          |
|                     |                                                                                                                                                                                                                                                                                                                                                        |                                                                 |                                                                                                                          |
|                     |                                                                                                                                                                                                                                                                                                                                                        | -   2 = Republicans                                             |                                                                                                                          |
|                     |                                                                                                                                                                                                                                                                                                                                                        |                                                                 |                                                                                                                          |
|                     |                                                                                                                                                                                                                                                                                                                                                        | -   Any integer                                                 |                                                                                                                          |
+---------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+-----------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------+
| PartyId             | -   pid_3 = Generally speaking, do you usually think of yourself as a Republican, a Democrat, an Independent or what?                                                                                                                                                                                                                                  | -   1 = Republican                                              | -   -3 = Strong democrat                                                                                                 |
|                     |                                                                                                                                                                                                                                                                                                                                                        |                                                                 |                                                                                                                          |
|                     | -   pid_d\_str\[pid_r\_str\] (if pid == 2\[1\]) = Would you call yourself a strong Democrat\[Republican\] or a not very strong Democrat\[Republican\]?                                                                                                                                                                                                 | -   2 = Democrat                                                |     -   pid == 2 & dem1 == 3                                                                                             |
|                     |                                                                                                                                                                                                                                                                                                                                                        |                                                                 |                                                                                                                          |
|                     | -   ind_lean (if pid == 3 \| 4) = Do you think of yourself as closer to the Republican or Democratic party?                                                                                                                                                                                                                                            | -   3 = Independent                                             | -   -2 = Democrat                                                                                                        |
|                     |                                                                                                                                                                                                                                                                                                                                                        |                                                                 |                                                                                                                          |
|                     |                                                                                                                                                                                                                                                                                                                                                        | -   4 = Other                                                   |     -   pid == 2 & dem1 == 2                                                                                             |
|                     |                                                                                                                                                                                                                                                                                                                                                        |                                                                 |                                                                                                                          |
|                     |                                                                                                                                                                                                                                                                                                                                                        | -   1 = Strong Democrat\[Republican\]                           | -   -1 = Leans Democratic                                                                                                |
|                     |                                                                                                                                                                                                                                                                                                                                                        |                                                                 |                                                                                                                          |
|                     |                                                                                                                                                                                                                                                                                                                                                        | -   2 = Not a very strong Democrat\[Republican\]                |     -   pid == 3 & ind1 == 2                                                                                             |
|                     |                                                                                                                                                                                                                                                                                                                                                        |                                                                 |                                                                                                                          |
|                     |                                                                                                                                                                                                                                                                                                                                                        | -   1 = Republican party                                        | -   0 = Independent                                                                                                      |
|                     |                                                                                                                                                                                                                                                                                                                                                        |                                                                 |                                                                                                                          |
|                     |                                                                                                                                                                                                                                                                                                                                                        | -   2 = Democratic party                                        |     -   pid == 3 & ind1 == 3                                                                                             |
|                     |                                                                                                                                                                                                                                                                                                                                                        |                                                                 |                                                                                                                          |
|                     |                                                                                                                                                                                                                                                                                                                                                        | -   3 = Neither party                                           | -   1 = Leans Republican                                                                                                 |
|                     |                                                                                                                                                                                                                                                                                                                                                        |                                                                 |                                                                                                                          |
|                     |                                                                                                                                                                                                                                                                                                                                                        |                                                                 |     -   pid == 3 & ind1 == 1                                                                                             |
|                     |                                                                                                                                                                                                                                                                                                                                                        |                                                                 |                                                                                                                          |
|                     |                                                                                                                                                                                                                                                                                                                                                        |                                                                 | -   2 = Republican                                                                                                       |
|                     |                                                                                                                                                                                                                                                                                                                                                        |                                                                 |                                                                                                                          |
|                     |                                                                                                                                                                                                                                                                                                                                                        |                                                                 |     -   pid == 1 & rep == 2                                                                                              |
|                     |                                                                                                                                                                                                                                                                                                                                                        |                                                                 |                                                                                                                          |
|                     |                                                                                                                                                                                                                                                                                                                                                        |                                                                 | -   3 = Strong Republican                                                                                                |
|                     |                                                                                                                                                                                                                                                                                                                                                        |                                                                 |                                                                                                                          |
|                     |                                                                                                                                                                                                                                                                                                                                                        |                                                                 |     -   pid == 1 & rep == 1                                                                                              |
+---------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+-----------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------+
| Party               | After viewing each image in each trial:                                                                                                                                                                                                                                                                                                                | -   1 = Republican                                              | 1 = Democrat                                                                                                             |
|                     |                                                                                                                                                                                                                                                                                                                                                        |                                                                 |                                                                                                                          |
|                     | Is this candidate a Republican, a Democrat, or Neither?                                                                                                                                                                                                                                                                                                | -   2 = Democrat                                                | 2 = neither                                                                                                              |
|                     |                                                                                                                                                                                                                                                                                                                                                        |                                                                 |                                                                                                                          |
|                     |                                                                                                                                                                                                                                                                                                                                                        | -   3 = Neither                                                 | 3 = Republican                                                                                                           |
+---------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+-----------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------+

: Study 1 Measures

{{< pagebreak >}}

## Results

{{< pagebreak >}}