### 2023-09-22 grants/ddirg/src/roberts_other_sources_of_support.qmd

---
subtitle: Other Sources of Support
format:
  fancy-doc-pdf:
    toc: false
    latex-output-dir: ../out/
    biblatex: true
    filters:
      #- "_extensions/andrewheiss/wordcount/citeproc.lua"
      - "_extensions/andrewheiss/wordcount/wordcount.lua"
author:
  - name: Damon C. Roberts
    affiliations:
      - name: Univeristy of Colorado Boulder
    email: damon.roberts-1@colorado.edu
bibliography: ../../../assets/references.bib
---

The first empirical chapter of this project has received a grant amounting to $2,054 to conduct the online experiment used in that component. This funding came from the Department of Political Science at the University of Colorado Boulder -- specifically, the Large Research Grant Fund.