### 2023-11-01 chapter_2/src/slides/color-book.qmd
updating the slides a bit more

---
title: "The Shape and <span style='color: #b70000'>Co</span><span style='color: #00AEF3'>lor</span> of Politics"
author:
    - name: Damon C. Roberts [{{< fa brands orcid size=xs >}}](https://orcid.org/0000-0002-4360-3675)
format:
    simpledarkjs-revealjs:
        multiplex: true
        output-file: color-book.html
        slide-level: 5
        transition: none
        slide-number: c
bibliography: ../../../assets/candidate_support_chapter_references.bib
execute:
    echo: false
    freeze: true
    eval: true
    cache: true
nocite: |
    @allaire_quarto_2022, @arel-bundock_marginaleffects_2022, @wickham_ggplot2_2016, @leeuw_jspsych_2015, @burkner_brms_2017, @zaller_simple_1992, @zaller_nature_1992, @lodge_rationalizing_2013, @taber_illusion_2016
---

```{r}
#| label: setup-block
# Set seed
set.seed(12102023)
# Load libraries
box::use(
  ggplot2[
    ggplot
    , aes
    , geom_density
    , theme_minimal
    , geom_vline
    , labs
    , scale_y_discrete
    , scale_x_discrete
  ]
  , ggdist[
    stat_halfeye
  ]
  , modelsummary[
    datasummary_skim
    , datasummary_balance
    , modelsummary
  ]
  , marginaleffects[
    avg_slopes
    , posterior_draws
  ]
  , ../R/predicted_prob_bar[...]
  , data.table[...]
)

# Load cleaned data
list_df <- list()
list_df[["cleaned"]] <- read.csv(
  file = "../../data/clean/cleaned_pavlovia.csv"
)
# Load fitted models
  #* Pre-test
load(
    file="../../data/models/prr/pre-test/pre-test_party_model.RDS"
  )
  #* Main studies
load("../../data/temp/models/study_1_fitted.RData")
# Load data for study 2
df_study_2 <- read.csv("../../data/clean/study_2.csv") |>
  as.data.table()
  #* convert state_district and year variables  to factors
df_study_2 <- df_study_2[
  , year := tstrsplit(year, "-", keep = 1)
][
  , year := as.numeric(year)
][
  , state_district := factor(state_district)
][
  , blue_prop := blue_percent / 100
][
  , red_prop := red_percent / 100
]

# Set default for plot color scheme

plot_color_scheme <- FALSE
```

## Available to you to follow along and view more

![QR Code](../../assets/slides/imgs/qr-code.png){#fig-qr}

---

- <span style="color: #b70000"> Motivation of the project</span>
- A snap-judgment model of politically-relevant information processing
- An experiment
- Is the world consistent with the snap-judgement model?
- The broader book project
- Q&A

## Yard signs are simple, yet effective forms of information

![Highway Yard Sign](../assets/img/highway_yardsign.jpg)

### Pre-test: Perceptions of candidate's party affiliation depends on the <span style='color: #b70000'>co</span><span style='color: #00AEF3'>lor</span> of the yard sign

```{r}
#| label: fig-hook-1
#| fig-cap: Perceptions of candidate's party affiliation based on color of yard sign

partyAMES <- avg_slopes(
  PartyModel,
  type = "link"
) |>
posterior_draws()

ggplot(
  data = partyAMES,
  aes(
    x = draw,
    y = term
  )
) +
stat_halfeye(
  slab_alpha = 0.45
) +
theme_minimal() +
labs(
  x = "Average Marginal Effect",
  y = "",
  caption = "Data source: Pre-test.\n Distribution of AME's for model posterior draws using HDI's.\n Interpreted as the logged odds of labelling the candidate as a partisan\n given whether they viewed a red yard sign or blue yard sign\n while holding all other variables at their mean value.\n Dot reflects median posterior draw where bars reflect 90% and 95% credible intervals."
) +
scale_y_discrete(
  labels = c(
    "Blue Treatment",
    "Red Treatment"
  )
) +
ggplot2::scale_x_continuous(
  breaks = c(-3, 0, 3)
  , labels = c(
    "Strong Democrat", "Independent", "Strong Republican"
  )
)
```

### <span style='color: #b70000'>Col</span><span style='color: #00AEF3'>ors</span> can convey identity

![Nebraska](../../assets/slides/imgs/nebraska.jpg){#fig-nebraska width=6.5in}

---

![[Example of brand UI]](../../assets/slides/imgs/green_bubble.jpeg){.absolute left=100 width=4.5in #fig-brand-ui}
![[The great divide]](../../assets/slides/imgs/green_bubble_funny.png){.absolute right=100 width=4.5in #fig-brand-divide}

## How are <span style='color: #b70000'>col</span><span style='color: #00AEF3'>ors</span> used in politics?

---

![[1860's Republican Ballots](https://slate.com/news-and-politics/2020/02/ballot-history-photo-essay.html)](../../assets/slides/imgs/ballots.avif){#fig-1860-ballot}

---

::: {#fig-1980-convention layout-ncol=2}

![[Democratic National Convention](https://untappedcities.com/2021/08/12/1980-democratic-national-convention/)](../../assets/slides/imgs/1980_dem_convention.jpg){#fig-1980-dem width=6.5in}

![[Republican National Convention]((../../assets/slides/imgs/1980_rep_convention.jpg))](../../assets/slides/imgs/1980_rep_convention.jpg){#fig-1980-rep width=6.5in}

1980 Presidential Party National Conventions
:::

---

![[1924 Election Map](https://collections.leventhalmap.org/search/commonwealth:q524n296b)](../../assets/slides/imgs/election_map.jpg){#fig-1924-map}

---

![[1976 Election Map](https://www.google.com/url?sa=i&url=https%3A%2F%2Fwww.washingtonpost.com%2Fnews%2Fthe-fix%2Fwp%2F2016%2F11%2F08%2Fred-vs-blue-a-brief-history-of-how-we-use-political-colors%2F&psig=AOvVaw1tL5UeACMye7NTm5hvZG-A&ust=1698159827386000&source=images&cd=vfe&opi=89978449&ved=0CBIQjhxqFwoTCKil57-4jIIDFQAAAAAdAAAAABAI)](../../assets/slides/imgs/1976_map.webp){#fig-1976-map}

---

![[2000 Presidential Election Map](https://www.google.com/url?sa=i&url=https%3A%2F%2Fwww.smithsonianmag.com%2Fhistory%2Fwhen-republicans-were-blue-and-democrats-were-red-104176297%2F&psig=AOvVaw3mlqWWKalyCl0gzVJDCIV5&ust=1694829600914000&source=images&cd=vfe&opi=89978449&ved=0CA8QjRxqFwoTCJCNg7bCq4EDFQAAAAAdAAAAABAT)](../../assets/slides/imgs/2000_election_map.webp){#fig-2000-election-map width=6.5in}

---

::: {#fig-party-brand layout-ncol=2}

![Republicans pre-2000](../assets/img/rnc_logo_pre-2000.png){#fig-republican}

![Democrats pre-2000](../assets/img/dnc_logo_1960.png){#fig-democrat}

![Republicans 2023](../assets/img/rnc_logo_2023.png){#fig-republican-post}

![Democrats 2023](../assets/img/dnc_logo_2023.png){#fig-democrat-post}

Party logos
:::

---

![[The public](https://www.brookings.edu/articles/what-is-a-brokered-convention-what-is-a-contested-convention/)](../../assets/slides/imgs/division.webp){#fig-public width=6.5in}

## What do we know so far about how we process political information?

### Political information is <span style='color: #b70000'>identity-based</span> and <span style='color: #00AEF3'>evokes affect</span>

::::: {.columns}

:::: {.column width="60%"}
::: {#fig-pip layout-ncol=2}
![](../../assets/slides/imgs/zaller.jpg){width=1.46in}
![](../../assets/slides/imgs/lodge_taber.jpg){width=1.46in}

Dominant models of political information processing
:::
::::

::::{.column width="40%"}

A key takeaway is that we are being pushed to react to information to before we even realize it.
::::

:::::

---

But what about visual information?

## [<span style='color: #b70000'>A snap-judgment model of</span> <span style='color: #00aef3'>visual political information processing</span>](#full-snap-judgment-model)

### Visual information is primitive

- We can process it faster, and automatically
- It is ubiquitous
- It can carry a significant amount of information
    - E.g., the color <span style='color: #FFCD00'>yellow</span> has associations with bananas, with our favorite sweater, with the Iowa Hawkeys, etc.
    - We have positive and negative affect with all of these.

### So, how can it be for politics?

- These associations have been strengthened over time as the consistent usage associating Republicans and Democrats.
- Our feelings toward the object that is associated with those colors can be activated because of those colors and the association.
- Because of the centrality of vision for many living organisms, our brain is designed around this. It happens first.
    - <span style='color: #b70000'>We process and re</span><span style='color: #00aef3'>act to color first.</span>
- Given this, we are already in a state of mind. We have made a snap-judgment that can be updated with new information, but nonetheless are influenced by it.

### Implications

:::: {.columns}

::: {.column width="50%"}

**Substantive Innovations**

- Presents a cognitive model that builds upon existing info. processing models to consider pre-conscious influences.
    - All of these steps happen <span style='color: #b70000'>bef</span><span style='color: #00aef3'>ore</span> the pre-existing info. processing models
:::

::: {.column width="50%"}

**Methodological Innovations**

- Using mousetracking with a survey experiment to examine pre-conscious motivations.
- Use image-based classification
:::

::::

---

- <span style="color: #616A6B">Motivation of the project</span>
- <span style="color: #616A6B">A snap-judgment model of politically-relevant information processing</span>
- <span style="color: #b70000">A survey experiment</span>
- Is the world consistent with the snap-judgement model?
- The broader book project
- Q&A

## Experiment

### Motivation of the experiment

- Do we associate the color <span style='color: #b70000'>red</span> with Republicans and the color <span style='color: #00aef3'>blue</span> with Democrats?
- Do these associations provide information that the public act on in politics?
- How do we interact with visual information?

### Setup of the experiment

#### [Background information](#tbl-study-1-measures)

- Age
- Education
- Gender Identity
- Sex
- Visual impairments
  - Including colorblindness
- Party Identification

#### Treatment

- Three trials

---

::: {.r-stack}

![](../study_1/assets/rep_yard_sign_round_1.png){.fragment width="6.5in"}

![](../study_1/assets/dem_yard_sign_round_1.png){.fragment width="6.5in"}

![](../study_1/assets/cont_yard_sign_round_1.png){.fragment width="6.5in"}

:::

---

::: {.r-stack}

![](../study_1/assets/rep_yard_sign_round_2.png){.fragment width=6.5in}

![](../study_1/assets/dem_yard_sign_round_2.png){.fragment width=6.5in}

![](../study_1/assets/cont_yard_sign_round_2_3.png){.fragment width=6.5in}

:::

---

::: {.r-stack}

![](../study_1/assets/rep_yard_sign_round_3.png){.fragment width=6.5in}

![](../study_1/assets/dem_yard_sign_round_3.png){.fragment width=6.5in}

![](../study_1/assets/cont_yard_sign_round_2_3.png){.fragment width=6.5in}

:::

#### Design

- Recruited [1,000 participants](#tbl-study-1-descriptives) from an online survey experiment platform; Prolific.
- Counterbalance treatment and background information ($\frac{1}{2}$ see treatment first; other $\frac{1}{2}$ see background questions first)
- Within each trial, participants have $\frac{1}{3}$ chance of seeing any given yard sign.
- MouseView.js [@anwyl-irvine_mouseviewjs_2022]
    - Use this instead of eye-tracking.
    - Do not need to use a convenience sample of students.

---

![Example of treatment](../../assets/slides/imgs/treatment.gif){#fig-treatment-example width=6.5in}

### A pre-registered expectation

- People associate the colors <span style='color: #b70000'>red</span> and <span style='color: #00aef3'>blue</span> with Republicans and Democrats, respectively.
  - Replicating the pre-test

#### [Red yard signs = Republican candidate](#h_2)

```{r}
#| label: fig-h-2-predictions-1
#| fig-cap: Red yard signs = Republican candidate
red <- predicted_prob_bar(
  fitted_model = list_fitted[["h_2"]][["red_multinomial"]]
  , x_axis = "trial_1_red_stimuli"
  , gray = plot_color_scheme
)
red
```

#### [<span style='color: #00aef3'>Blue yard signs = Democratic candidate</span>](#h_2)

```{r}
#| label: fig-h-2-predictions-2
#| fig-cap: Blue yard signs = Democratic candidate
blue <- predicted_prob_bar(
  fitted_model = list_fitted[["h_2"]][["blue_multinomial"]]
  , x_axis = "trial_1_blue_stimuli"
  , treatment = "Blue"
  , gray = plot_color_scheme
)
blue
```

### A pre-registered expectation

Subjects will use color as a basis for making a decision on whom they should vote for.

- Republican subjects will express preference for a candidate with a pre-dominantly <span style='color: #b70000'>red</span> yard sign.
- Democratic subjects will express preference for a candidate with a pre-dominantly <span style='color: #00aef3'>blue</span> yard sign.

#### [Republicans are more likely to vote for a candidate with a red yard sign, Democrats are less likely](#h_3)

```{r}
#| label: fig-h-3-predictions-1
#| fig-cap: Republicans are more likely to vote for candidate with red yard sign, Democrats less likely
red <- predicted_prob_bar(
  fitted_model = list_fitted[["h_3"]][["red"]]
  , x_axis = "trial_1_red_stimuli"
  , treatment = "Red"
  , hypothesis = "H3"
  , x_label = "Party ID"
  , y_label = "Pr(Vote for candidate)"
  , legend_title = "Treatment"
  , gray = plot_color_scheme
)

red
```

#### [<span style='color: #00aef3'>Democrats are more likely to vote for a candidate with a blue yard sign, Republicans less likely</span>](#h_3)

```{r}
#| label: fig-h-3-predictions-2
#| fig-cap: Democrats are more likely to vote for candidate with blue yard signs, Republicans less likely
blue <- predicted_prob_bar(
  fitted_model = list_fitted[["h_3"]][["blue"]]
  , x_axis = "trial_1_blue_stimuli"
  , treatment = "Blue"
  , hypothesis = "H3"
  , x_label = "Party ID"
  , y_label = "Pr(Vote for candidate)"
  , legend_title = "Treatment"
  , gray = plot_color_scheme
)

blue
```

### A pre-registered expectation

When you include conflicting information (<span style='color: #00aef3'>blue</span> on a <span style='color: #b70000'>red</span> yard sign; <span style='color: #b70000'>red</span> on a <span style='color: #00aef3'>blue</span> yard sign), subjects will be less certain about the partisanship of the candidate

#### [Mixing red and blue in a yard sign <span style='color: #00aef3'>increases uncertainty about partisanship</span>](#h_2a)

```{r}
#| label: fig-h-2a-predictions-trial-2
#| layout-ncol: 2
#| fig-cap: Mixing red and blue in a yard sign increases uncertainty about partisanship
#| fig-subcap:
#|   - Trial 2 Red treatment
#|   - Trial 2 Blue treatment
plot_trial_2_red <- predicted_prob_bar(
  fitted_model = list_fitted[["h_2a"]][["trial_2_red_multinomial"]]
  , x_axis = "trial_2_red_stimuli"
  , gray = plot_color_scheme
)
plot_trial_2_blue <- predicted_prob_bar(
  fitted_model = list_fitted[["h_2a"]][["trial_2_blue_multinomial"]]
  , x_axis = "trial_2_blue_stimuli"
  , treatment = "Blue"
  , gray = plot_color_scheme
)
plot_trial_3_red <- predicted_prob_bar(
  fitted_model = list_fitted[["h_2a"]][["trial_3_red_multinomial"]]
  , x_axis = "trial_3_red_stimuli"
  , gray = plot_color_scheme
)
plot_trial_3_blue <- predicted_prob_bar(
  fitted_model = list_fitted[["h_2a"]][["trial_3_blue_multinomial"]]
  , x_axis = "trial_3_blue_stimuli"
  , treatment = "Blue"
  , gray = plot_color_scheme
)
plot_trial_2_red
plot_trial_2_blue
```

---

```{r}
#| label: fig-h-2a-predictions-trial-3
#| layout-ncol: 2
#| fig-cap: Mixing red and blue in a yard sign increases uncertainty about partisanship (cont.)
#| fig-subcap:
#|   - Trial 3 Red treatment
#|   - Trial 3 Blue treatment

plot_trial_3_red
plot_trial_3_blue
```

### A pre-registered expectation

Subjects will spend more time and will move their cursor more when looking at <span style='color: #00aef3'>blue</span> and <span style='color: #b70000'>red</span> yard signs

#### When viewing a white yard sign, subjects explored the image more.

```{r}
#| label: tbl-h1-trial-1
#| tbl-cap: Average time between cursor movements (Trial 1)
df_movement_trial_1 <- list_df[["cleaned"]][
  , c(
    "trial_1_difference_Time"
    , "trial_1_stimulus"
    , "trial_1_difference_X"
    , "trial_1_difference_Y"
  )
]
setnames(
  df_movement_trial_1
  , old = c(
    "trial_1_difference_Time"
    , "trial_1_difference_X"
    , "trial_1_difference_Y"
  )
  , new = c(
    "&#x394<sub>t</sub> (milliseconds)"
    , "&#x394<sub>x</sub> (pixels)"
    , "&#x394<sub>y</sub> (pixels)"
  )
)

datasummary_balance(
  ~trial_1_stimulus
  , data = df_movement_trial_1
  , notes = "Data source: Prolific sample."
  , escape = FALSE
)
```

#### A pre-registered expectation

Subjects will take longer than they did in trial 1 to explore the yard signs when there are conflicting colors on the yard sign

#### The average subject in trial 2 explored the image more than the average participant in trial 1 for <span style='color: #b70000'>red</span> and <span style='color: #00aef3'>blue</span> yard signs

```{r}
#| label: tbl-h1-trial-2
#| tbl-cap: Average time between cursor movements (Trial 2)
df_movement_trial_2 <- list_df[["cleaned"]][
  , c(
    "trial_2_difference_Time"
    , "trial_2_stimulus"
    , "trial_2_difference_X"
    , "trial_2_difference_Y"
  )
]
setnames(
  df_movement_trial_2
  , old = c(
    "trial_2_difference_Time"
    , "trial_2_difference_X"
    , "trial_2_difference_Y"
  )
  , new = c(
    "&#x394<sub>t</sub> (milliseconds)"
    , "&#x394<sub>x</sub> (pixels)"
    , "&#x394<sub>y</sub> (pixels)"
  )
)
modelsummary::datasummary_balance(
  ~trial_2_stimulus
  , data = df_movement_trial_2
  , notes = "Data source: Prolific sample."
  , escape = FALSE
)
```

#### The average subject in trial 3 spent more time exploring the <span style='color: #b70000'>red</span> and <span style='color: #00aef3'>blue</span> yard signs than the average participant in trial 1

```{r}
#| label: tbl-h1-trial-3
#| tbl-cap: Average time between cursor movements (Trial 3)
df_movement_trial_3 <- list_df[["cleaned"]][
  , c(
    "trial_3_difference_Time"
    , "trial_3_stimulus"
    , "trial_3_difference_X"
    , "trial_3_difference_Y"
  )
]
setnames(
  df_movement_trial_3
  , old = c(
    "trial_3_difference_Time"
    , "trial_3_difference_X"
    , "trial_3_difference_Y"
  )
  , new = c(
    "&#x394<sub>t</sub> (milliseconds)"
    , "&#x394<sub>x</sub> (pixels)"
    , "&#x394<sub>y</sub> (pixels)"
  )
)

modelsummary::datasummary_balance(
  ~trial_3_stimulus
  , data = df_movement_trial_3
  , notes = "Data source: Prolific sample."
  , escape = FALSE
)
```

---

- <span style="color: #616A6B">Motivation of the project</span>
- <span style="color: #616A6B">A snap-judgment model of politically-relevant information processing</span>
- <span style="color: #616A6B">A survey experiment</span>
- <span style="color: #b70000">Is the world consistent with the snap-judgement model?</span>
- The broader book project
- Q&A

## Study 2 - Is the world consistent with my theory?

### A pre-registered expectation

Campaigns use the colors more or less <span style='color: #b70000'>red</span> and <span style='color: #00aef3'>blue</span> depending on partisan composition of electoral district

####  Talking to practitioners

> There obviously is a strong association with blue and the Democratic Party and red and the Republican Party. We've seen in some of our testing that specifically MAGA red tends to elicit a slightly stronger association with Republicans/conservatives/Trump than other shades.

---

> A lot of decisions are still made on purely gut feelings as opposed to taking a more multi-disciplinary evidence-based approach.

---

It remains unclear whether this is just ***one*** group of practitioners, however.

#### [<span style='color: #b70000'>D</span><span style='color: #00aef3'>V</span>: the amount of red <span style='color: #00aef3'>and blue in political yard signs</span>](#tbl-yard-sign-summary)

- Collect around 1,000 yard signs from House elections
  - 2016-2022
- Measuring the colors used on a yard sign:
  - Use a machine learning Python Library, `OpenCV` [@bradski_opencv_2000], that reads images and classifies their components.
    - Trained to classify each pixel in the image
    - Classifies it based on BGR value (reverse of RGB)
    - Looks for a range of values around the "Republican red" and "Democratic blue"

---

```{python}
#| label: example detection
#| echo: false

from sys import path # for path management
from cv2 import imread, imwrite # for color detection
    #* user-defined
sys.path.append("../")
from PY.helper import colorDetector

# Define colors to detect
    #* White
        #** Not defined. Default for colorDetector()
    #* Red
republican_red = [232, 27, 35] # target color
red_lower = [93, 9, 12] # lower end of spectrum for red
red_higher = [255, 170, 170] # higher end of spectrum for red

# Load gop image to read
img = imread("../assets/img/gop_2022.png")

# Detect colors
percent, img_transformed, result = colorDetector(img = img, color_upper = red_higher, color_lower = red_lower)

transformed= imwrite("../../data/temp/study_2/gop_2022_transformed.png", img_transformed) # save transformed image

masked= imwrite("../../data/temp/study_2/gop_2022_detected.png", result) # save masked image
```

::: {#fig-color-detection-example layout-ncol="2"}
![Resized original image](../../data/temp/study_2/gop_2022_transformed.png){#fig-resized}

![Masked](../../data/temp/study_2/gop_2022_detected.png){#fig-masked}

Detecting colors in the GOP logo
:::

```{python}
#| output: asis
# Print calculated percentage of example
print("For the example in @fig-color-detection-example, about {:.2f}".format(percent) + " of the image is red.")
```

#### IV: partisan composition of the district

- MIT Election Lab data
  - House election results
  - 1970-2020
- Measuring partisan composition of district:
  - Calculate rolling 5-year average of the Democratic party's vote share for each district

### [<span style='color: #00aef3'>More blue on the yard signs in Democratic strongholds</span>](#h_5)

```{r}
#| label: fig-study-2-pred-1
#| fig-cap: More blue on the yard signs in Democratic strongholds

blue <- predicted_prob_bar(
  fitted_model = list_fitted[["h_5"]][["blue"]]
  , treatment = "Blue"
  , hypothesis = "H5"
  , gray = plot_color_scheme
)

blue
```

### [More red on the yard signs in Republican strongholds](#h_5)

```{r}
#| label: fig-study-2-pred-2
#| fig-cap: Less red on the yard signs in Democratic strongholds

red <- predicted_prob_bar(
  fitted_model = list_fitted[["h_5"]][["red"]]
  , treatment = "Red"
  , hypothesis = "H5"
  , gray = plot_color_scheme
)
red
```

---

- <span style="color: #616A6B">Motivation of the project</span>
- <span style="color: #616A6B">A snap-judgment model of politically-relevant information processing</span>
- <span style="color: #616A6B">A survey experiment</span>
- <span style="color: #616A6b">Is the world consistent with the snap-judgement model?</span>
- <span style="color: #b70000">The broader book project</span>
- Q&A

## How this fits in with the rest of the book?

- <span style="color: #616A6B">How do colors convey political information and effect individual attitudes? </span>
- How do colors influence political deliberation?
- How do colors influence neighborhood evaluations?

## How my book fits in with the rest of my work?

![](../../assets/slides/imgs/roberts_gs.png)

## Questions and Comments

## References

::: {#refs}
:::

## Appendix

### Full snap-judgment model

{{< include ../../../assets/_theory.qmd >}}

### Experiment

#### Characteristics of experimental sample

```{r}
#| label: tbl-study-1-descriptives
#| tbl-cap: Descriptive statistics of Prolific sample

df_demographics_subset <- list_df[["cleaned"]][
    , c(
        "Age" = "age"
        , `Color blind` = "color_blind_dummy"
        , "Female" = "female_sex"
        , "White" = "white"
        , "Hispanic" = "hispanic"
        , "Black" = "black"
        , `Party ID` = "pid_7"
    )
] |> 
as.data.frame()

datasummary_skim(
    df_demographics_subset
    , histogram = FALSE
    , notes = c(
      "Data source: Prolific."
      , "Sample characteristics."
    )

)
```

#### Measures

+-----------------------------+-----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| Measure                     | Question                                                                                                                                                                                                                                                                                                  | Coded as                                                                                                                                                           |
+=============================+:=========================================================================================================================================================================================================================================================================================================:+:==================================================================================================================================================================:+
| Age                         | Confirm your age                                                                                                                                                                                                                                                                                          | Text entry                                                                                                                                                         |
+-----------------------------+-----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| Gender                      | What is your gender identity?                                                                                                                                                                                                                                                                             | Non-binary, Female, Male, Prefer not to choose, Other                                                                                                              |
+-----------------------------+-----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| Sex                         | What is your sex assigned at birth?                                                                                                                                                                                                                                                                       | Female, Male, Other                                                                                                                                                |
+-----------------------------+-----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| Race                        | Which ethnic or racial category best describes you?                                                                                                                                                                                                                                                       | White, non-Hispanic; Black, non-Hispanic; Hispanic; Asian or Native Hawaiian/other Pacific Islander; Native American/Alaska Native or other race; Multiple races   |
+-----------------------------+-----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| Color blind                 | Have you been diagnosed with any of the following visual impairments that are not treated with corrective lenses?                                                                                                                                                                                         | Color blindness (any form); Blurred vision (not treatable by corrective lenses); Macular degeneration; Glaucoma; Cataract; Diabetic retinopathy; None of the above |
+-----------------------------+-----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| Attention                   | How often do you pay attention to what is going on in government and politics?                                                                                                                                                                                                                            | Always; Most of the time; About half the time; Some of the time; Never                                                                                             |
+-----------------------------+-----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| Party ID                    | Generally speaking, do you usually think of yourself as a Democrat, a Republican, an Independent, or what?; Would you say that you are a strong [Republican/Democrat] or not a very strong [Republican/Democrat]?; Do you think of yourself as closer to the Republican party or to the Democratic party? | Republican, Democrat, Independent, Other; Strong, Not very strong; Closer to the Republican party, Neither, Closer to Democratic                                   |
+-----------------------------+-----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| Attn. Check                 | In a few words describe what is your first memory of a political event.                                                                                                                                                                                                                                   | Text entry                                                                                                                                                         |
+-----------------------------+-----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| Trial stimuli               | Yard sign assigned                                                                                                                                                                                                                                                                                        | 0 = White, 1 = Red/Blue                                                                                                                                            |
+-----------------------------+-----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| Trial stimuli alt           | Yard sign assigned                                                                                                                                                                                                                                                                                        | 0 = Not red/blue, 1 = Red/Blue                                                                                                                                     |
+-----------------------------+-----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| Time elapsed                | NA                                                                                                                                                                                                                                                                                                        | Difference between time at start of trial and stop                                                                                                                 |
+-----------------------------+-----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| $\Delta_{t}$ (milliseconds) | NA                                                                                                                                                                                                                                                                                                        | Time elapsed between cursor movements                                                                                                                              |
+-----------------------------+-----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| $\Delta_{x}$ (pixels)       | NA                                                                                                                                                                                                                                                                                                        | Number of pixels moved along x-axis                                                                                                                                |
+-----------------------------+-----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| $\Delta_{y}$ (pixels)       | NA                                                                                                                                                                                                                                                                                                        | Number of pixels moved along y-axis                                                                                                                                |
+-----------------------------+-----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| Candidate party             | Is this candidate a Republican, a Democrat, or Neither?                                                                                                                                                                                                                                                   | Republican, Democrat, Neither                                                                                                                                      |
+-----------------------------+-----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| Candidate vote              | Of the three candidates who would you most likely vote for?                                                                                                                                                                                                                                               | The first yard sign, The second yard sign, The third yard sign                                                                                                     |
+-----------------------------+-----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| Candidate vote indicator    | NA                                                                                                                                                                                                                                                                                                        | 0 = Would not vote for first yard sign, 1 = Would vote for first yard sign                                                                                         |
+-----------------------------+-----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------+

: Study 1 Measures {#tbl-study-1-measures}

### List of model specifications for Experiment

- [Moderated by age](#moderated-by-age)
- [Moderated by colorblindness](#moderated-by-colorblindness)
- [Dependent on the amount of the color red and blue](#h_2a)
- [Include treatment indicator variables in the same model](#single-model)
- [Multinomial (main) versus Ordered logistic regressions](#h_2)

### $H_2$

#### Multinomial Logit

```{r}
#| label: tbl-hypothesis-2-main
#| tbl-cap: Full results for reported models in $H_2$
# Redefine names
list_fitted_split_1 <- list(
  "Red treatment" = list_fitted[["h_2"]][["red_multinomial"]]
  , "Blue treatment" = list_fitted[["h_2"]][["blue_multinomial"]]
)

# Change goodness of fit stuff
gm <- list(
  list("raw" = "nobs", "clean" = "Num.Obs.", "fmt" = 1)
  , list("raw" = "r.squared", "clean" = "R<sup>2</sup>", "fmt" = 2)
  , list("raw" = "waic", "clean" = "WAIC", "fmt" = 2)
)
# Make table
modelsummary(
  list_fitted_split_1
  , coef_map = c(
    "b_mu2_trial_1_red_stimuli" = "Red yard sign (Independent)"
    , "b_mu3_trial_1_red_stimuli" = "Red yard sign (Republican)"
    , "b_mu2_trial_1_blue_stimuli" = "Blue yard sign (Independent)"
    , "b_mu3_trial_1_blue_stimuli" = "Blue yard sign (Republican)"
    , "b_mu2_Intercept" = "Intercept - Independent"
    , "b_mu3_Intercept" = "Intercept - Republican"
    )
  , statistic = "conf.int"
  , conf_level = 0.9
  , gof_map = gm 
  , notes = c(
    "Data source: Pavlovia"
    , "Positive coefficients reflect the logged-odds"
    , "of claiming candidate is not a Democrat."
    , "Numbers inside brackets reflect the 5th and 95th percentile"
    , "of the draws from the posterior distribution."
  )
  , escape = FALSE

)
```

##### Posterior Predictive Check

```{r}
#| label: fig-h-2-ppc-multinomial
#| fig-cap: Posterior predictive check of Multinomial models for $H_2$
#| fig-subcap:
#|   - Red treatment
#|   - Blue treatment

list_ppc[["h_2"]][["red_multinomial"]]
list_ppc[["h_2"]][["blue_multinomial"]]
```

#### Ordered Logit

```{r}
#| label: tbl-hypothesis-2-ordered
#| tbl-cap: Full results for reported ordered logistic regression models for $H_2$
# Redefine names
list_fitted_split_1_alt <- list(
  "Red treatment" = list_fitted[["h_2"]][["red_ordinal"]]
  , "Blue treatment" = list_fitted[["h_2"]][["blue_ordinal"]]
)

# Change goodness of fit stuff
gm <- list(
  list("raw" = "nobs", "clean" = "Num.Obs.", "fmt" = 1)
  , list("raw" = "r.squared", "clean" = "R<sup>2</sup>", "fmt" = 2)
  , list("raw" = "waic", "clean" = "WAIC", "fmt" = 2)
)
# Make table
modelsummary(
  list_fitted_split_1_alt
  , coef_map = c(
    "b_trial_1_red_stimuli" = "Red yard sign"
    , "b_trial_1_blue_stimuli" = "Blue yard sign"
    , "b_Intercept[1]" = "Intercept 1"
    , "b_Intercept[2]" = "Intercept 2"
    )
  , statistic = "conf.int"
  , conf_level = 0.9
  , gof_map = gm 
  , notes = c(
    "Data source: Pavlovia"
    , "Positive coefficients reflect the logged-odds"
    , "of claiming candidate is Republican."
    , "Numbers inside brackets reflect the 5th and 95th percentile"
    , "of the draws from the posterior distribution."
  )
  , escape = FALSE

)
```

##### Posterior Predictive Check

```{r}
#| label: fig-h-2-ppc-ordered
#| fig-cap: Posterior predictive checks of Ordered Logit models in $H_2$
#| fig-subcap:
#|   - Red treatment
#|   - Blue treatment

list_ppc[["h_2"]][["red_ordinal"]]
list_ppc[["h_2"]][["blue_ordinal"]]
```

#### Single model

```{r}
#| label: tbl-robustness-alt-specification-h-2
#| tbl-cap: $H_2$ models with alternative specification
# Redefine names
list_fitted_split_3 <- list(
  "Single model" = list_fitted[["h_2"]][["single_model"]]
)

# Make table
modelsummary(
  list_fitted_split_3
  , coef_map = c(
    "b_trial_1_red_stimuli_alt" = "Red yard sign"
    , "b_trial_1_blue_stimuli_alt" = "Blue yard sign"
    , "b_Intercept[1]" = "Intercept 1"
    , "b_Intercept[2]" = "Intercept 2"
  )
  , statistic = "conf.int"
  , conf_level = 0.9
  , gof_map = gm
  , notes = c(
    "Data source: Pavlovia"
    , "Positive coefficients reflect the logged-odds"
    , " of claiming candidate is Republican."
    , "Numbers inside brackets reflect the 5th and 95th percentile"
    , " of the draws from the posterior distribution."
  )
  , escape = FALSE

)
```

##### Posterior Predictive Check

```{r}
#| label: fig-h-2-ppc-single
#| fig-cap: Posterior predictive checks of single model specification for $H_2$

list_ppc[["h_2"]][["single_model"]]

```

#### Moderated by age

```{r}
#| label: tbl-interacted-models-h-2
#| tbl-cap: $H_2$ models moderated by age
# Redefine names
list_fitted_split_2 <- list(
  "Red treatment - Age moderating" = list_fitted[["h_2"]][["red_interact"]]
  , "Blue treatment - Age moderating" = list_fitted[["h_2"]][["blue_interact"]]
)

# Make table
modelsummary(
  list_fitted_split_2
  , coef_map = c(
    "b_trial_1_red_stimuli" = "Red yard sign"
    , "b_trial_1_blue_stimuli" = "Blue yard sign"
    , "b_age" = "Age"
    , "b_trial_1_red_stimuli:age" = "Red yard sign × Age"
    , "b_trial_1_blue_stimuli:age" = "Blue yard sign × Age"
    , "b_Intercept[1]" = "Intercept 1"
    , "b_Intercept[2]" = "Intercept 2"
    )
  , statistic = "conf.int"
  , conf_level = 0.9
  , gof_map = gm 
  , notes = c(
    "Data source: Pavlovia"
    , "Positive coefficients reflect the logged-odds"
    , " of claiming candidate is Republican."
    , "Numbers inside brackets reflect the 5th and 95th percentile"
    , " of the draws from the posterior distribution."
  )
  , escape = FALSE

)
```

##### Posterior Predictive Check

```{r}
#| label: fig-h-2-ppc-age
#| fig-cap: Posterior predictive checks of models in $H_2$ moderated by age
#| fig-subcap:
#|   - Red treatment moderated by age
#|   - Blue treatment moderated by age

list_ppc[["h_2"]][["red_interact"]]
list_ppc[["h_2"]][["blue_interact"]]
```

#### Moderated by colorblindness

```{r}
#| label: tbl-colorblindness
#| tbl-cap: $H_2$ moderated by colorblindness

# Redefine names
list_fitted_split_4 <- list(
  "Red - Color blindness moderating" = list_fitted[["h_2"]][["red_colorblind"]]
  , "Blue - Color blindness moderating" = list_fitted[["h_2"]][["blue_colorblind"]]
)
# Make table
modelsummary(
  list_fitted_split_4
  , coef_map = c(
    "b_trial_1_red_stimuli" = "Red yard sign"
    , "b_trial_1_blue_stimuli" = "Blue yard sign"
    , "b_trial_1_red_stimuli:color_blind_dummy" = "Red yard sign × Color blind"
    , "b_trial_1_blue_stimuli:color_blind_dummy" = "Blue yard sign × Color blind"
    , "b_Intercept[1]" = "Intercept 1"
    , "b_Intercept[2]" = "Intercept 2"
  )
  , statistic = "conf.int"
  , conf_level = 0.9
  , gof_map = gm 
  , notes = c(
    "Data source: Pavlovia"
    , "Positive coefficients reflect the logged-odds"
    , " of claiming candidate is Republican."
    , "Numbers inside brackets reflect the 5th and 95th percentile"
    , " of the draws from the posterior distribution."
  )
  , escape = FALSE

)
```

##### Posterior Predictive Check

```{r}
#| label: fig-h-2-ppc-colorblind
#| fig-cap: Posterior predictive checks of models in $H_2$ moderated by colorblindness
#| fig-subcap:
#|   - Red treatment moderated by colorblindness
#|   - Blue treatment moderated by colorblindness

list_ppc[["h_2"]][["red_colorblind"]]
list_ppc[["h_2"]][["blue_colorblind"]]
```

### $H_{2a}$

#### Multinomial

```{r}
#| label: tbl-hypothesis-2a
#| tbl-cap: Full results for models testing bounds of $H_2$
# Redefine names
model_names <- list(
  "Trial 2 - Red treatment" = list_fitted[["h_2a"]][["trial_2_red_multinomial"]]
  , "Trial 2 - Blue treatment" = list_fitted[["h_2a"]][["trial_2_blue_multinomial"]]
  , "Trial 2 - Single model" = list_fitted[["h_2a"]][["trial_2_single_model"]]
)

# Change goodness of fit stuff
gm <- list(
  list("raw" = "nobs", "clean" = "Num.Obs.", "fmt" = 1)
  , list("raw" = "r.squared", "clean" = "R<sup>2</sup>", "fmt" = 2)
  , list("raw" = "waic", "clean" = "WAIC", "fmt" = 2)
)
# Make table
modelsummary(
  model_names
  , coef_map = c(
    "b_mu2_trial_2_red_stimuli" = "Red yard sign (Independent)"
    , "b_mu3_trial_2_red_stimuli" = "Red yard sign (Republican)"
    , "b_mu2_trial_2_blue_stimuli" = "Blue yard sign (Independent)"
    , "b_mu3_trial_2_blue_stimuli" = "Blue yard sign (Republican)"
    , "b_trial_2_red_stimuli_alt" = "Red yard sign"
    , "b_trial_2_blue_stimuli_alt" = "Blue yard sign"
    , "b_mu2_Intercept" = "Intercept - Independent"
    , "b_mu3_Intercept" = "Intercept - Republican"
    , "b_Intercept[1]" = "Intercept 1"
    , "b_Intercept[2]" = "Intercept 2"
    )
  , statistic = "conf.int"
  , conf_level = 0.9
  , gof_map = gm
  , notes = c(
    "Data source: Pavlovia"
    , "Positive coefficients reflect the logged-odds"
    , " of claiming candidate is not a Democrat (Columns 1 and 2) or is a Republican (Column 3)."
    , "Numbers inside brackets reflect the 5th and 95th percentile"
    , " of the draws from the posterior distribution."
  )
  , escape = FALSE

)
```

##### Posterior Predictive Check

```{r}
#| label: fig-h-2a-multinomial-trial-2-ppc
#| fig-cap: Posterior Predictive Checks for Multinomial models testing $H_{2a}$, Trial 2
#| fig-subcap:
#|   - Red treatment
#|   - Blue treatment
#|   - Single model

list_ppc[["h_2a"]][["trial_2_red_multinomial"]]
list_ppc[["h_2a"]][["trial_2_blue_multinomial"]]
list_ppc[["h_2a"]][["trial_2_single_model"]]
```

#### Multinomial Cont.

```{r}
#| label: tbl-hypothesis-2a-cont
#| tbl-cap: Full results for models testing bounds of $H_2$ (cont.)
# Redefine names
model_names <- list(
  "Trial 3 - Red treatment" = list_fitted[["h_2a"]][["trial_3_red_multinomial"]]
  , "Trial 3 - Blue treatment" = list_fitted[["h_2a"]][["trial_3_blue_multinomial"]]
  , "Trial 3 - Single model" = list_fitted[["h_2a"]][["trial_3_single_model"]]
)

# Change goodness of fit stuff
gm <- list(
  list("raw" = "nobs", "clean" = "Num.Obs.", "fmt" = 1)
  , list("raw" = "r.squared", "clean" = "R<sup>2</sup>", "fmt" = 2)
  , list("raw" = "waic", "clean" = "WAIC", "fmt" = 2)
)
# Make table
modelsummary(
  model_names
  , coef_map = c(
    "b_mu2_trial_3_red_stimuli" = "Red yard sign (Independent)"
    , "b_mu3_trial_3_red_stimuli" = "Red yard sign (Republican)"
    , "b_mu2_trial_3_blue_stimuli" = "Blue yard sign (Independent)"
    , "b_mu3_trial_3_blue_stimuli" = "Blue yard sign (Republican)"
    , "b_trial_3_red_stimuli_alt" = "Red yard sign"
    , "b_trial_3_blue_stimuli_alt" = "Blue yard sign"
    , "b_mu2_Intercept" = "Intercept - Independent"
    , "b_mu3_Intercept" = "Intercept - Republican"
    , "b_Intercept[1]" = "Intercept 1"
    , "b_Intercept[2]" = "Intercept 2"
    )
  , statistic = "conf.int"
  , conf_level = 0.9
  , gof_map = gm 
  , notes = c(
    "Data source: Pavlovia"
    , "Positive coefficients reflect the logged-odds"
    , " of claiming candidate is Republican."
    , "Numbers inside brackets reflect the 5th and 95th percentile"
    , " of the draws from the posterior distribution."
  )
  , escape = FALSE

)
```

##### Posterior Predictive Check

```{r}
#| label: fig-h-2a-multinomial-trial-3-ppc
#| fig-cap: Posterior Predictive Checks for Multinomial models testing $H_{2a}$, Trial 3
#| fig-subcap:
#|   - Red treatment
#|   - Blue treatment
#|   - Single model

list_ppc[["h_2a"]][["trial_3_red_multinomial"]]
list_ppc[["h_2a"]][["trial_3_blue_multinomial"]]
list_ppc[["h_2a"]][["trial_3_single_model"]]
```

#### Ordinal model

```{r}
#| label: tbl-hypothesis-2a-ordered
#| tbl-cap: Full results for ordered logistic regressions testing bounds of $H_2$
# Redefine names
model_names <- list(
  "Trial 2 - Red treatment" = list_fitted[["h_2a"]][["trial_2_red_ordinal"]]
  , "Trial 2 - Blue treatment" = list_fitted[["h_2a"]][["trial_2_blue_ordinal"]]
)

# Change goodness of fit stuff
gm <- list(
  list("raw" = "nobs", "clean" = "Num.Obs.", "fmt" = 1)
  , list("raw" = "r.squared", "clean" = "R<sup>2</sup>", "fmt" = 2)
  , list("raw" = "waic", "clean" = "WAIC", "fmt" = 2)
)
# Make table
modelsummary(
  model_names
  , coef_map = c(
    "b_trial_2_red_stimuli" = "Red yard sign"
    , "b_trial_2_blue_stimuli" = "Blue yard sign"
    , "b_Intercept[1]" = "Intercept 1"
    , "b_Intercept[2]" = "Intercept 2"
    )
  , statistic = "conf.int"
  , conf_level = 0.9
  , gof_map = gm
  , notes = c(
    "Data source: Pavlovia"
    , "Positive coefficients reflect the logged-odds"
    , " of claiming candidate is Republican."
    , "Numbers inside brackets reflect the 5th and 95th percentile"
    , " of the draws from the posterior distribution."
  )
  , escape = FALSE

)
```

##### Posterior Predictive Check

```{r}
#| label: fig-h-2a-ordinal-trial-2-ppc
#| fig-cap: Posterior Predictive Checks for Ordinal models testing $H_{2a}$, Trial 2
#| fig-subcap:
#|   - Red treatment
#|   - Blue treatment

list_ppc[["h_2a"]][["trial_2_red_ordinal"]]
list_ppc[["h_2a"]][["trial_2_blue_ordinal"]]
```

#### Ordered Cont.

```{r}
#| label: tbl-hypothesis-2a-cont-ordered
#| tbl-cap: Full results for models testing bounds of $H_2$ (cont.)
# Redefine names
model_names <- list(
  "Trial 3 - Red treatment" = list_fitted[["h_2a"]][["trial_3_red_ordinal"]]
  , "Trial 3 - Blue treatment" = list_fitted[["h_2a"]][["trial_3_blue_ordinal"]]
)

# Change goodness of fit stuff
gm <- list(
  list("raw" = "nobs", "clean" = "Num.Obs.", "fmt" = 1)
  , list("raw" = "r.squared", "clean" = "R<sup>2</sup>", "fmt" = 2)
  , list("raw" = "waic", "clean" = "WAIC", "fmt" = 2)
)
# Make table
modelsummary(
  model_names
  , coef_map = c(
    "b_trial_3_red_stimuli" = "Red yard sign"
    , "b_trial_3_blue_stimuli" = "Blue yard sign"
    , "b_Intercept[1]" = "Intercept 1"
    , "b_Intercept[2]" = "Intercept 2"
    )
  , statistic = "conf.int"
  , conf_level = 0.9
  , gof_map = gm 
  , notes = c(
    "Data source: Pavlovia"
    , "Positive coefficients reflect the logged-odds"
    , " of claiming candidate is Republican."
    , "Numbers inside brackets reflect the 5th and 95th percentile"
    , " of the draws from the posterior distribution."
  )
  , escape = FALSE

)
```

##### Posterior Predictive Check

```{r}
#| label: fig-h-2a-ordinal-trial-3-ppc
#| fig-cap: Posterior Predictive Checks for Ordinal models testing $H_{2a}$, Trial 3
#| fig-subcap:
#|   - Red treatment
#|   - Blue treatment

list_ppc[["h_2a"]][["trial_3_red_ordinal"]]
list_ppc[["h_2a"]][["trial_3_blue_ordinal"]]
```

### $H_3$

#### Main results

```{r}
#| label: tbl-h-3-main
#| tbl-cap: Support of candidate based on partisanship

# Redefine names
model_names <- list(
  "Red stimuli" = list_fitted[["h_3"]][["red"]]
  , "Blue stimuli" = list_fitted[["h_3"]][["blue"]]
)

# Make table
modelsummary(
  model_names
  , coef_map = c(
    "b_trial_1_red_stimuli" = "Red yard sign"
    , "b_trial_1_blue_stimuli" = "Blue yard sign"
    , "b_pid_7" = "Party ID"
    , "b_trial_1_red_stimuli:pid_7" = "Red yard sign × Party ID"
    , "b_trial_1_blue_stimuli:pid_7" = "Blue yard sign × Party ID"
    , "b_attention" = "Attention"
    , "b_age" = "Age"
    , "b_gender_idMale" = "Gender - Male"
    , "b_gender_idNonMbinary" = "Gender - Non-binary"
    , "b_gender_idPrefernottochoose" = "Gender - Prefer not to choose"
    , "b_race_idBlacknonMhispanic" = "Black"
    , "b_race_idHispanice" = "Hispanic"
    , "b_race_idMultipleraces" = "Multiple races"
    , "b_race_idNativeAmericanDAlaskaNativeOrOtherRace" = "Other race"
    , "b_race_idWhitenonMhispanic" = "White"
  )  
  , statistic = "conf.int"
  , conf_level = 0.9
  , gof_map = gm 
  , notes = c(
    "Data source: Pavlovia"
    , "Positive coefficients reflect the logged-odds"
    , " of claiming candidate is Republican."
    , "Numbers inside brackets reflect the 5th and 95th percentile"
    , " of the draws from the posterior distribution."
  )
  , escape = FALSE

)
```

##### Posterior Predictive Check

```{r}
#| label: fig-h-3-ppc
#| fig-cap: Posterior Predictive checks for $H_3$ models
#| fig-subcap:
#|   - Red treatment
#|   - Blue treatment
list_ppc[["h_3"]][["red"]]
list_ppc[["h_3"]][["blue"]]
```

### $H_4$

#### ✅ for over-dispersion

```{r}
#| label: tbl-h-4-dispersion
#| tbl-cap: Checking for over-dispersion of data
df_over_dispersed <- list_df[["cleaned"]][, c("trial_1_time_elapsed")]
names(df_over_dispersed)[names(df_over_dispersed) == 'trial_1_time_elapsed'] <- 'Time Elapsed'
datasummary_skim(
  df_over_dispersed
  , notes = "Data source: Prolific sample."
  , histogram = FALSE
)
```

#### Main results

```{r}
#| label: tbl-h-4
#| tbl-cap: The effect of partisan congruency on time spent looking at a yard sign
# Redefine names
model_names <- list(
  "Red stimuli" = list_fitted[["h_4"]][["red"]]
  , "Blue stimuli" = list_fitted[["h_4"]][["blue"]]
)

# Make table
modelsummary(
  model_names
  , coef_map = c(
    "b_trial_1_red_stimuli" = "Red yard sign"
    , "b_trial_1_blue_stimuli" = "Blue yard sign"
    , "b_pid_7" = "Party ID"
    , "b_trial_1_red_stimuli:pid_7" = "Red yard sign × Party ID"
    , "b_trial_1_blue_stimuli:pid_7" = "Blue yard sign × Party ID"
    , "b_attention" = "Attention"
    , "b_age" = "Age"
    , "b_gender_idMale" = "Gender - Male"
    , "b_gender_idNonMbinary" = "Gender - Non-binary"
    , "b_gender_idPrefernottochoose" = "Gender - Prefer not to choose"
    , "b_race_idBlacknonMhispanic" = "Black"
    , "b_race_idHispanice" = "Hispanic"
    , "b_race_idMultipleraces" = "Multiple races"
    , "b_race_idNativeAmericanDAlaskaNativeOrOtherRace" = "Other race"
    , "b_race_idWhitenonMhispanic" = "White"
  )  
  , statistic = "conf.int"
  , conf_level = 0.9
  , gof_map = gm 
  , notes = c(
    "Data source: Pavlovia"
    , "Positive coefficients reflect the logged-odds"
    , " of claiming candidate is Republican."
    , "Numbers inside brackets reflect the 5th and 95th percentile"
    , " of the draws from the posterior distribution."
  )
  , escape = FALSE

)
```

##### Posterior Predictive Check

```{r}
#| label: fig-h-4-ppc
#| fig-cap: Posterior predictive checks of $H_4$ models
#| fig-subcap:
#|   - Red
#|   - Blue
list_ppc[["h_4"]][["red"]]
list_ppc[["h_4"]][["blue"]]
```

### $H_5$

#### Color proportion descriptive statistics

```{r}
#| label: tbl-yard-sign-summary
#| tbl-cap: Summary statistics
names(df_study_2)[names(df_study_2) == 'red_percent'] <- '% Red'
names(df_study_2)[names(df_study_2) == 'blue_percent'] <- '% Blue'
names(df_study_2)[names(df_study_2) == 'white_percent'] <- '% White'

df_study_2_subset <- df_study_2[, c('% Red', '% Blue', '% White')]
datasummary_skim(
  df_study_2_subset
  , notes = "Data source: CAPD Yard Signs."
  , histogram = FALSE
)

```

#### Main results

```{r}
#| label: tbl-hypothesis-5
#| tbl-cap: The effect of democratic vote share on color choices for yard sign

names(list_fitted[["h_5"]]) <- c("Prop. yard sign Red", "Prop. yard sign Blue")
modelsummary(
  list_fitted[["h_5"]]
  , stat = "conf.int"
  , conf_level = 0.9
  , metrics = c("n.obs", "RMSE")
  , coef_map = c(
    "b_dem_vote_share" = "Democratic vote share (percent)"
    , "b_Intercept" = "Intercept"
  )
  , notes = c(
    "Data source: Center for Political Design and MIT Election Lab"
    , "Positive coefficients reflect the logged-odds"
    , "of claiming candidate is Republican."
    , "Numbers inside brackets reflect the 5th and 95th percentile"
    , "of the draws from the posterior distribution."
  )
  , escape = FALSE

)
```

##### Posterior Predictive Check

```{r}
#| label: fig-hypothesis-5-ppc
#| layout-ncol: 2
#| fig-cap: Posterior Predictive checks for $H_5$
#| fig-subcap:
#|   - Proportion of yard signs red
#|   - Proportion of yard signs blue

list_ppc[["h_5"]][["red"]]$continuous
list_ppc[["h_5"]][["blue"]]$continuous
```

## Copyright Information

<span class="faux-block">{{< fa brands creative-commons >}} 2021-2023 Damon Charles Roberts</span>
<span class="faux-block"><br>All content licensed under {{< fa brands creative-commons >}} {{< fa brands creative-commons-by >}} {{< fa brands creative-commons-sa >}} [Creative Commons CC BY-SA 4.0](https://creativecommons.org/licenses/by-sa/4.0/)</span>