### 2023-12-04 chapter_1/src/manuscript.qmd
drafting the theory chapter a bit

# A theory of processing politically-relevant visual information

In this chapter I will explain a theory of how visually-based political marketing works. This theory encompasses many types of visual information. As you will see, this theory builds upon existing theories of political information processing by incorporating theories from fields like psychology and neuroscience.

While I intend for the theory to be one in which scholars can apply to a number of different types of visual information, the rest of this project will focus on an almost universal and commonly covered form of visual information used in politics: color. Color is so salient in politics that we often describe partisanship by associated colors. For example, Iowa is a red state whereas New York is a blue state. Professional political scientists also tend to use the colors red and blue in the titles of their peer-reviewed papers to indicate the paper's topic being on partisanship. One example would be @goggin_what_2020 with the title of a paper being, "What Goes with Red and Blue? Mapping Partisan and Ideological Associations in the Minds of Voters".

This is symbolic of the field of political science as a whole. While the associations of the colors red and blue emerged in the 2000 Presidential Election [@elving_color_2014] and a handful of scholars [see @schloss_politics_2014;@williams_whats_2022] have documented these strong associations with color and the two dominant parties, there remains to be a theory about how these associations come to be and what the implications are for voters.

Why is this the case? Visual information, at large, is comparatively understudied in the field. The field largely considers information that are quite common and clearly contain political information such as stated policy preferences, conversations with our friends about politics, and text with political information from the news or campaigns. Relative to these more traditionally considered forms of information, visual information is much harder to pin down. Though, for many of us, visual information is ubiquitous, it is highly contextual information. As I will elaborate upon more later in this chapter and in [Chapter 3](../../chapter_3/manuscript.qmd), colors like red and blue can represent different things in different contexts and, as a result, have different effects on other forms of cognitive processing. Not only is visual information quite complex, but theorizing about it becomes harder when we consider how interdisciplinary the theorizing becomes. To not operate in a vacuum, a scholar who wants to theorize about the role of visual information in politics should consider fields such as neuroscience, psychology, marketing, and communication. I agree with @bucy_editors_2021 who discuss this challenge to the study of visual politics when they argue that each of these fields have their own methodological and generative modelling traditions, but they also apply their theories to many different contexts that may not necessarily be easily mappable to political science. These features make it quite difficult for political scientists to take some of these theories and apply it to politics though it may seem to be important low-hanging fruit.

Before examining what these fields would inform political scientists who want to theorize about visual information processing (or in this case, color), I will first discuss existing, and highly interdisciplinary efforts, to theorize about how people process political information in the political context of the United States. I will then explain how I can build upon these existing theories and fold in rich theorizing from these other fields about visual information processing. While doing so, I will present the theory that is the focus of the book. I will refer to this theory as the snap-judgment model of political information processing. Once I explain this theory, I will close this chapter with a discussion about how this contrasts with existing theories of political information processing, but I will more importantly discuss how this theory contributes to these existing models.

## The two canonical models of political information processing

There are two dominant models of how the public in the United States process political information and express attitudes in response to such information. There are, of course, many more than these two; particularly in the fields of survey and public opinion research [see @tourangeau_psychology_2000]. I will focus on the two most commonly used in political science, however, as they synthesize many of these other alternative models and are quite commonly used by political psychologists.

Zaller's [-@zaller_nature_1992;@zaller_simple_1992] "memory" model of political information processing and attitude expression is one of the most popular psychological models in political science. As opposed to viewing citizens as Aristotle's ideal democratic citizens, this "memory" model characterizes citizens as flawed Bayesian updaters. While the application of this model largely focuses on the model's implications for how we express an attitude given existing knowledge, this model characterizes how we "Receive" and "Accept" new information.

In the "memory" model, @zaller_nature_1992 characterizes the public as Bayesian updaters who, when gathering information about politics, will develop a schema about how that information relates to existing information. That is, citizens are not passively consuming political information from political elites; which is much of the focus in @zaller_nature_1992. These "predispositions" that people have influence the degree to which and *how* we organize this new information. The degree to which this new information becomes part of our cognitive schema depends upon these values. As a result, and like other related models at the time that were interested in explaining inconsistency in survey responses, the information that we have in our central cognitive schema for some political object depends upon those predispositions. As a result of these features, not all new political information will prove influential in our expressed political attitudes all of the time.

To help readers view the key contributions of the memory model, I created @fig-zaller-model which emphasizes the key processes involved in processing new information, incorporating, and then using that information to express an attitude. Though extremely simplistic, each node reflects a step in the model. We see that at each step, various predispositions mediate the relationship between the Receive-Accept and Accept-Sample nodes.

```{mermaid}
%%| label: fig-zaller-model
%%| fig-cap: Memory model of political information processing
%%| fig-width: 5
%%{init: {'theme':'base', 'themeVariables':{'primaryColor':'#ffffff', 'primaryBorderColor': '#000000'}}}%%
flowchart TB
    direction TB
    A[New information from a political elite] -->|"Predispositions (e.g., Ideology and sociodemographics)" | B[Receiver adds the new information to their schema]
    B -->|"Predispositions (e.g., Ideology and sociodemographics)"| C[Receiver expresses an attitude based on the schema]
```

The next canonical model of political information processing is Lodge and Taber's [-@lodge_rationalizing_2013] "online" model. This model also views the public as selectively incorporating Bayesian updaters. Like Zaller's [-@zaller_nature_1992] "memory" model, the Lodge and Taber's [-@lodge_rationalizing_2013] "online" model is heavily interdisciplinary. However, they characterize this "Receive" and "Accept" processes of political information as slightly different with significant implications.

One of the psychological theories that Lodge and Taber incorporate into their model of psychological processing is motivated reasoning is referred to as "motivated reasoning" [see @taber_motivated_2006]. The psychological theory of motivated reasoning suggests that the degree to which we gather new information that goes against our pre-existing viewpoints depends on which motivation we presently have. In summarizing this theory, @kunda_case_1990 argues that in contexts where we are motivated to gather information so that we are accurate we are more likely to incorporate all new information; even new information that challenges our pre-existing views. However, according to @kunda_case_1990, we often default to the motivation of selectively gathering information that fits with our pre-existing views in an attempt to have more evidence that we can cite to defend our position to those who may wish to challenge them. This theory of motivated reasoning has been and is still quite popular among social psychologists who examine how members of social groups might justify their negative views of those in another group while maintaining positive views of those in their own group, despite the presence any contradictory evidence.

Though, Zaller's [-@zaller_nature_1992] considers the role of ideology as one of the "predispositions" that influence the reception and acceptance of new information, Lodge and Taber's key innovation is that partisan identity and the degree to which we are motivated to defend the political perspectives of our partisan confederates is **the** pre-disposition that influences the degree to which we consider new information [@lodge_rationalizing_2013]. This perspective predicts that rather than the public appearing to be somewhat regularly biased, but also inconsistent, in expressing their attitudes about something in the face of new information, this model suggests that the public are somewhat predictably going to gather information and express attitudes that align with those of those holding the same partisan identification. This model has been significant for informing theories around political polarization in the United States as well as conversations around the degree to which political persuasion is possible with the public.

Another feature of Lodge and Taber's [-@lodge_rationalizing_2013] model is that it considers the role of affect. Throughout this book, I conceptualize affect as valanced (positive or negative) reactions to some political object. This departs from conceptualizations popular among political scientists where some see it as somewhat synonymous with emotion [see @marcus_affective_2000]. However, I depart with this to align with popular conceptualizations [see @sander_cambridge_2013] that distinguish affect from emotion where affect is a pre-conscious valanced reaction to an object and I conceptualize emotion as a discrete or categorical classification of these valanced reactions which require some degree of conscious processing. In the online model, they include the theory of affect transfer which, when we conceptualize the brain as a network of interconnected nodes [see @collins_spreading-activation_1975], suggests that if an individual dislikes a directive to ICE that limits the number of deportations that they can carry out then we would predict that the same individual would also dislike a directive that extends protection to children of non-documented parents that immigrated to the United States by enrolling in the DACA program due to the person's preferences for restrictive immigration.

Relative to Zaller's [-@zaller_nature_1992] memory model, the online model is much more permissive of context. Besides the rhetorical devices that the particular messenger uses and the pre-dispositions that the receiver has, Lodge and Taber's [-@lodge_rationalizing_2013] model presents a situation where there are many more mediating factors that influence these processes. The degree to which we are motivated to engage with this information in the first place depends on particular motivations that we may have such as to gather accurate information or information that makes us feel more confident that our pre-existing preferences are accurate. As a result of emotion, pre-dispositions such as anxiety can induce this accuracy motivation [@marcus_affective_2000]. Further, we do not just consume this information but the degree to which this information evokes a directionally valanced evaluation depends on what prior information is salient (or "hot" in line with Lodge and Taber's [-@lodge_rationalizing_2013] use of the phrase from the theory of hot cognition) when we consume this new information.

While there are some differences between these models, they also agree on foundational perspectives of how memory is used to generate attitudes. The first is the perspective that information is organized in a schema. This conceptualization borrows significantly from the perspective that the brain is organized as a set of associated nodes in a network. This perspective has been a powerful perspective in a number of fields that study cognition. In this perspective, the cognition works because we have nodes which identify a particular piece of information. These nodes are connected to each other in a mental schema. Or some structured organization of information or memories. These connections between nodes, however are not equally strong. Rather, the strength of these nodes depend on a variety of features. Though the memory and online model vary in which features matter for influencing the strength of these nodes, they still consider the influence of activated confounding information that can make some of these connections between nodes more relevant than others. As a result, both models suggest that we do not just gather information and then just calculate the average between a pros and cons list. Rather, the degree to which a new node is added and used in the future depends on which nodes it is attached to when being encoded as memory.