### 2024-02-19 chapter_4/src/prr/feedback-memo.qmd
re-writing based on feedback for final chapter. Will need to re-write code for outcome variables. Then re-run simulations

---
subtitle: Response memo
format:
  fancy-doc-pdf:
    toc: false
    latex-output-dir: ../out/
    biblatex: true
    #filters:
      #- "_extensions/andrewheiss/wordcount/citeproc.lua"
      #- "_extensions/andrewheiss/wordcount/wordcount.lua"
author:
  - name: Damon C. Roberts
    affiliations:
      - name: Univeristy of Colorado Boulder
    email: damon.roberts-1@colorado.edu
bibliography: 
  - "../../../assets/references.bib"
---

Thank you everyone for your comments. To ensure that I am covering all of your feedback before I submit IRB materials and sort out funding for this final chapter, I want to place the changes in a memo so that you don't have to re-read the pre-registered version of the chapter.